### 2013-03-01 paper/main.tex
Re-arranged structure.

% LaTeX file for a 1 page document
\documentclass[12pt]{proc}

\title{...}

\author{Kevin Peterson\\
\small Mayo Clinic\\
\small Rochester, MN, USA\\
\small \texttt{peterson.kevin@mayo.edu}\\
}

\begin{document}
\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

\section{Conclusions and Future Work}

\bibliographystyle{plain}
\bibliography{bibliography}
hi\cite{lakhani2003open}
\end{document}

### 2013-03-01 paper/publish.sh

@@ -1,0 +1,1 @@
+latexmk -output-directory=build

### 2013-03-03 code/research.py
Added new SQL.

@@ -45,7 +45,7 @@ commit_table = Table('commit', metadata,
 metadata.create_all()
 
 session = requests.Session()
-session.auth = ('kevinpeterson', 'Sept2907$$')
+session.auth = ('kevinpeterson', '*******')
 
 response = urllib2.urlopen('https://raw.github.com/gist/4669395')
 random_words = response.readlines()

### 2013-03-03 paper/publish.sh

@@ -1,1 +1,1 @@
-latexmk -output-directory=build
+latexmk main.tex -output-directory=build -pdf

### 2013-03-03 code/config.py
Added a config file.

@@ -1,0 +1,10 @@
+#Database Config
+db_username=
+db_password=
+db_host=
+db_port=
+db_database=
+
+#GitHub API Config
+github_username=
+github_password=

### 2013-03-03 code/get_summary_data.py
Added summary data retrieval.

@@ -1,1 +1,21 @@
-#
+import MySQLdb as mdb
+
+try:
+	con = mdb.connect('localhost', 'root', '')
+
+	cur = con.cursor()
+
+	with open ("research.sql", "r") as sql_file:
+		sql = sql_file.read()
+
+	cur.execute(sql)
+
+	desc = cur.description
+
+	rows = cur.fetchall()
+
+	for row in rows:
+		print row[0]
+
+finally:
+	con.close()

### 2013-03-03 code/config.py
Added a config file.

@@ -1,10 +1,0 @@
-#Database Config
-db_username=
-db_password=
-db_host=
-db_port=
-db_database=
-
-#GitHub API Config
-github_username=
-github_password=

### 2013-03-03 code/get_summary_data.py
Updated summary data pull.

@@ -1,7 +1,13 @@
 import MySQLdb as mdb
+from config import *
 
 try:
-	con = mdb.connect('localhost', 'root', '')
+	con = mdb.connect(
+		host=db_host,
+		user=db_username,
+		passwd=db_password,
+		db=db_database,
+		port=db_port)
 
 	cur = con.cursor()
 
@@ -16,6 +22,9 @@ try:
 
 	for row in rows:
 		print row[0]
+
+except Exception as e:
+	print e
 
 finally:
 	con.close()

### 2013-03-03 code/research.py

@@ -8,8 +8,10 @@ import random
 from sqlalchemy import create_engine
 from sqlalchemy import MetaData, Column, Table, ForeignKey
 from sqlalchemy import Integer, String, DateTime
+
+from config import *
  
-engine = create_engine('mysql://root@localhost:3306/github', echo=False)
+engine = create_engine('mysql://'+db_username+'@'+db_location+':'+db_port+'/'+db_database, echo=False)
  
 metadata = MetaData(bind=engine)
  
@@ -45,7 +47,7 @@ commit_table = Table('commit', metadata,
 metadata.create_all()
 
 session = requests.Session()
-session.auth = ('kevinpeterson', '*******')
+session.auth = (github_username, github_password)
 
 response = urllib2.urlopen('https://raw.github.com/gist/4669395')
 random_words = response.readlines()

### 2013-03-03 paper/main.tex
Added new SQL.
Updated structure.

@@
-% LaTeX file for a 1 page document
-\documentclass[12pt]{proc}
+% SEng 5852 - Spring 2013 White Paper Review
+\documentclass{article}
 
-\title{...}
+\title{
+Exploring the Open Source Process
+\\[2\baselineskip]
+A Review of \\``A Case Study of Open Source Software Development: The Apache Server''
+\\[2\baselineskip]
+SEng 5852 - Spring 2013 White Paper Review
+\\[2\baselineskip]
+}
+\author{Kevin Peterson\\
+\small \texttt{pete1968@umn.edu}
+}
 
-\author{Kevin Peterson\\
-\small Mayo Clinic\\
-\small Rochester, MN, USA\\
-\small \texttt{peterson.kevin@mayo.edu}\\
-}
+\date{Feb 8, 2013}
 
@@
 \begin{abstract}
-
+my abstract
 \end{abstract}
@@
 
-\section{Conclusions and Future Work}
+\section{Methods}
+
+\section{Results}
+
+\section{Data}
+\begin{table}
+\caption{Table My Table}
+\begin{center}
+\begin{tabular}{rrrrrrr}
+\hline
+\hline
+Variable & Mean & Min & Max & Std. Dev. \\ 
+\hline
+\input{summary_statistics}
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+\section{Acknowledgments}
+
+\section{Reviewed Article}
+In``A Case Study of Open Source Software Development: The Apache Server''\cite{mockus2000case}, the authors examine the popular Open Source Software (OSS) project Apache Server.\footnote{http://httpd.apache.org/} The article defines six research questions which it answers after an examination of the project.
+These questions examine the process, team makeup, and general quality of the software produced in order to determine if an OSS project can compete with commercially developed equivalents.
+This review summarizes the authors' key points, as well as includes my analysis and summary.
+
+\section{Article Key Points}
+
+\subsection{OSS Process} The development process of Apache does not include many of the traditional Software Engineering components, such as project plans, schedules, or even formal design -- attributes it shares with many OSS projects\cite{dibona1999open}.
+Despite this, many OSS projects (including Apache) are thought to be of equal or higher quality than commercial equivalents.\\
+Although there is no defined process, a process does emerge organically.
+Problems are typically reported through a USENET mailing list.
+After a problem or new feature has been identified, how development progresses depends heavily on the ideas of a {\it developer hierarchy} and {\it code ownership}.
+In absence of a dedicated Quality Assurance (QA) team, or even defined system-testing phases, {\it quality and support} is a distributed, community undertaking.
+
+\subsection{Developer Hierarchy} Apache is built around a small set of {\it Core Developers}, followed by {\it Defect Repairers} and {\it Defect Reporters}.
+Each level of this hierarchy brings with it an order of magnitude increase in number of participants.
+The 10-15 {\it Core Developers} contribute around 80\% of the new functionality, while the rest of the 400 code contributors focus primarily on bug fixes.
+The {\it Defect Reporters} were by far the largest group, with over 3000 individuals submitting bug reports.
+
+\subsection{Code Ownership} {\it Core Developers} rely on mutual trust in each other's expertise to avoid having to strictly partition the code base to avoid conflicts.
+If this expertise and trust is sufficient, {\it Core Developers} do not need to segment or limit commit rights amongst themselves.
+The set of core developers is surrounded by ``satellite'' groups, which then own the development of additional functionality.
+As stated in the article, this effectively breaks the OSS project into a collection of related OSS projects.
+This is necessary because a small core of developers has a natural limit to their code output.
+
+\subsection{Quality and Support} Given the same amount of testing, OSS will have a lower defect density than commercial software.
+There are several reasons for this.
+Notably, most OSS is written by passionate domain experts.
+Also, given the sheer number of people that review changes, defects are more likely to be caught without having to go through formal testing cycles.
+Regarding customer support, a rapid response to problems can be obtained because OSS is not bound to release schedules in the same way commercial software is.
+``Patches'' may be released at any time, by any member of the community.
+
+\section{Analysis of Key Points}
+
+\subsection{OSS Process}
+The authors suggest pairing these OSS ideas with a more structured commercial software process, and I find that idea very interesting.
+I am not convinced that simply ``evolving'' without a high level design or project plan is advisable, even though there are success stories.
+I am unsure as to whether it is {\it in spite of} or {\it because of} the process that this project was successful.
+There seems to be very real limitations in this process, such as code base size.
+The idea of the ``community as an asset'' is well described as very useful in reviewing code and finding defects.
+
+\subsection{Developer Hierarchy}
+Skill and experience of developers is one of the key factors in OSS quality\cite{hedberg2007assuring}.
+If OSS can be the vehicle that brings together a talented core of domain experts, that in itself is an accomplishment regardless of process.
+Also, participation expands as the amount of potential impact decreases -- which is a logical but interesting point.
+Bug reporters, for example, do not have to be necessarily trusted with core functionality, but can still contribute a valuable service -- and there can be many of them.
+
+\subsection{Code Ownership}
+Collective code ownership is an important part of the Extreme Programing (XP) process, and I believe is generally useful regardless of process or software license.
+It is interesting that lacking a defined process, OSS comes to the same conclusion on this point as XP, as XP explicitly lists collective code ownership as one of their core values\cite{beck1999embracing}.
+
+\subsection{Quality and Support}
+Quality is governed by the group as a whole, primarily through mailing list discussions and the opportunity for any change/patch to be reviewed by the community.
+This peer scrutiny, along with the pride of work that comes with working in a domain that the contributors are passionate about, are highlighted as important factors in a lower (pre-test) defect count.
+However, I believe that the intrinsic reward\cite{lakhani2003open} felt by many from supporting OSS projects will be a challenge to duplicate in commercial offerings.
+
+\section{Summary}
+OSS development, as mentioned in the article, is its own unique process.
+It is interesting to see, given a set of developers and a shared project, the process that evolves.
+A process that has evolved organically (as opposed to one dictated by management) can perhaps give us a better understanding of how we as Software Engineers really work.\\
+The article shows that even though the OSS process may appear to be unstructured and unorganized, in reality roles and hierarchies appear.
+This structure may be organically created -- but is structure none the less.\\
+I believe that the ideas of Code Ownership were highlighted especially well in this article.
+The authors conclude that these ideas are successful because of shared respect for experience and ability.
+This point is powerful, and something that all processes should strive for.\\
+Data analysis was also done comparing Apache to several similar commercial products.
+This data helps to confirm the hypotheses of the authors, and makes a very strong argument for the viability of this process.
 
@@
 \bibliography{bibliography}
-hi\cite{lakhani2003open}
+
 \end{document}

### 2013-03-03 code/get_summary_data.py

@@ -1,4 +1,5 @@
 import MySQLdb as mdb
+from decimal import Decimal
 from config import *
 
 try:
@@ -21,7 +22,10 @@ try:
 	rows = cur.fetchall()
 
 	for row in rows:
-		print row[0]
+		for i in range(0,5):
+			print row[i],
+			print ' \\\\' if i == 4 else ' & ',
+		print
 
 except Exception as e:
 	print e

### 2013-03-03 code/plot.py
Added some plots.

@@ -1,0 +1,32 @@
+import sys
+import matplotlib.pyplot as plt
+import MySQLdb
+
+
+db = MySQLdb.connect(host="localhost", user="root", passwd="", db="github")
+cursor = db.cursor()
+
+query = '''
+select forks, count(*) FROM github.repository r inner join github.commit c on r.id = c.repository_id group by repository_id
+'''
+cursor.execute(query)
+
+result = cursor.fetchall()
+
+t = []
+s = []
+
+for record in result:
+  t.append(record[0])
+  s.append(record[1])
+
+plt.plot(t, s, 'ko')
+plt.ylim(0,3000)
+plt.xlim(0,400)
+plt.title("Forks V Commits")
+plt.xlabel("Forks",fontsize=12)
+plt.ylabel("Commits",fontsize=12)
+plt.show()
+#F = gcf()
+#DPI = F.get_dpi()
+#F.savefig('plot.png',dpi = (80))

### 2013-03-03 code/plots.py

@@ -1,0 +1,9 @@
+from plotter import Histograph
+
+committers_query = '''select count(distinct committer) from github.commit c group by repository_id'''
+committers_histogram = Histograph(committers_query, "Committers", "Repositories", (1,25))
+committers_histogram.plot("committers_histogram")
+
+issue_reporters_query = '''select count(distinct creator) from github.issue group by repository_id'''
+issue_reporters_histogram = Histograph(issue_reporters_query, "Issue Reporters", "Repositories")
+issue_reporters_histogram.plot("issue_reporters_histogram")

### 2013-03-03 code/plotter.py

@@ -1,0 +1,35 @@
+import sys
+import MySQLdb
+from config import *
+from pylab import *
+from matplotlib import rc
+
+class Histograph:
+	def __init__(self, sql, xlabel, ylabel, range=None):
+		self.sql = sql
+		self.xlabel = xlabel
+		self.ylabel = ylabel
+		self.range = range
+
+	def plot(self, filename):
+		db = MySQLdb.connect(host=db_host, user=db_username, passwd=db_password, db=db_database)
+		cursor = db.cursor()
+
+		cursor.execute(self.sql)
+
+		result = cursor.fetchall()
+
+		y = []
+
+		for record in result:
+		  y.append(record[0])
+
+		import matplotlib.pyplot as plt
+
+		plt.hist(y, range=self.range)
+		plt.xlabel(self.xlabel)
+		plt.ylabel(self.ylabel)
+
+		F = gcf()
+		F.savefig("../paper/images/"+filename+".png")
+		plt.close()

### 2013-03-04 code/plots.py
Added new queries.

@@ -1,9 +1,56 @@
-from plotter import Histograph
+from plotter import Histograph, ScatterPlot
 
 committers_query = '''select count(distinct committer) from github.commit c group by repository_id'''
 committers_histogram = Histograph(committers_query, "Committers", "Repositories", (1,25))
 committers_histogram.plot("committers_histogram")
 
+committers_percentage_query = '''
+select 
+	cast(cast( (count(*) / commits.cnt ) as decimal(2,2)) * 100 as unsigned integer)
+from
+    (select 
+        count(*) as cnt, c1.repository_id
+    from
+        github.commit c1
+    group by c1.repository_id) commits
+        inner join
+    github.commit c1 ON c1.repository_id = commits.repository_id
+group by c1.repository_id , c1.committer
+'''
+committers_percentage_histogram = Histograph(committers_percentage_query, "% Of Project Commits", "Committers")
+committers_percentage_histogram.plot("committers_percentage_histogram")
+
 issue_reporters_query = '''select count(distinct creator) from github.issue group by repository_id'''
-issue_reporters_histogram = Histograph(issue_reporters_query, "Issue Reporters", "Repositories")
+issue_reporters_histogram = Histograph(issue_reporters_query, "Issue Reporters", "Repositories", (1,40))
 issue_reporters_histogram.plot("issue_reporters_histogram")
+
+issue_close_time_query = '''
+select cast(avg(datediff(close_date, open_date)) as unsigned) FROM github.issue issues 
+where issues.close_date is not null 
+group by repository_id
+'''
+issue_close_time_histogram = Histograph(issue_close_time_query, "Average Issue Close Time (Days)", "Repositories", (0,140))
+issue_close_time_histogram.plot("issue_close_time_histogram")
+
+watcher_forks_scatterplot_query = '''select watchers, forks from github.repository'''
+watcher_forks_scatterplot = ScatterPlot(watcher_forks_scatterplot_query, "Watchers", "Forks", (0,1000), (0,150))
+print watcher_forks_scatterplot.plot("watcher_forks_scatter_scatterplot")
+
+issue_reporters_committers_query = '''
+select 
+    issues.reporters_count, commits.committers_count
+from
+    (select 
+        repository_id, count(distinct committer) as committers_count
+    from
+        github.commit
+    group by repository_id) commits
+        inner join
+    (select 
+        repository_id, count(distinct creator) as reporters_count
+    FROM
+        github.issue
[…]

### 2013-03-04 code/plotter.py

@@ -2,7 +2,50 @@ import sys
 import MySQLdb
 from config import *
 from pylab import *
-from matplotlib import rc
+from matplotlib import rc		
+import matplotlib.pyplot as plt
+import numpy
+
+class ScatterPlot:
+	def __init__(self, sql, xlabel, ylabel, xrange=None, yrange=None):
+		self.sql = sql
+		self.xlabel = xlabel
+		self.ylabel = ylabel
+		self.xrange = xrange
+		self.yrange = yrange
+
+	def plot(self, filename):
+		db = MySQLdb.connect(host=db_host, user=db_username, passwd=db_password, db=db_database)
+		cursor = db.cursor()
+
+		cursor.execute(self.sql)
+
+		result = cursor.fetchall()
+
+		x = []
+		y = []
+
+		for record in result:
+			x.append(record[0])
+			y.append(record[1])
+
+		fig = plt.figure() 
+		m,b = numpy.polyfit(x,y,1)
+		print("Slope: %s, B: %s") % (m,b)
+
+		plt.plot(x,y,'bo',x,m*numpy.array(x)+b,'-k',linewidth=2)
+
+		plt.scatter(x,y)
+		plt.xlabel(self.xlabel)
+		plt.ylabel(self.ylabel)
+		plt.xlim(self.xrange)
+		plt.ylim(self.yrange)
+	
+		F = gcf()
+		F.savefig("../paper/images/"+filename+".png")
+		plt.close(fig)
+
+		return numpy.corrcoef(x,y)
 
 class Histograph:
 	def __init__(self, sql, xlabel, ylabel, range=None):
@@ -19,17 +62,16 @@ class Histograph:
 
 		result = cursor.fetchall()
 
-		y = []
+		x = []
 
 		for record in result:
-		  y.append(record[0])
+		  x.append(record[0])
 
-		import matplotlib.pyplot as plt
-
-		plt.hist(y, range=self.range)
+		fig = plt.figure() 
+		plt.hist(x, range=self.range)
 		plt.xlabel(self.xlabel)
 		plt.ylabel(self.ylabel)
 
 		F = gcf()
 		F.savefig("../paper/images/"+filename+".png")
-		plt.close()
+		plt.close(fig)

### 2013-03-05 paper/main.tex
Added some plots.
Progress on the paper...

--- 2013-03-01 paper/main.tex
@@
-% LaTeX file for a 1 page document
-\documentclass[12pt]{proc}
+% SEng 5852 - Spring 2013 White Paper Review
+\documentclass{proc}
+\usepackage{graphicx}
 
-\title{...}
+\title{
+The GitHub Open Source Development Process
+\author{Kevin Peterson\\
+Department of Information Technology\\
+Clinical Informatics Support Systems\\
+Mayo Clinic\\
+200 1st Street\\
+Rochester, MN  55905\\
+\small \texttt{peterson.kevin@mayo.edu}
+}
+}
 
-\author{Kevin Peterson\\
-\small Mayo Clinic\\
-\small Rochester, MN, USA\\
-\small \texttt{peterson.kevin@mayo.edu}\\
-}
+\date{}
 
@@
 \begin{abstract}
-
+my abstract
 \end{abstract}
 
+\noindent \textbf{Keywords.} list of keywords
+
 \section{Introduction}
+Open Source Software (OSS) has fundamentally changed how we view the Software Development Process\cite{raymond1999cathedral}. OSS projects are not only viable, but successful and thriving.
 
-\section{Conclusions and Future Work}
+A case study of the Apache Server project\cite{mockus2000case} shows that a dedicated community can produce software that rivals or exceeds commercial offerings.
+This study showed that even in the unstructured context of OSS, certain structures, hierarchies, and codes of conduct emerge.
+Contrast the Apache Server project to a successful project on GitHub.
+Although the spirit and intent may be the same, the tool set is drastially different.
+The intent of this paper is to explore how GitHub facilitates this process, and also how it may cause it to evolve.
+
+\subsection{Hypothesis}
+\emph{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
+In Open Source Software, a feeling of community belonging can be intrinsicly motivating to developers\cite{lakhani2003hackers}.
+This feeling of belonging can be expressed passively, as a \emph{watch}, or actively, as a \emph{fork} of a repository.
+As \emph{watches} don't necessarily signal intent to contribute, a positive direction for an OSS project is to not only increase the number of \emph{watchers}, but also the number of \emph{forks}.
+As both signify a general general interest in the project, it is expected that there be some correllation between the two.\\
+
+\emph{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
+In OSS, a project \emph{fork} has at times carried negative connotations, and has even been referred to as a hazzard\cite{kogut2001open}.
+In the context of Git and GitHub, however, a \emph{fork} is a positive occurence for a project, as it signifies greater project involvement.
+As Git allows for easy merging of \emph{forks}, code contributions in the form of defect fixes can be incorporated quickly.
+Because of this, it is expected that \emph{Issue Resolution Time} will decrease as the number of \emph{forks} increases.\\
+
+The Apache Project noted that a rapid response to problems can be obtained because OSS is not bound to release schedules in the same way commercial software is.
+``Patches'' may be released at any time, by any member of the community\cite{mockus2000case}.
+``Patches"" in the context of GitHub could be equated to Pull Requests.\\
+
+\emph{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude.}\\
+Research into the Apache Server project observed that there were far more \emph{Issue Reporters} than there were code \emph{Committers}\cite{mockus2000case}.
+The Apache Server project is an excellent case study in this type of Developer Hierarchy.
+Apache is built around a small set of {\it Core Developers}, followed by {\it Defect Repairers} and {\it Defect Reporters}.
+Each level of this hierarchy brings with it an order of magnitude increase in number of participants.
+The 10-15 {\it Core Developers} contribute around 80\% of the new functionality, while the rest of the 400 code contributors focus primarily on bug fixes.
+The {\it Defect Reporters} were by far the largest group, with over 3000 individuals submitting bug reports.\\
+
+Because issue reporting is of low risk to the code base, but has potentially high value, it is a perfect way for large numbers of people to contribute.
+It is expected that the findings of the Apache Server project research will hold true for GitHub projects as well.
+
+\section{Methods}
+
+\section{Results}
+\emph{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
+\includegraphics[height=3in,width=3in]{images/issue_close_time_forks_scatterplot.png}
+
+\emph{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
+\includegraphics[height=3in,width=3in]{images/watcher_forks_scatterplot.png}
+
+\emph{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude.}\\
+\includegraphics[height=3in,width=3in]{images/issue_reporters_histogram.png}
+\includegraphics[height=3in,width=3in]{images/committers_histogram.png}
+\includegraphics[height=3in,width=3in]{images/issue_reporters_committers_scatterplot.png}
+
+\section{Acknowledgments}
 

### 2013-03-05 code/plots.py

@@ -34,7 +34,7 @@ issue_close_time_histogram.plot("issue_close_time_histogram")
 
 watcher_forks_scatterplot_query = '''select watchers, forks from github.repository'''
 watcher_forks_scatterplot = ScatterPlot(watcher_forks_scatterplot_query, "Watchers", "Forks", (0,1000), (0,150))
-print watcher_forks_scatterplot.plot("watcher_forks_scatter_scatterplot")
+print watcher_forks_scatterplot.plot("watcher_forks_scatterplot")
 
 issue_reporters_committers_query = '''
 select 
@@ -54,3 +54,22 @@ from
 '''
 issue_reporters_committers_scatterplot = ScatterPlot(issue_reporters_committers_query, "Committers", "Issue Reoprters", (0,140), (0,140))
 print issue_reporters_committers_scatterplot.plot("issue_reporters_committers_scatterplot")
+
+issue_close_time_forks_query = '''
+select 
+    cast(issues.close_time as unsigned), repository.forks
+from
+    github.repository repository
+        inner join
+    (select 
+        avg(DATEDIFF(close_date, open_date)) as close_time,
+            repository_id
+    from
+        github.issue issue
+    where
+        issue.close_date is not null
+    group by repository_id) issues ON repository.id = issues.repository_id
+group by repository_id
+'''
+issue_close_time_forks_scatterplot = ScatterPlot(issue_close_time_forks_query, "Close Time", "Forks", (0,200), (0,200))
+print issue_close_time_forks_scatterplot.plot("issue_close_time_forks_scatterplot")

### 2013-03-05 code/plotter.py

@@ -36,6 +36,10 @@ class ScatterPlot:
 		plt.plot(x,y,'bo',x,m*numpy.array(x)+b,'-k',linewidth=2)
 
 		plt.scatter(x,y)
+
+		correlation_coefficient = numpy.corrcoef(x,y)[0,1]
+		title('Correlation Coefficient: ' + str(correlation_coefficient))
+
 		plt.xlabel(self.xlabel)
 		plt.ylabel(self.ylabel)
 		plt.xlim(self.xrange)
@@ -45,7 +49,7 @@ class ScatterPlot:
 		F.savefig("../paper/images/"+filename+".png")
 		plt.close(fig)
 
-		return numpy.corrcoef(x,y)
+		return correlation_coefficient
 
 class Histograph:
 	def __init__(self, sql, xlabel, ylabel, range=None):

### 2013-03-05 paper/main.tex
Added more sources and some research questions.

@@
 
-\emph{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude.}\\
+\emph{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude, but the number of Committers will remain small.}\\
 Research into the Apache Server project observed that there were far more \emph{Issue Reporters} than there were code \emph{Committers}\cite{mockus2000case}.
@@
 Because issue reporting is of low risk to the code base, but has potentially high value, it is a perfect way for large numbers of people to contribute.
-It is expected that the findings of the Apache Server project research will hold true for GitHub projects as well.
+It is expected that the findings of the Apache Server project research will hold true for GitHub projects as well.\\
+
+\subsection{Research Questions}
+
+\emph{Research Question 1: There will be more Issue Reporters than Committers by an order of magnitude, but the number of Committers will remain small.}\\
+A smal core of developers, however, seems to be a historial fact of OSS development\cite{mockus2000case,mockus2000two,krishnamurthy2002cave}.
+At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers.
+Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?
+
+\emph{Research Question 2: What is the GitHub Philosophy?}\\
+The socal aspects of GitHub are an important part of the experience\cite{dabbish2012social}.
+User interactions and social pressures can drive OSS development in interesting ways.
+Social aspects of OSS development have existed before GitHub, with mailing lists\cite{mockus2000case}, gift giving mechanisms\cite{bergquist2008power}, and other code hosting sites.
+What is GitHub doing to facilitate change on OSS development, and how does it see OSS development changing moving forward?
 
@@
 \emph{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
-\includegraphics[height=3in,width=3in]{images/issue_close_time_forks_scatterplot.png}
+\includegraphics[height=3in,width=3in]{images/watcher_forks_scatterplot.png}
 
 \emph{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
-\includegraphics[height=3in,width=3in]{images/watcher_forks_scatterplot.png}
+\includegraphics[height=3in,width=3in]{images/issue_close_time_forks_scatterplot.png}
 
@@
 \includegraphics[height=3in,width=3in]{images/issue_reporters_committers_scatterplot.png}
+
+\subsection{Other Findings}
+\emph{The Git SCM }\\
+
+\emph{The GitHub Process applies to more than code artifacts}\\
 

### 2013-03-06 paper/main.tex
Started on the abstract.

@@
 \usepackage{graphicx}
+\setlength\parindent{0pt}
 
@@
 \begin{abstract}
-my abstract
+Open Source Software (OSS) has produced many successful projects.
+The development process by which these projects are produced is generally unstructured compared to commercial software -- but a definite pattern does arrive, and is no less a pattern.
+GitHub, a popular OSS code hosting website, and Git, the site's SCM of choice, may have the potential to fundamentally change ths process.\\
+By analyzing a subset of GitHub repositories, this report will show how GitHub has influenced some very intrinsic aspects of traditional OSS develoment, such as developer hierarchies and issue close velicity.
+We find that...
+TODO
 \end{abstract}
 
-\noindent \textbf{Keywords.} list of keywords
+\noindent \textbf{Keywords.} Git, GitHub, OpenSource
 
@@
 
-\emph{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude, but the number of Committers will remain small.}\\
+\emph{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude}\\
 Research into the Apache Server project observed that there were far more \emph{Issue Reporters} than there were code \emph{Committers}\cite{mockus2000case}.
@@
 
-\emph{Research Question 1: There will be more Issue Reporters than Committers by an order of magnitude, but the number of Committers will remain small.}\\
+\emph{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
 A smal core of developers, however, seems to be a historial fact of OSS development\cite{mockus2000case,mockus2000two,krishnamurthy2002cave}.
@@
 
+\emph{Research Question 3: Can GitHub be used for more that code artifacts?}\\
+The Object Managment Group\textregistered\footnote{http://www.omg.org} computer industry consortium focused on technology standards.
+Potential standard summission teams must go through a vetting process\cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard.
+Gathering and recording this feedback is a challenge, but GitHub's issue tracking system may be able to streamline the process.
+
 \section{Methods}
@@
 \section{Results}
+\subsection{Hypotheses}
 \emph{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
@@
 
-\subsection{Other Findings}
-\emph{The Git SCM }\\
+\subsection{Research Questions}
+\emph{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
 
-\emph{The GitHub Process applies to more than code artifacts}\\
+\emph{Research Question 2: What is the GitHub Philosophy?}\\
+
+\emph{Research Question 3: Can GitHub be used for more that code artifacts?}\\
 

### 2013-03-07 paper/main.tex
Added a "Linux Law" section.
Started on summary statistics.

@@
 \usepackage{graphicx}
+\usepackage{mathtools}
+\usepackage{algorithmicx}
+\usepackage{algpseudocode}
 \setlength\parindent{0pt}
@@
 
+This is, ultimetely, a proof of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub.
+In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.
+
 \emph{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude}\\
@@
 \section{Methods}
+Data collection was done using the GitHub REST API\footnote{http://developer.github.com/v3/}.
+A random selection of 1000 repositories was selected using the following algorithm:\footnote{https://github.com/kevinpeterson/github-process-research/blob/master/code/research.py}
+
+\begin{algorithmic}
+\While{$i < 1000$} 
+\State $i \gets i + 1$
+\State $word \gets random(word\_list)$
+\State $repos \gets git\_search(word)$
+\State $j \gets random(0...size(repos))$
+\State $repo \gets repos_j$
+\State \Call{process}{repo}
+\EndWhile
+\end{algorithmic}\footnote
+
+$\displaystyle\frac{\sum\limits_{i=1}^{n} f( \left\{{i_j,i_{j+1},...,i_{j+k}  }\right\} ) } {n} $.
 
 \section{Results}
+
+\subsection{Summary Statistics}
+\begin{table}[!ht]
+\begin{center}
+\caption{GitHub Summary Statistics}
+\begin{tabular}{rrrrrrr}
+\hline
+Variable & Mean & Min & Max & Std. Dev. \\
+\hline
+\input{summary_statistics}
+\hline
+\end{tabular}
+N = 100 Repositories
+\end{center}
+\end{table}
+
 \subsection{Hypotheses}

### 2013-03-07 code/get_summary_data.py

@@ -21,9 +21,17 @@ try:
 
 	rows = cur.fetchall()
 
+	col_processor = {
+		0:lambda x: x,
+		1:lambda x: round(x,2),
+		2:lambda x: int(x),
+		3:lambda x: int(x),
+		4:lambda x: round(x,2)
+	}
+
 	for row in rows:
 		for i in range(0,5):
-			print row[i],
+			print col_processor[i](row[i]),
 			print ' \\\\' if i == 4 else ' & ',
 		print
 

### 2013-03-09 code/plotter.py
Started a PieChart function.

@@ -6,6 +6,28 @@ from matplotlib import rc
 import matplotlib.pyplot as plt
 import numpy
 
+class PieChart:
+	def __init__(self, sql):
+		self.sql = sql
+
+	def plot(self, filename):
+		db = MySQLdb.connect(host=db_host, user=db_username, passwd=db_password, db=db_database)
+		cursor = db.cursor()
+
+		cursor.execute(self.sql)
+
+		result = cursor.fetchall()
+
+		x = []
+		y = []
+
+		for record in result:
+			x.append(record[0])
+
+		plt.pie([4,4,10,4,4], autopct='%.2f')
+
+		plt.show()
+
 class ScatterPlot:
 	def __init__(self, sql, xlabel, ylabel, xrange=None, yrange=None):
 		self.sql = sql

### 2013-03-09 paper/main.tex
Added some CTS2 text.
Updates to sql and paper.

@@
 \usepackage{mathtools}
+\usepackage{fixltx2e}
 \usepackage{algorithmicx}
 \usepackage{algpseudocode}
+\MakeRobust{\Call}
 \setlength\parindent{0pt}
@@
 \State $repos \gets git\_search(word)$
-\State $j \gets random(0...size(repos))$
+\State $j \gets \Call{random}{0... \Call{size}{repos}}$
 \State $repo \gets repos_j$
-\State \Call{process}{repo}
+\State \Call{store}{repo}
 \EndWhile
-\end{algorithmic}\footnote
+\end{algorithmic}
 
-$\displaystyle\frac{\sum\limits_{i=1}^{n} f( \left\{{i_j,i_{j+1},...,i_{j+k}  }\right\} ) } {n} $.
+The \textsc{store} function accepts as an input a given random repository. The purpose of this function is to persist the given repository to a database for further analysis.
+
+\begin{figure}
+\includegraphics[height=3in,width=3in]{images/er.png}
+\caption{GitHub Date Entity Relationship diagram}
+\end{figure}
+
+$\displaystyle\frac{\sum\limits_{i=1}^{n} f(  S_{ij} ) } {n} $.
 
@@
 \begin{center}
-\caption{GitHub Summary Statistics}
 \begin{tabular}{rrrrrrr}
@@
 N = 100 Repositories
+\caption{GitHub Summary Statistics}
 \end{center}
@@
 \emph{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
+\begin{figure}
 \includegraphics[height=3in,width=3in]{images/watcher_forks_scatterplot.png}
+\caption{Repostitory watchers and forks}
+\end{figure}
 
 \emph{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
+\begin{figure}
 \includegraphics[height=3in,width=3in]{images/issue_close_time_forks_scatterplot.png}
+\caption{Repostitory issue close time and number of forks}
+\end{figure}
 
 \emph{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude.}\\
+\begin{figure}
 \includegraphics[height=3in,width=3in]{images/issue_reporters_histogram.png}
+\caption{Repostitory issue reporters}
+\end{figure}
+\begin{figure}
 \includegraphics[height=3in,width=3in]{images/committers_histogram.png}
+\caption{Repostitory committers}
+\end{figure}
+\begin{figure}
 \includegraphics[height=3in,width=3in]{images/issue_reporters_committers_scatterplot.png}
+\caption{Repostitory issue reporters and committers}
+\end{figure}
 
@@
 \emph{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
+\begin{figure}
+\includegraphics[height=3in,width=3in]{images/committers_percentage_pie_chart.png}
+\caption{Individual committer commit percentage per repository}
+\end{figure}
 
@@
 \emph{Research Question 3: Can GitHub be used for more that code artifacts?}\\
+As part of the OMG\textregistered standardization process, the CTS2 Specification used GitHub issue tracker to track specification changes.
 

### 2013-03-09 code/plots.py

@@ -1,6 +1,6 @@
-from plotter import Histograph, ScatterPlot
+from plotter import Histograph, ScatterPlot, PieChart
 
-committers_query = '''select count(distinct committer) from github.commit c group by repository_id'''
+committers_query = '''select count(distinct committer) from commit c group by repository_id'''
 committers_histogram = Histograph(committers_query, "Committers", "Repositories", (1,25))
 committers_histogram.plot("committers_histogram")
 
@@ -11,28 +11,28 @@ from
     (select 
         count(*) as cnt, c1.repository_id
     from
-        github.commit c1
+        commit c1
     group by c1.repository_id) commits
         inner join
-    github.commit c1 ON c1.repository_id = commits.repository_id
+    commit c1 ON c1.repository_id = commits.repository_id
 group by c1.repository_id , c1.committer
 '''
 committers_percentage_histogram = Histograph(committers_percentage_query, "% Of Project Commits", "Committers")
 committers_percentage_histogram.plot("committers_percentage_histogram")
 
-issue_reporters_query = '''select count(distinct creator) from github.issue group by repository_id'''
+issue_reporters_query = '''select count(distinct creator) from issue group by repository_id'''
 issue_reporters_histogram = Histograph(issue_reporters_query, "Issue Reporters", "Repositories", (1,40))
 issue_reporters_histogram.plot("issue_reporters_histogram")
 
 issue_close_time_query = '''
-select cast(avg(datediff(close_date, open_date)) as unsigned) FROM github.issue issues 
+select cast(avg(datediff(close_date, open_date)) as unsigned) FROM issue issues 
 where issues.close_date is not null 
 group by repository_id
 '''
 issue_close_time_histogram = Histograph(issue_close_time_query, "Average Issue Close Time (Days)", "Repositories", (0,140))
 issue_close_time_histogram.plot("issue_close_time_histogram")
 
-watcher_forks_scatterplot_query = '''select watchers, forks from github.repository'''
+watcher_forks_scatterplot_query = '''select watchers, forks from repository'''
 watcher_forks_scatterplot = ScatterPlot(watcher_forks_scatterplot_query, "Watchers", "Forks", (0,1000), (0,150))
 print watcher_forks_scatterplot.plot("watcher_forks_scatterplot")
 
@@ -43,13 +43,13 @@ from
     (select 
         repository_id, count(distinct committer) as committers_count
     from
-        github.commit
+        commit
     group by repository_id) commits
[…]

### 2013-03-09 code/plotter.py

@@ -18,15 +18,27 @@ class PieChart:
 
 		result = cursor.fetchall()
 
-		x = []
-		y = []
+		ranges = {'0-5':0,'5-10':0,'10-50':0,'50-95':0,'95-100':0}
 
 		for record in result:
-			x.append(record[0])
+			percentage = record[0]
+			if 0 <= percentage < 5:
+				ranges['0-5'] += 1
+			elif 5 <= percentage < 10:
+				ranges['5-10'] += 1
+			elif 10 <= percentage < 50:
+				ranges['10-50'] += 1
+			elif 50 <= percentage < 95:
+				ranges['50-95'] += 1
+			elif 95 <= percentage <= 100:
+				ranges['95-100'] += 1
 
-		plt.pie([4,4,10,4,4], autopct='%.2f')
+		fig = plt.figure()
+		plt.pie(ranges.values(), labels=[key + "%" for key in ranges.keys()], autopct='%.2f%%', shadow=True)
 
-		plt.show()
+		F = gcf()
+		F.savefig("../paper/images/"+filename+".png")
+		plt.close(fig)
 
 class ScatterPlot:
 	def __init__(self, sql, xlabel, ylabel, xrange=None, yrange=None):

### 2013-03-09 code/research.py

@@ -11,7 +11,7 @@ from sqlalchemy import Integer, String, DateTime
 
 from config import *
  
-engine = create_engine('mysql://'+db_username+'@'+db_location+':'+db_port+'/'+db_database, echo=False)
+engine = create_engine('mysql://'+db_username+'@'+db_host+':'+str(db_port)+'/'+db_database, echo=False)
  
 metadata = MetaData(bind=engine)
  
@@ -135,4 +135,4 @@ def crawl(sample_size):
 			print e
 
 
-crawl(313)
+crawl(999)

### 2013-03-10 paper/main.tex
Added some more references.

@@
 At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers.
-Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?
+Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?\\
 
-\emph{Research Question 2: What is the GitHub Philosophy?}\\
+\emph{Research Question 2: How is GitHub changing the OSS process?}\\
 The socal aspects of GitHub are an important part of the experience\cite{dabbish2012social}.
@@
 Social aspects of OSS development have existed before GitHub, with mailing lists\cite{mockus2000case}, gift giving mechanisms\cite{bergquist2008power}, and other code hosting sites.
-What is GitHub doing to facilitate change on OSS development, and how does it see OSS development changing moving forward?
+What is GitHub doing to facilitate change on OSS development, and how does it see OSS development changing moving forward?\\
 
@@
 \section{Methods}
+\subsection{Collection and Storage}
 Data collection was done using the GitHub REST API\footnote{http://developer.github.com/v3/}.
@@
 \caption{GitHub Date Entity Relationship diagram}
+\label{fig:er_diagram}
 \end{figure}
 
-$\displaystyle\frac{\sum\limits_{i=1}^{n} f(  S_{ij} ) } {n} $.
+The database schema to store the data is decribed in \ref{fig:er_diagram}.
+
+\subsection{Analysis}
+
+Where the Set $S$ is the random GitHub repository sample.
+Each randomly collected repository $S_i$ contains a Set of attriubtes $A$.
+Thus, a set of attributes for a given repository can be referred to as $A_{S_i}$.
+
+Given the total set of attributes $A$, each individual attribute set can be denoted as $A_j$.
+Each individual attribute processed by some aggregate function $f$, or $f(  A_j )$.
+For this analysis, the aggregate functions are: \textsc{mean}, \textsc{min}, \textsc{max}, \textsc{stddev}.
+
+Results are then averaged over all repositories, yeilding:
+\[ \frac{\sum\limits_{i=1}^{n} f(  A_{S{_i}j}  ) } {n} \]
+
+The individual attrubites of $A$, or $A_j$, are described below:\\
+\textit{Issues}\\
+The number of issues posted to a repository's issue tracker\footnote{https://github.com/blog/411-github-issue-tracker}.
+
+\textit{Commits}\\
+The number of individual source code commits to a repository.
+
+\textit{Closed Issues}\\
+The number of issues in a repository in a state of \textsc{closed} at the time of processing.
+
+\textit{Open Issues}\\
+The number of issues in a repository in a state of \textsc{open} at the time of processing.
+
+\textit{Issue Close Time}\\
+The average time taken to close a given issue in a repository. Non-\textsc{closed} issues were not couted.
+
+The average is calculated using the following method. Let $d(x)$ be a function that converts seconds to a Date, and let $s(x)$ be a function that converts a Date to seconds.
+
+\[ d\left( \frac{\sum\limits_{i=1}^{n} s(i_{\textrm{closed}}) - s(i_{\textrm{opened}})  } {n} \right) \]
+
+\textit{Issue Creators}\\
+
+\textit{Committers}\\
 
@@
 \emph{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
+No significant correlation was found between the number of repository forks and issue resolution time.
+It is important to note, however, that issue resolution time is not necessarily related to software product quality in the context of GitHub.
+We can assert that for serveral reasons.
+\textit{First}, GitHub issues are not necessarily product defects.
+Issues may be feature requests.
+TODOs, or general support questions.
+\textit{second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all.
+\textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub -- something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric \cite{phamcreating}.
 \begin{figure}
@@
 
-\emph{Research Question 2: What is the GitHub Philosophy?}\\
+\emph{Research Question 2: How is GitHub changing the OSS process?}\\
+Thung \textit{et al.} show that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enable high collaboration rates.\cite{thung2013network} In fact, the developer connectivity was shown to be higher than SourceForge and even Facebook.
+In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal for.
 

### 2013-03-10 paper/main.tex
Added correlation outputs.

@@
 \MakeRobust{\Call}
+\usepackage{amsmath}
 \setlength\parindent{0pt}
@@
 \emph{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
-A smal core of developers, however, seems to be a historial fact of OSS development\cite{mockus2000case,mockus2000two,krishnamurthy2002cave}.
+A smal core of developers, however, seems to be a historial fact of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}.
 At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers.
@@
 
-The \textsc{store} function accepts as an input a given random repository. The purpose of this function is to persist the given repository to a database for further analysis.
+The \textsc{store} function accepts as an input a given random repository.
+The purpose of this function is to persist the given repository to a database for further analysis.
+Results were stored in a MySQL relational database management system.
 
@@
 \caption{GitHub Summary Statistics}
+\label{table:summary_stats}
 \end{center}
@@
 
+Table \ref{table:summary_stats} describes summary statistics of the repository variables above.
+The most evident feature of the result is the large variance.
+In fact, for the set of tested variables $A$, each variable $A_j$ is observed to have a larger standard deviation than the variable mean.
+\begin{equation}
+\forall j \in A \colon \overline{j} < \sigma_{j}
+\label{eq:variance}
+\end{equation}
+All measured variables demonstrated this characteristic.
+
 \subsection{Hypotheses}
 \emph{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
+A correlation coefficient of \input{watcher_forks_scatterplot-correlation} was observed betwee the number of Watchers and Forks of a repository.
+This suggests that a correlation between Watchers and Forks exists.
 \begin{figure}
@@
 \end{figure}
+The data reinforces the hypothesis, as does other research in the area.
+Even though there are aguments that the number of Forks is the true measure of project success\cite{baudry2012towards}, the data suggests that the two measures are related.
+Also, \textit{Dabbish et al.} have noted that the number of Watches can be a key for others to gauge interest in a project, and thus determine if they should participate \cite{dabbish2013leveraging}.
+If the number of Watchers of a repository serves as a social cue for participation, even ``passive'' involvement with a project (Watching) could influence the ``active'' participation (Forking).
+This could possibly contribute to the correlation of the two variables.
 

### 2013-03-10 code/plotter.py

@@ -83,7 +83,9 @@ class ScatterPlot:
 		F.savefig("../paper/images/"+filename+".png")
 		plt.close(fig)
 
-		return correlation_coefficient
+		f = open("../paper/"+filename+"-correlation.tex",'w')
+		f.write(str(correlation_coefficient))
+		f.close()
 
 class Histograph:
 	def __init__(self, sql, xlabel, ylabel, range=None):

### 2013-03-11 paper/main.tex
Expanded the CTS2 part.

@@
 \subsection{Hypothesis}
-\emph{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
+\textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
 In Open Source Software, a feeling of community belonging can be intrinsicly motivating to developers\cite{lakhani2003hackers}.
@@
 
-\emph{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
+\textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
 In OSS, a project \emph{fork} has at times carried negative connotations, and has even been referred to as a hazzard\cite{kogut2001open}.
@@
 This is, ultimetely, a proof of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub.
-In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.
+In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.\\
 
-\emph{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude}\\
+\textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude}\\
 Research into the Apache Server project observed that there were far more \emph{Issue Reporters} than there were code \emph{Committers}\cite{mockus2000case}.
@@
 
-\emph{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
+\textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
 A smal core of developers, however, seems to be a historial fact of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}.
@@
 
-\emph{Research Question 2: How is GitHub changing the OSS process?}\\
+\textbf{Research Question 2: How is GitHub changing the OSS process?}\\
 The socal aspects of GitHub are an important part of the experience\cite{dabbish2012social}.
@@
 
-\emph{Research Question 3: Can GitHub be used for more that code artifacts?}\\
+\textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
 The Object Managment Group\textregistered\footnote{http://www.omg.org} computer industry consortium focused on technology standards.
@@
 Results were stored in a MySQL relational database management system.
+The database schema to store the data is decribed in figure \ref{fig:er_diagram}.
 
@@
 \end{figure}
-
-The database schema to store the data is decribed in \ref{fig:er_diagram}.
 
@@
 \subsection{Hypotheses}
-\emph{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
+\textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
 A correlation coefficient of \input{watcher_forks_scatterplot-correlation} was observed betwee the number of Watchers and Forks of a repository.
@@
 
-\emph{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
+\textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
 No significant correlation was found between the number of repository forks and issue resolution time.
@@
 
-\emph{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude.}\\
-\begin{figure}
-\includegraphics[height=3in,width=3in]{images/issue_reporters_histogram.png}
-\caption{Repostitory issue reporters}
-\end{figure}
-\begin{figure}
-\includegraphics[height=3in,width=3in]{images/committers_histogram.png}
-\caption{Repostitory committers}
-\end{figure}
+\textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude.}\\
 \begin{figure}
@@
 \caption{Repostitory issue reporters and committers}
+\label{fig:issue_reporters_committers_scatterplot}
 \end{figure}
+Data does not support an exponentially increasing number of distinct Issue Reporters as compared to distinct Committers.
+Figure\ref{fig:issue_reporters_committers_scatterplot}, does however suggest that that the two variables may be related.
+The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation cofficient of \input{issue_reporters_committers_scatterplot-correlation}.
+This does not seem to align with findings in the Apache Server project\cite{mockus2000case}.
+More investigation would be necessary to further examine this finding.
 
 \subsection{Research Questions}
-\emph{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
+\textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
 \begin{figure}
@@
 \caption{Individual committer commit percentage per repository}
+\label{fig:committers_percentage_pie_chart}
+\end{figure}
+The data in figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown.
+The inner percentages represent the percentage of committers in a project that that have committed the outer percentage amount of commits to that project, or
+
+ \[ P = \left\{ \frac{ C_u } { |C| } \times 100 \Big| u \in U \right\} \]
+
+where, given the set of commits of a repository $C$ and the set of committers to the repository $U$, the commits by each user $C_u$ are divided by the total number of commits $|C|$, yielding the set of contribution percentages $P$ for the repository.
+Elements of $P$ are then assigned to ranges $0-5\%$, $5-10\%$, $10-50\%$, $50-95\%$, and $95-100\%$ (with shared values inclusive to the larger range).
+The distribution of this range assignments is plotted on figure \ref{fig:committers_percentage_pie_chart}.
+
+There does appear to be evidence of a small set of core developers, denoted by the relatively high percentage of developers that have authored $95-100\%$ of the commits of a given repository.
+Most notably, however, is that the data shows that by a large percentage of developers committed small amounts ($0-5\%$ of the total repository commits).
+This means that a small core are not entirely monopolizing the commits, but rather opportunity exists for users to contribute, even if in small ways.
+
+\begin{figure}
+\includegraphics[height=3in,width=3in]{images/committers_histogram.png}
+\caption{Count of individual committers per repository}
+\label{fig:committers_histogram}
 \end{figure}
 
-\emph{Research Question 2: How is GitHub changing the OSS process?}\\
+When looking strictly at the number of committers to a repository, however, the notion of a small set of core Committers is strongly supported, as the vast majority of repositoris have fewer than 10 committers (figure \ref{fig:committers_histogram}).
+
+\textbf{Research Question 2: How is GitHub changing the OSS process?}\\
 Thung \textit{et al.} show that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enable high collaboration rates.\cite{thung2013network} In fact, the developer connectivity was shown to be higher than SourceForge and even Facebook.
@@
 
-\emph{Research Question 3: Can GitHub be used for more that code artifacts?}\\
-As part of the OMG\textregistered standardization process, the CTS2 Specification used GitHub issue tracker to track specification changes.
+\textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
+The OMG\textregistered standardization process is a process by which products strive to produce an interoperability statndard by way of an industry standards consortium.\cite{omgTechProcess}
+The standardization process itself aims to, at its conclusion, produce a set of model artificats containing sufficient detail such that when interpreted and implemented, produce interoperable software.
+At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process.
+If so,  of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,\cite{sommerville1998requirements} we can show that GitHub is an apporiate forum for many of them.
+
+As a case study, the CTS2 OMG\textregistered  Specification\footnote{https://github.com/cts2/cts2-specification} used GitHub issue tracker to track specification changes.
+Elicitation, Analysis and Negotiation were natural fits for the issue tracker.
+This was chosen in part because the barrier to participation was very small (only a GitHub account was needed), and issues tracker itself allowed for dialogs in the form of comments.
+Because customer involvement is critical to these activities\cite{paetsch2003requirements}, ease of contribution was a main concern.
+OMG has a predescribed format for requirements Documentation, but using the GitHub issue API\footnote{http://developer.github.com/v3/issues/}, much of this documentation can be automated.
+Validation and Management are areas of requirements engineering that GitHub has interesting solutions for.
+First, in order to track (or validate) changes to the specification, when the actual specification is modified, the commit that modified it can be traced to a named issue by simply putting the issue number in the commit log.
+This level of traceability is powerful, as it allows every change to the specification artifacts to be tracked consistently to an open issue.
+As the normal OMG process proceeds, requirements can be managed through ``tagging'' the issues.
+By allowing issues to be grouped by arbitrary tags, users are more free to adapt their own workflow.
+For instance, when community voting on issues needed to be done for CTS2, a simple ``ready for vote'' tag signified to the voting board exactly what was under consideration.
+
+As mentioned, the CTS2 specification made heavy use of the GitHub issue tracker, as well as another OMG specification, ServD\footnote{https://github.com/servd/servd-specification}.
+For open source specifications, or requirements engineering activities, GitHub has shown to be a viable platform.
+
+\section{Conclusion}
 

### 2013-03-18 paper/main.tex
Removed the date.
Added a CTS2 citation.

@@
 }
-
-\date{}
 
@@
 
-As a case study, the CTS2 OMG\textregistered  Specification\footnote{https://github.com/cts2/cts2-specification} used GitHub issue tracker to track specification changes.
+As a case study, the CTS2 OMG\textregistered Specification\cite{cts2} used a GitHub issue tracker\footnote{https://github.com/cts2/cts2-specification/issues} to track specification changes.
 Elicitation, Analysis and Negotiation were natural fits for the issue tracker.

### 2013-03-19 paper/main.tex
Still working...

@@
 
-\subsection{Hypothesis}
+\subsection{Hypotheses}
+The following hypotheses
+
 \textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
@@
 The individual attrubites of $A$, or $A_j$, are described below:\\
-\textit{Issues}\\
-The number of issues posted to a repository's issue tracker\footnote{https://github.com/blog/411-github-issue-tracker}.
+\textit{Issues:}
+The number of issues posted to a repository's issue tracker\footnote{https://github.com/blog/411-github-issue-tracker}.\\
 
-\textit{Commits}\\
+\textit{Commits:}
 The number of individual source code commits to a repository.
+It is important to note that a textit{commit} is a single transaction that changes the code contents of a repository.
+Actual commit ``size'' (lines of code modified, for examlple), is not taken into account.
 
-\textit{Closed Issues}\\
+\textit{Closed Issues:}
 The number of issues in a repository in a state of \textsc{closed} at the time of processing.
 
-\textit{Open Issues}\\
+\textit{Open Issues:}
 The number of issues in a repository in a state of \textsc{open} at the time of processing.
 
-\textit{Issue Close Time}\\
+\textit{Issue Close Time:}
 The average time taken to close a given issue in a repository. Non-\textsc{closed} issues were not couted.
@@
 
-\textit{Issue Creators}\\
+\textit{Issue Creators:}
+The number of distinct users that have opened at least one \textit{Issue}.
 
-\textit{Committers}\\
+\textit{Committers:}
+The number of distinct users that have summitted at least one \textit{Commit}.
 
@@
 \end{tabular}
-N = 100 Repositories
+N = 1000 Repositories
 \caption{GitHub Summary Statistics}
@@
 
+The sample ($N = 1000$) GitHub repositories were analyzed, out of a total of 5.7 million\cite{githubPress}.
+
 \subsection{Hypotheses}
@@
 If the number of Watchers of a repository serves as a social cue for participation, even ``passive'' involvement with a project (Watching) could influence the ``active'' participation (Forking).
-This could possibly contribute to the correlation of the two variables.
+This could possibly contribute to the correlation of the two variables.\\
 
 \textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
+\begin{figure}
+\includegraphics[height=3in,width=3in]{images/issue_close_time_forks_scatterplot.png}
+\caption{Repostitory issue close time and number of forks}
+\label{fig:issue_close_time_forks_scatterplot}
+\end{figure}
 No significant correlation was found between the number of repository forks and issue resolution time.
@@
 \textit{second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all.
-\textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub -- something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric \cite{phamcreating}.
-\begin{figure}
-\includegraphics[height=3in,width=3in]{images/issue_close_time_forks_scatterplot.png}
-\caption{Repostitory issue close time and number of forks}
-\end{figure}
+\textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub -- something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric \cite{phamcreating}.\\
+
+It is possible that other factors, such as issue prioritization, may have a greater impact on issue resolution time.
+Research does support a relationship\cite{mockus2002two} between the stated priority and fix interval.
+As the GitHub issue tracker can only support prioritization by free-text \textit{tags} placed on issues, it is not feasable to programmattically gather issue priority metrics.
 
@@
 Data does not support an exponentially increasing number of distinct Issue Reporters as compared to distinct Committers. 
-Figure\ref{fig:issue_reporters_committers_scatterplot}, does however suggest that that the two variables may be related.
+As shown in Figure\ref{fig:issue_reporters_committers_scatterplot}, the slope of the relationship between the two variables is less than one.
+This does not seem to completely align with findings in the Apache Server project\cite{mockus2000case}.
+
+Figure\ref{fig:issue_reporters_committers_scatterplot}, does however suggest, however, that that the two variables may be related.
 The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation cofficient of \input{issue_reporters_committers_scatterplot-correlation}.
-This does not seem to align with findings in the Apache Server project\cite{mockus2000case}.
-More investigation would be necessary to further examine this finding.
+More investigation would be necessary to further examine this finding.\\
 
@@
 Elements of $P$ are then assigned to ranges $0-5\%$, $5-10\%$, $10-50\%$, $50-95\%$, and $95-100\%$ (with shared values inclusive to the larger range).
-The distribution of this range assignments is plotted on figure \ref{fig:committers_percentage_pie_chart}.
+The distribution of this range assignments is plotted on Figure \ref{fig:committers_percentage_pie_chart}.
 
@@
 
-When looking strictly at the number of committers to a repository, however, the notion of a small set of core Committers is strongly supported, as the vast majority of repositoris have fewer than 10 committers (figure \ref{fig:committers_histogram}).
+When looking strictly at the number of committers to a repository, however, the notion of a small set of core Committers is strongly supported, as the vast majority of repositoris have fewer than 10 committers (figure \ref{fig:committers_histogram}).\\
 
@@
 Thung \textit{et al.} show that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enable high collaboration rates.\cite{thung2013network} In fact, the developer connectivity was shown to be higher than SourceForge and even Facebook.
-In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal for.
+In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal for.\\
+
+Ye also observed\cite{ye2003toward} that the contribution environment, or ``society,'' is important in providing a framework for develper trust building:
+\begin{quote}
+Only in a society where technical supremacy is highly appreciated can developers acquire good reputations among their peers by displaying their skills through free distribution, and often wider acceptance, of their systems.The good reputation attracts attention, trust, and cooperation from others and lays the foundation for advancing the original developers agenda and the establishment and development of OSS communities.
+\end{quote}
+
+To further analyze this research question, and to explore further the GitHub ``society'' and its impact on OSS development, an interview with GitHub staff was conducted.\\ 
+\textit{Interview Q1: How is GitHub changing the way OSS is being developed?}\\
+
+\textit{Interview Q2: Does GitHub aim to change the way OSS is developed, or simply facilitate?}\\
+
+\textit{Interview Q3: How is GitHub changing the way OSS is being developed?}\\
 
@@
 At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process.
-If so,  of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,\cite{sommerville1998requirements} we can show that GitHub is an apporiate forum for many of them.
+If so,  of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,\cite{sommerville1998requirements} we can show that GitHub is an apporiate forum for many of them.\\
 

### 2013-03-19 paper/main.tex
Added more method descriptions.

@@
 \usepackage{algpseudocode}
+\usepackage{url}
 \MakeRobust{\Call}
@@
 By analyzing a subset of GitHub repositories, this report will show how GitHub has influenced some very intrinsic aspects of traditional OSS develoment, such as developer hierarchies and issue close velicity.
-We find that...
-TODO
+We find that many of the traditional aspects of OSS development remain, such most project developlment is done by a small group of core developers.
+Others assumptions about project hierarchy, such as a large number of Issue Reporters compared to Committers, seems unsupported by the data.
+
 \end{abstract}
 
-\noindent \textbf{Keywords.} Git, GitHub, OpenSource
+\noindent \\\textbf{Keywords.} Git, GitHub, Open Source
 
@@
 \subsection{Collection and Storage}
-Data collection was done using the GitHub REST API\footnote{http://developer.github.com/v3/}.
-A random selection of 1000 repositories was selected using the following algorithm:\footnote{https://github.com/kevinpeterson/github-process-research/blob/master/code/research.py}
+Data collection was done using the GitHub REST API\footnote{http://developer.github.com/v3/}. A random selection of 1000 repositories was selected using the following algorithm:
 
@@
 Results were stored in a MySQL relational database management system.
-The database schema to store the data is decribed in figure \ref{fig:er_diagram}.
+The database schema to store the data is decribed in Figure \ref{fig:er_diagram}.
 
@@
 
+Data was collected and analyzed using a Python\footnote{http://python.org/} script.
+Because the GitHub REST API constrains the number of API calls per hour made by a client, logic was included to pause execution when the allotment was exhausted.
+
+Matplotlib\cite{Hunter2007}, a 2D graphics package for Python, was used to render graphs and charts.
+NumPY and SciPY\cite{scipy} were used for data analysis calculations, such as correllation coefficients.
+
+All source code and materials used in data collection and analysis are hosted at GitHub (naturally!), and may be accessed by the following URL: \url{https://github.com/kevinpeterson/github-process-research}
+
 \subsection{Analysis}
-
 Where the Set $S$ is the random GitHub repository sample.
@@
 This suggests that a correlation between Watchers and Forks exists.
+
 \begin{figure}
@@
 \end{figure}
+
 The data reinforces the hypothesis, as does other research in the area.
@@
 \textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
+
 \begin{figure}
@@
 \end{figure}
+
 No significant correlation was found between the number of repository forks and issue resolution time.
@@
 \textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude.}\\
+Data does not support an exponentially increasing number of distinct Issue Reporters as compared to distinct Committers. 
+As shown in Figure\ref{fig:issue_reporters_committers_scatterplot}, the slope of the relationship between the two variables is less than one.
+This does not seem to completely align with findings in the Apache Server project\cite{mockus2000case}.
+
 \begin{figure}
@@
 \end{figure}
-Data does not support an exponentially increasing number of distinct Issue Reporters as compared to distinct Committers. 
-As shown in Figure\ref{fig:issue_reporters_committers_scatterplot}, the slope of the relationship between the two variables is less than one.
-This does not seem to completely align with findings in the Apache Server project\cite{mockus2000case}.
 
@@
 \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
+The data in figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown.
+The inner percentages represent the percentage of committers in a project that that have committed the outer percentage amount of commits to that project, or
+
+ \[ P = \left\{ \frac{ C_u } { |C| } \times 100 \Big| u \in U \right\} \]
+
 \begin{figure}
@@
 \end{figure}
-The data in figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown.
-The inner percentages represent the percentage of committers in a project that that have committed the outer percentage amount of commits to that project, or
-
- \[ P = \left\{ \frac{ C_u } { |C| } \times 100 \Big| u \in U \right\} \]
 

### 2013-03-19 paper/main.tex
Added README.

@@
 
-\textit{Interview Q2: Does GitHub aim to change the way OSS is developed, or simply facilitate?}\\
+\textit{Interview Q2: Many new Software Developers are will be starting their education/training already having experience with GitHub.
+How will Software Engineering as a discipline (and how it is taught) change to accodate the ideas popularized by GitHub?}\\
 
-\textit{Interview Q3: How is GitHub changing the way OSS is being developed?}\\
+\textit{Interview Q3: There are many non-code artifacts being hosted on GitHub, for example, Object Management Group Specifications, Laws and Statutes, books...
+etc.
+Do you see GitHub participation expanding in the non-code areas?}\\
 

### 2013-03-19 paper/main.tex
Fixed some typos.

@@
 The development process by which these projects are produced is generally unstructured compared to commercial software -- but a definite pattern does arrive, and is no less a pattern.
-GitHub, a popular OSS code hosting website, and Git, the site's SCM of choice, may have the potential to fundamentally change ths process.\\
+GitHub, a popular OSS code hosting website, and Git, the site's Source Code Management (SCM) tool of choice, may have the potential to fundamentally change ths process.\\
 By analyzing a subset of GitHub repositories, this report will show how GitHub has influenced some very intrinsic aspects of traditional OSS develoment, such as developer hierarchies and issue close velicity.
@@
 Because issue reporting is of low risk to the code base, but has potentially high value, it is a perfect way for large numbers of people to contribute.
-It is expected that the findings of the Apache Server project research will hold true for GitHub projects as well.\\
+It is expected that the findings of the Apache Server project research will hold true for GitHub projects as well.
 

### 2013-03-19 paper/main.tex
Ran spellcheck.

@@
 GitHub, a popular OSS code hosting website, and Git, the site's Source Code Management (SCM) tool of choice, may have the potential to fundamentally change ths process.\\
-By analyzing a subset of GitHub repositories, this report will show how GitHub has influenced some very intrinsic aspects of traditional OSS develoment, such as developer hierarchies and issue close velicity.
-We find that many of the traditional aspects of OSS development remain, such most project developlment is done by a small group of core developers.
+
+By analyzing a subset of GitHub repositories, this report will show how GitHub has influenced some very intrinsic aspects of traditional OSS development, such as developer hierarchies and issue close velocity.
+We find that many of the traditional aspects of OSS development remain, such most project development is done by a small group of core developers.
 Others assumptions about project hierarchy, such as a large number of Issue Reporters compared to Committers, seems unsupported by the data.
@@
 Although the spirit and intent may be the same, the tool set is drastially different.
-The intent of this paper is to explore how GitHub facilitates this process, and also how it may cause it to evolve.
+The intent of this work is to explore how GitHub facilitates this process, and also how it may cause it to evolve.\\
+
+To better understand quantitatively the GitHub process, three hypotheses are introduced.
+The goal of these hypotheses is to bridge the gap between the more qualitative research of GitHub\cite{dabbish2012social,begel2013social} and the statistical analysis of data from a random sampling of GitHub projects.\\
+
+In addition, three supporting research questions are proposed.
+These questions will draw upon data from this work, but also reference interviews with GitHub staff to better represent how GitHub may be changing how OSS development is conducted.
 
@@
 \textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
-In Open Source Software, a feeling of community belonging can be intrinsicly motivating to developers\cite{lakhani2003hackers}.
+In Open Source Software, a feeling of community belonging can be intrinsically motivating to developers\cite{lakhani2003hackers}.
 This feeling of belonging can be expressed passively, as a \emph{watch}, or actively, as a \emph{fork} of a repository.
 As \emph{watches} don't necessarily signal intent to contribute, a positive direction for an OSS project is to not only increase the number of \emph{watchers}, but also the number of \emph{forks}.
-As both signify a general general interest in the project, it is expected that there be some correllation between the two.\\
+As both signify a general general interest in the project, it is expected that there be some correlation between the two.\\
 
@@
 In OSS, a project \emph{fork} has at times carried negative connotations, and has even been referred to as a hazzard\cite{kogut2001open}.
-In the context of Git and GitHub, however, a \emph{fork} is a positive occurence for a project, as it signifies greater project involvement.
+In the context of Git and GitHub, however, a \emph{fork} is a positive occurrence for a project, as it signifies greater project involvement.
 As Git allows for easy merging of \emph{forks}, code contributions in the form of defect fixes can be incorporated quickly.
@@
 
-This is, ultimetely, a proof of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub.
+This is, ultimately, a proof of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub.
 In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.\\
@@
 \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
-A smal core of developers, however, seems to be a historial fact of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}.
+A smal core of developers, however, seems to be a historical fact of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}.
 At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers.
@@
 \textbf{Research Question 2: How is GitHub changing the OSS process?}\\
-The socal aspects of GitHub are an important part of the experience\cite{dabbish2012social}.
+The social aspects of GitHub are an important part of the experience\cite{dabbish2012social}.
 User interactions and social pressures can drive OSS development in interesting ways.
@@
 \textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
-The Object Managment Group\textregistered\footnote{http://www.omg.org} computer industry consortium focused on technology standards.
-Potential standard summission teams must go through a vetting process\cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard.
+The Object Management Group\textregistered\footnote{http://www.omg.org} computer industry consortium focused on technology standards.
+Potential standard submission teams must go through a vetting process\cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard.
 Gathering and recording this feedback is a challenge, but GitHub's issue tracking system may be able to streamline the process.
@@
 Results were stored in a MySQL relational database management system.
-The database schema to store the data is decribed in Figure \ref{fig:er_diagram}.
+The database schema to store the data is described in Figure \ref{fig:er_diagram}.
 
@@
 Matplotlib\cite{Hunter2007}, a 2D graphics package for Python, was used to render graphs and charts.
-NumPY and SciPY\cite{scipy} were used for data analysis calculations, such as correllation coefficients.
+NumPY and SciPY\cite{scipy} were used for data analysis calculations, such as correlation coefficients.
 
@@
 
-Results are then averaged over all repositories, yeilding:
+Results are then averaged over all repositories, yielding:
 \[ \frac{\sum\limits_{i=1}^{n} f(  A_{S{_i}j}  ) } {n} \]
 
-The individual attrubites of $A$, or $A_j$, are described below:\\
+The individual attributes of $A$, or $A_j$, are described below:\\
 \textit{Issues:}
@@
 It is important to note that a textit{commit} is a single transaction that changes the code contents of a repository.
-Actual commit ``size'' (lines of code modified, for examlple), is not taken into account.
+Actual commit ``size'' (lines of code modified, for example), is not taken into account.
 
@@
 \textit{Issue Close Time:}
-The average time taken to close a given issue in a repository. Non-\textsc{closed} issues were not couted.
+The average time taken to close a given issue in a repository. Non-\textsc{closed} issues were not counted.
 
@@
 \textit{Committers:}
-The number of distinct users that have summitted at least one \textit{Commit}.
+The number of distinct users that have summited at least one \textit{Commit}.
 
@@
 \textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
-A correlation coefficient of \input{watcher_forks_scatterplot-correlation} was observed betwee the number of Watchers and Forks of a repository.
+A correlation coefficient of \input{watcher_forks_scatterplot-correlation} was observed between the number of Watchers and Forks of a repository.
 This suggests that a correlation between Watchers and Forks exists.
@@
 The data reinforces the hypothesis, as does other research in the area.
-Even though there are aguments that the number of Forks is the true measure of project success\cite{baudry2012towards}, the data suggests that the two measures are related.
+Even though there are arguments that the number of Forks is the true measure of project success\cite{baudry2012towards}, the data suggests that the two measures are related.
 Also, \textit{Dabbish et al.} have noted that the number of Watches can be a key for others to gauge interest in a project, and thus determine if they should participate \cite{dabbish2013leveraging}.
@@
 Research does support a relationship\cite{mockus2002two} between the stated priority and fix interval.
-As the GitHub issue tracker can only support prioritization by free-text \textit{tags} placed on issues, it is not feasable to programmattically gather issue priority metrics.
+As the GitHub issue tracker can only support prioritization by free-text \textit{tags} placed on issues, it is not feasable to programmattically gather issue priority metrics.\\
 
@@
 
-When looking strictly at the number of committers to a repository, however, the notion of a small set of core Committers is strongly supported, as the vast majority of repositoris have fewer than 10 committers (figure \ref{fig:committers_histogram}).\\
+When looking strictly at the number of committers to a repository, however, the notion of a small set of core Committers is strongly supported, as the vast majority of repositories have fewer than 10 committers (figure \ref{fig:committers_histogram}).\\
 
@@
 At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process.
-If so,  of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,\cite{sommerville1998requirements} we can show that GitHub is an apporiate forum for many of them.\\
+If so,  of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,\cite{sommerville1998requirements} we can show that GitHub is an appropriate forum for many of them.\\
 

### 2013-03-20 paper/main.tex
Made progress...

@@
-% SEng 5852 - Spring 2013 White Paper Review
 \documentclass{proc}
@@
 \usepackage{fixltx2e}
-\usepackage{algorithmicx}
+\usepackage{algorithm}
 \usepackage{algpseudocode}
+\usepackage{caption}
 \usepackage{url}
-\MakeRobust{\Call}
 \usepackage{amsmath}
-\setlength\parindent{0pt}
+%\setlength{\parskip}{\baselineskip}
+%\setlength{\parindent}{0pt}
 
@@
 The development process by which these projects are produced is generally unstructured compared to commercial software -- but a definite pattern does arrive, and is no less a pattern.
-GitHub, a popular OSS code hosting website, and Git, the site's Source Code Management (SCM) tool of choice, may have the potential to fundamentally change ths process.\\
+GitHub, a popular OSS code hosting website, and Git, the site's Source Code Management (SCM) tool of choice, may have the potential to fundamentally change ths process.
 
@@
 Others assumptions about project hierarchy, such as a large number of Issue Reporters compared to Committers, seems unsupported by the data.
-
 \end{abstract}
@@
 Although the spirit and intent may be the same, the tool set is drastially different.
-The intent of this work is to explore how GitHub facilitates this process, and also how it may cause it to evolve.\\
+The intent of this work is to explore how GitHub facilitates this process, and also how it may cause it to evolve.
+
+GitHub\footnote{https://github.com/} is a popular code hosting website that uses Git Source Code Management (SCM)\footnote{http://git-scm.com/}.
+GitHub...
+Git allows for development to take place in a more distributed manner than previously available in other SCM systems\cite{spinellis2012git}.
+Git also allows for a variety of work flows\cite{chacon2009pro}.
+These workflows may be tailored to the individual needs of a project.
+The Linux Kernel, for example, uses a Dictator and
+Lieutenants Workflow\cite{platschekfloss}, which is hierarchical in nature.
+Other work flows may be more distributed, or resemble more traditional centralized SCM systems.
+Although work flow techniques aren't explicitly addressed in this work, it is important to note that work flow change may contribute to \textit{process change}.
 
 To better understand quantitatively the GitHub process, three hypotheses are introduced.
-The goal of these hypotheses is to bridge the gap between the more qualitative research of GitHub\cite{dabbish2012social,begel2013social} and the statistical analysis of data from a random sampling of GitHub projects.\\
+The goal of these hypotheses is to bridge the gap between the more qualitative research of GitHub\cite{dabbish2012social,begel2013social} and the statistical analysis of data from a random sampling of GitHub projects.
 
@@
 
-\textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
+\noindent \textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
 In Open Source Software, a feeling of community belonging can be intrinsically motivating to developers\cite{lakhani2003hackers}.
@@
 As \emph{watches} don't necessarily signal intent to contribute, a positive direction for an OSS project is to not only increase the number of \emph{watchers}, but also the number of \emph{forks}.
-As both signify a general general interest in the project, it is expected that there be some correlation between the two.\\
+As both signify a general general interest in the project, it is expected that there be some correlation between the two.
 
-\textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
+\noindent \textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
 In OSS, a project \emph{fork} has at times carried negative connotations, and has even been referred to as a hazzard\cite{kogut2001open}.
@@
 As Git allows for easy merging of \emph{forks}, code contributions in the form of defect fixes can be incorporated quickly.
-Because of this, it is expected that \emph{Issue Resolution Time} will decrease as the number of \emph{forks} increases.\\
+Because of this, it is expected that \emph{Issue Resolution Time} will decrease as the number of \emph{forks} increases.
 
@@
 ``Patches'' may be released at any time, by any member of the community\cite{mockus2000case}.
-``Patches"" in the context of GitHub could be equated to Pull Requests.\\
+``Patches'' in the context of GitHub could be equated to Pull Requests.
 
 This is, ultimately, a proof of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub.
-In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.\\
+In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.
 
-\textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude}\\
+\noindent \textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude}\\
 Research into the Apache Server project observed that there were far more \emph{Issue Reporters} than there were code \emph{Committers}\cite{mockus2000case}.
@@
 The 10-15 {\it Core Developers} contribute around 80\% of the new functionality, while the rest of the 400 code contributors focus primarily on bug fixes.
-The {\it Defect Reporters} were by far the largest group, with over 3000 individuals submitting bug reports.\\
+The {\it Defect Reporters} were by far the largest group, with over 3000 individuals submitting bug reports.
 
@@
 
-\textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
+\noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
 A smal core of developers, however, seems to be a historical fact of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}.
 At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers.
-Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?\\
+Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?
 
-\textbf{Research Question 2: How is GitHub changing the OSS process?}\\
+\noindent \textbf{Research Question 2: How is GitHub changing the OSS process?}\\
 The social aspects of GitHub are an important part of the experience\cite{dabbish2012social}.
@@
 Social aspects of OSS development have existed before GitHub, with mailing lists\cite{mockus2000case}, gift giving mechanisms\cite{bergquist2008power}, and other code hosting sites.
-What is GitHub doing to facilitate change on OSS development, and how does it see OSS development changing moving forward?\\
+What is GitHub doing to facilitate change on OSS development, and how does it see OSS development changing moving forward?
 
-\textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
+\noindent \textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
 The Object Management Group\textregistered\footnote{http://www.omg.org} computer industry consortium focused on technology standards.
@@
 \subsection{Collection and Storage}
-Data collection was done using the GitHub REST API\footnote{http://developer.github.com/v3/}. A random selection of 1000 repositories was selected using the following algorithm:
+Data collection was done using the GitHub REST API\footnote{http://developer.github.com/v3/}. A random selection of 1000 repositories was selected using Algorithm \ref{algo:data_collection_algo}.
 
-\begin{algorithmic}
+\begin{algorithm}[H]
+\captionof{algorithm}{Data Collection Algorithm}\label{algo:data_collection_algo}
+\begin{algorithmic}[1]
 \While{$i < 1000$} 
@@
 \State $repos \gets git\_search(word)$
-\State $j \gets \Call{random}{0... \Call{size}{repos}}$
+\State $j \gets $%\Call{random}{0... \Call{size}{repos}}$
 \State $repo \gets repos_j$
-\State \Call{store}{repo}
+\State %\Call{store}{repo}
 \EndWhile
 \end{algorithmic}
+\end{algorithm}
 
@@
 Where the Set $S$ is the random GitHub repository sample.
-Each randomly collected repository $S_i$ contains a Set of attriubtes $A$.
+Each randomly collected repository $S_i$ contains a Set of attributes $A$.
 Thus, a set of attributes for a given repository can be referred to as $A_{S_i}$.
@@
 For this analysis, the aggregate functions are: \textsc{mean}, \textsc{min}, \textsc{max}, \textsc{stddev}.
+This analysis is applied to all $n$ repositories for which data was collected.
+Results were then averaged over all repositories.
+Each attribute is analyzed independently, so given:
+\[ g(j) = \frac{\sum\limits_{i=1}^{n} f(  A_{S{_i}j}  ) } {n} \]
+each individual attribute is processed via $g(j)$ and will result in the average of the given aggregate function.
 
-Results are then averaged over all repositories, yielding:
-\[ \frac{\sum\limits_{i=1}^{n} f(  A_{S{_i}j}  ) } {n} \]
+The individual attributes of $A$, or $A_j$, are described below. For detailed summary statistics of these attributes, see Figure  \ref{table:summary_stats}.\\\\
+\noindent \textbf{\textit{Issues:}}
+The number of issues posted to a repository's issue tracker\footnote{https://github.com/blog/411-github-issue-tracker}.
 
-The individual attributes of $A$, or $A_j$, are described below:\\
-\textit{Issues:}
-The number of issues posted to a repository's issue tracker\footnote{https://github.com/blog/411-github-issue-tracker}.\\
-
-\textit{Commits:}
+\noindent \textbf{\textit{Commits:}}
 The number of individual source code commits to a repository.
@@
 
-\textit{Closed Issues:}
+\noindent \textbf{\textit{Closed Issues:}}
 The number of issues in a repository in a state of \textsc{closed} at the time of processing.
 
-\textit{Open Issues:}
+\noindent \textbf{\textit{Open Issues:}}
 The number of issues in a repository in a state of \textsc{open} at the time of processing.
 
-\textit{Issue Close Time:}
-The average time taken to close a given issue in a repository. Non-\textsc{closed} issues were not counted.
+\noindent \textbf{\textit{Issue Close Time:}}
+The average time taken to close a given issue in a repository. Non-\textsc{closed} issues were not counted.\\
 
-The average is calculated using the following method. Let $d(x)$ be a function that converts seconds to a Date, and let $s(x)$ be a function that converts a Date to seconds.
+The average is calculated using the following method.
+Let $d(x)$ be a function that converts seconds to a Date, and let $s(x)$ be a function that converts a Date to seconds.
+The variable $n$ denotes the total number of issues, or $I$, assigned to a particular repository.
 
-\[ d\left( \frac{\sum\limits_{i=1}^{n} s(i_{\textrm{closed}}) - s(i_{\textrm{opened}})  } {n} \right) \]
+\[ \tau = d\left( \frac{\sum_{i=1}^{n} s(C_{I_i} ) - s(O_{I_i} )  } {n} \right) \]
 
-\textit{Issue Creators:}
+For each issue $I_i$ in $I$, the \textsc{opened\_date} $O_{I_i}$ is subtracted from the \textsc{closed\_date} $C_{I_i}$.
+The result of this is averaged and converted into a date via $d(x)$, resulting in the average issue close time interval $\tau$.
+
+\noindent \textbf{\textit{Issue Creators:}}
 The number of distinct users that have opened at least one \textit{Issue}.
 
-\textit{Committers:}
+\noindent \textbf{\textit{Committers:}}
 The number of distinct users that have summited at least one \textit{Commit}.
@@
 \textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
-
-\begin{figure}
-\includegraphics[height=3in,width=3in]{images/issue_close_time_forks_scatterplot.png}
-\caption{Repostitory issue close time and number of forks}
-\label{fig:issue_close_time_forks_scatterplot}
-\end{figure}
-
 No significant correlation was found between the number of repository forks and issue resolution time.
@@
 \textit{second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all.
-\textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub -- something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric \cite{phamcreating}.\\
+\textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub -- something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric \cite{phamcreating}.
+
+\begin{figure}
+\includegraphics[height=3in,width=3in]{images/issue_close_time_forks_scatterplot.png}
+\caption{Repostitory issue close time and number of forks}
+\label{fig:issue_close_time_forks_scatterplot}
+\end{figure}
 
@@
 Thung \textit{et al.} show that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enable high collaboration rates.\cite{thung2013network} In fact, the developer connectivity was shown to be higher than SourceForge and even Facebook.
-In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal for.\\
+In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal for.
 
@@
 
-To further analyze this research question, and to explore further the GitHub ``society'' and its impact on OSS development, an interview with GitHub staff was conducted.\\ 
+To further analyze this research question, and to explore further the GitHub ``society'' and its impact on OSS development, an interview with GitHub staff was conducted.\\\\
 \textit{Interview Q1: How is GitHub changing the way OSS is being developed?}\\
@@
 At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process.
-If so,  of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,\cite{sommerville1998requirements} we can show that GitHub is an appropriate forum for many of them.\\
+If so,  of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,\cite{sommerville1998requirements} we can show that GitHub is an appropriate forum for many of them.
 
@@
 By allowing issues to be grouped by arbitrary tags, users are more free to adapt their own workflow.
-For instance, when community voting on issues needed to be done for CTS2, a simple ``ready for vote'' tag signified to the voting board exactly what was under consideration.
+For instance, when community voting on issues needed to be done for CTS2, a simple ``ready for vote'' tag signified to the voting board exactly what was under consideration
 

### 2013-03-22 paper/main.tex
Upadated some equations.

@@
 \State $repo \gets repos_j$
-\State %\Call{store}{repo}
+\State \Call{store}{repo}
 \EndWhile
@@
 \subsection{Analysis}
-Where the Set $S$ is the random GitHub repository sample.
-Each randomly collected repository $S_i$ contains a Set of attributes $A$.
-Thus, a set of attributes for a given repository can be referred to as $A_{S_i}$.
+In this context, a \textit{Repository} is defined as a Set of \textit{Attributes}.
+An \textit{Attribute} is a metric of interest.
+See Section \ref{sec:attributes} for further information on individual \textit{Attributes}.
+These metrics of interest are used as an Indexing Set $A^\prime$, and have labels as follows:
 
-Given the total set of attributes $A$, each individual attribute set can be denoted as $A_j$.
-Each individual attribute processed by some aggregate function $f$, or $f(  A_j )$.
+\[A^\prime = \{i, c, i^\prime, i^{\prime\prime}, \Delta, \alpha, \beta\}\]
+
+Where $i$ = {\textit{Issues}, $c$ = {\textit{Commits}, $i^\prime$ = {\textit{Closed Issues}, $i^{\prime\prime}$ = {\textit{Open Issues}, $\Delta$ = {\textit{Issue Close Time}, $\alpha$ = {\textit{Issue Creators}, and $\beta$ = {\textit{Committers}.
+
+All analyzed \textit{Respositories} can be thought of as the Indexed Family of \textit{Attribute} Sets $A$, or a Set of Sets of \textit{Attributes}.
+Let the Set of these Sets of \textit{Attribute} Sets be called $S$, or the \textit{Repository Sample}.
+We can label the \textit{Attribute} Sets by arbitrary integers 0 through $|S|$, or:
+\[ \{A_i\}_{i \in \{0,1,2 ... |S|\} }\]
+
+In this way, we can reference a Set of \textit{Attributes} (or, an individual \textit{Repository}), by $A_i$.
+
+Going further, we can also denote an individual \textit{Attribute} of a given \textit{Repository} by:
+\[ \{A_{(i,r)}\}_{(i,r) \in A^\prime \times \{0,1,2 ... |S|\} } \]
+or, $A_{i,r}$. For example, $A_{\Delta,10}$ would denote the value of the \textit{Issue Close Time} ($\Delta$) of Repository \textit{10}.
+
+Each individual Attribute processed by some aggregate function $f$, or $f(  A_{i,r} )$.
 For this analysis, the aggregate functions are: \textsc{mean}, \textsc{min}, \textsc{max}, \textsc{stddev}.
@@
 Each attribute is analyzed independently, so given:
-\[ g(j) = \frac{\sum\limits_{i=1}^{n} f(  A_{S{_i}j}  ) } {n} \]
+\[ g(j) = \frac{\sum\limits_{r = 0}^{|S|} f(  A_{r,j}  ) }{|S|} \]
 each individual attribute is processed via $g(j)$ and will result in the average of the given aggregate function.
 
-The individual attributes of $A$, or $A_j$, are described below. For detailed summary statistics of these attributes, see Figure  \ref{table:summary_stats}.\\\\
+\subsubsection{Attributes}
+\label{sec:attributes}
+The individual \textit{Attribute} names, or the elements of $A^\prime$, are described below. For detailed summary statistics of these \textit{Attributes}, see Figure  \ref{table:summary_stats}.\\\\
 \noindent \textbf{\textit{Issues:}}
@@
 Let $d(x)$ be a function that converts seconds to a Date, and let $s(x)$ be a function that converts a Date to seconds.
-The variable $n$ denotes the total number of issues, or $I$, assigned to a particular repository.
+Let $I$ equal the Set of \textit{Issues} in a given repository, with elements comprised of tuples $(\tau, \tau^\prime)$, where $\tau$ is \textit{Issue Open Time} and $\tau^\prime$ is \textit{Issue Close Time}.
+We calculate \textit{Issue Close Time} by:
 
-\[ \tau = d\left( \frac{\sum_{i=1}^{n} s(C_{I_i} ) - s(O_{I_i} )  } {n} \right) \]
+\[ \Delta = d\left( \frac{\sum_{i \in I} s( p_{(1)}i ) - s( p_{(2)}i )  } {n} \right) \]
 
-For each issue $I_i$ in $I$, the \textsc{opened\_date} $O_{I_i}$ is subtracted from the \textsc{closed\_date} $C_{I_i}$.
-The result of this is averaged and converted into a date via $d(x)$, resulting in the average issue close time interval $\tau$.
+Where for each issue $i$ in $I$, the \textit{Issue Open Date} $s( p_{(1)}i)$ is subtracted from the \textit{Issue Close Date} $s( p_{(2)}i)$, and $p_{(1)}i$ and $p_{(2)}i$ denote the elements of the tuple, or \textit{Issue Open Date} and \textit{Issue Close Date}, respectively.
+The result of this is averaged and converted into a Date via $d(x)$, resulting in the average issue close time interval $\Delta$.
 
@@
 The data in figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown.
-The inner percentages represent the percentage of committers in a project that that have committed the outer percentage amount of commits to that project, or
+The inner percentages represent the percentage of committers in a project that that have committed the outer percentage amount of commits to that project, or:
 
- \[ P = \left\{ \frac{ C_u } { |C| } \times 100 \Big| u \in U \right\} \]
+\begin{equation}
+P = \left\{ \frac{ C_u } { |C| } \times 100 \Big| u \in U \right\}
+\label{eq:commit_percentage}
+\end{equation}
+
+Where $U$ is the Set of \textit{Committers} to a given \textit{Repository} and $C$ is the MultiSet of \textit{Commits} by an individual \textit{Committers} to a given \textit{Repository}, or:
+ \[ \{ C_u \}_{u \in U} \]
 
@@
 
-where, given the set of commits of a repository $C$ and the set of committers to the repository $U$, the commits by each user $C_u$ are divided by the total number of commits $|C|$, yielding the set of contribution percentages $P$ for the repository.
+To find the percentages Set $P$ of equation\ref{eq:commit_percentage}, the commits by each user $C_u$ are divided by the total number of commits $|C|$, yielding the set of contribution percentages $P$ for the repository.
 Elements of $P$ are then assigned to ranges $0-5\%$, $5-10\%$, $10-50\%$, $50-95\%$, and $95-100\%$ (with shared values inclusive to the larger range).

### 2013-03-29 paper/main.tex
Updated python code to only sample organization repositories.

@@
 \begin{abstract}
-Open Source Software (OSS) has produced many successful projects.
-The development process by which these projects are produced is generally unstructured compared to commercial software -- but a definite pattern does arrive, and is no less a pattern.
-GitHub, a popular OSS code hosting website, and Git, the site's Source Code Management (SCM) tool of choice, may have the potential to fundamentally change ths process.
+Open Source Software (OSS) projects, or software projects with publicly available source code, are realizing ever more important roles in both personal and business computing.
+As such, shifts in the way OSS is developed could have impacts on both the quantity and the quality of OSS projects.
 
-By analyzing a subset of GitHub repositories, this report will show how GitHub has influenced some very intrinsic aspects of traditional OSS development, such as developer hierarchies and issue close velocity.
-We find that many of the traditional aspects of OSS development remain, such most project development is done by a small group of core developers.
-Others assumptions about project hierarchy, such as a large number of Issue Reporters compared to Committers, seems unsupported by the data.
+The development process by which these projects are produced is generally unstructured compared to commercial software, but many projects do exhibit general development patterns.
+GitHub, a popular OSS code hosting website, along with Git, the site's Source Code Management (SCM) tool of choice, may have the potential to fundamentally change this process by facilitating new patterns and opportunities for developers.
+
+By analyzing a subset of GitHub repositories, this report will show how GitHub has influenced some intrinsic aspects of traditional OSS development, such as developer hierarchies and issue close velocity.
+We find that many of the traditional aspects of OSS development remain, such as most project development is done by a small group of core developers.
+Other assumptions about project hierarchy, such as a large number of Issue Reporters compared to Committers, seems unsupported by the data.
 \end{abstract}
@@
 Contrast the Apache Server project to a successful project on GitHub.
-Although the spirit and intent may be the same, the tool set is drastially different.
+Although the spirit and intent may be the same, the tool set is drastically different.
 The intent of this work is to explore how GitHub facilitates this process, and also how it may cause it to evolve.
@@
 These questions will draw upon data from this work, but also reference interviews with GitHub staff to better represent how GitHub may be changing how OSS development is conducted.
+These qualitative observations, paired with the more qualitative data analysis, allows for a broad description of the way that GitHub is changing the Software Engineering landscape for OSS.
 
 \subsection{Hypotheses}
-The following hypotheses
+The following hypotheses are presented in an effort to better understand the GitHub development process by analyzing data from actual GitHub projects.
+Each hypothesis will be backed by previous research in an attempt to show concrete ways in which the GitHub process may be changing the OSS development process.
+Where possible, results from previous research on the Apache Server[mockus2000case] is compared, as this work closely aligns with Hypothesis 2 and 3.
 
@@
 \subsection{Research Questions}
+To expand the analysis scope slightly, three qualitative research questions are posed.
+These, although backed by analyzed data where possible, are intended to show broader patterns, ideas, and motivations around the use of GitHub.
+Where possible, interviews\cite{begel2013social} where conducted to explore high level concepts and motivations.
 
@@
 \textit{Interview Q2: Many new Software Developers are will be starting their education/training already having experience with GitHub.
-How will Software Engineering as a discipline (and how it is taught) change to accodate the ideas popularized by GitHub?}\\
+How will Software Engineering as a discipline (and how it is taught) change to accommodate the ideas popularized by GitHub?}\\
 

### 2013-03-29 code/research.py

@@ -8,12 +8,16 @@ import random
 from sqlalchemy import create_engine
 from sqlalchemy import MetaData, Column, Table, ForeignKey
 from sqlalchemy import Integer, String, DateTime
+from sqlalchemy import func
+from sqlalchemy.orm import sessionmaker
 
 from config import *
  
 engine = create_engine('mysql://'+db_username+'@'+db_host+':'+str(db_port)+'/'+db_database, echo=False)
  
 metadata = MetaData(bind=engine)
+Session = sessionmaker(bind=engine)
+db_session = Session()
  
 repository_table = Table('repository', metadata,
                     Column('id', Integer, primary_key=True),
@@ -42,7 +46,6 @@ commit_table = Table('commit', metadata,
                     mysql_engine='InnoDB'
                     )
  
- 
 # create tables in database
 metadata.create_all()
 
@@ -52,8 +55,14 @@ session.auth = (github_username, github_password)
 response = urllib2.urlopen('https://raw.github.com/gist/4669395')
 random_words = response.readlines()
 
+def _get_user_organizations(user):
+	response = session.get('https://api.github.com/users/'+user+"/orgs")
+	_check_quota(response)
+	if(response.ok):
+		orgs = json.loads(response.text or response.content)
+		return [org['login'] for org in orgs]
+
 def _get_random_repo():
-	repo_json = None
 	while True:
 		keyword = random.choice(random_words)
 		response = session.get('https://api.github.com/legacy/repos/search/'+keyword)
@@ -62,19 +71,24 @@ def _get_random_repo():
 			repos = json.loads(response.text or response.content)
 			if(len(repos['repositories']) > 0):
 				repo = random.choice(repos['repositories'])
-				response = session.get('https://api.github.com/repos/'+repo['username']+'/'+repo['name'])
-				if(response.ok):
-					return json.loads(response.text or response.content)
+				userame = repo['username']
+				orgs = _get_user_organizations(userame)
+				for org in orgs:
+					response = session.get('https://api.github.com/users/'+org+'/repos')
+					_check_quota(response)
+					if(response.ok):
+						repos = json.loads(response.text or response.content)
+						return random.choice(repos)
 
 def _check_quota(response):
 	requests_left = int(response.headers['X-RateLimit-Remaining'])
 	if(requests_left == 0):
 		print "Sleeping for 65 minutes... Good Night."
 		time.sleep(65 * 60)
[…]

### 2013-04-01 code/research.py
Run using organisation data only.

@@ -32,7 +32,6 @@ issue_table = Table('issue', metadata,
 					Column('id', Integer, autoincrement=True, primary_key=True),
                     Column('repository_id', Integer, ForeignKey('repository.id')),
                     Column('creator', String(60)),
-                    Column('name', String(60)),
                     Column('number', Integer),
                     Column('open_date', DateTime),
                     Column('close_date', DateTime),
@@ -148,4 +147,4 @@ def crawl(sample_size):
 		except Exception as e:
 			print e
 
-crawl(100)
+crawl(300)

### 2013-04-11 paper/main.tex
Updated paper.

@@
 Although the spirit and intent may be the same, the tool set is drastically different.
-The intent of this work is to explore how GitHub facilitates this process, and also how it may cause it to evolve.
+The intent of this work is to explore how GitHub facilitates this process, and also how it may be causing it to evolve.
 
 GitHub\footnote{https://github.com/} is a popular code hosting website that uses Git Source Code Management (SCM)\footnote{http://git-scm.com/}.
-GitHub...
-Git allows for development to take place in a more distributed manner than previously available in other SCM systems\cite{spinellis2012git}.
+Along with hosting software repositories, GitHub incorporates social aspects of development.
+GitHub users have community-visible profiles, and user actions can be tracked and followed by other community actions.
+This type of social integration of a user's identity and actions is unique to GitHub\cite{dabbish2012social}.
+
+Git itself, the SCM of GitHub, lends itself to the collaborative nature of GitHub by allowing for development to take place in a more distributed manner than previously available in other SCM systems\cite{spinellis2012git}.
 Git also allows for a variety of work flows\cite{chacon2009pro}.
 These workflows may be tailored to the individual needs of a project.
-The Linux Kernel, for example, uses a Dictator and
-Lieutenants Workflow\cite{platschekfloss}, which is hierarchical in nature.
+The Linux Kernel, for example, uses a Dictator and Lieutenants Workflow\cite{platschekfloss}, which is hierarchical in nature.
 Other work flows may be more distributed, or resemble more traditional centralized SCM systems.
@@
 Each hypothesis will be backed by previous research in an attempt to show concrete ways in which the GitHub process may be changing the OSS development process.
-Where possible, results from previous research on the Apache Server[mockus2000case] is compared, as this work closely aligns with Hypothesis 2 and 3.
+Where possible, results from previous research on the Apache Server[mockus2000case] is compared, as this work closely aligns with Hypothesis 2 and 3.\\
 
@@
 As \emph{watches} don't necessarily signal intent to contribute, a positive direction for an OSS project is to not only increase the number of \emph{watchers}, but also the number of \emph{forks}.
-As both signify a general general interest in the project, it is expected that there be some correlation between the two.
+As both signify a general general interest in the project, it is expected that there be some correlation between the two.\\
 
@@
 This is, ultimately, a proof of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub.
-In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.
+In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.\\
 
@@
 These, although backed by analyzed data where possible, are intended to show broader patterns, ideas, and motivations around the use of GitHub.
-Where possible, interviews\cite{begel2013social} where conducted to explore high level concepts and motivations.
+Where possible, interviews\cite{begel2013social} where conducted to explore high level concepts and motivations.\\
 
@@
 At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers.
-Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?
+Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?\\
 
@@
 Matplotlib\cite{Hunter2007}, a 2D graphics package for Python, was used to render graphs and charts.
-NumPY and SciPY\cite{scipy} were used for data analysis calculations, such as correlation coefficients.
+NumPy and SciPy\cite{scipy} were used for data analysis calculations, such as correlation coefficients.
 
@@
 \subsection{Hypotheses}
-\textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
-A correlation coefficient of \input{watcher_forks_scatterplot-correlation} was observed between the number of Watchers and Forks of a repository.
+\noindent \textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
+A correlation coefficient of \input{watcher_forks_scatterplot-correlation}}\unskip was observed between the number of Watchers and Forks of a repository.
 This suggests that a correlation between Watchers and Forks exists.
@@
 
-\textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
+\noindent \textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
 No significant correlation was found between the number of repository forks and issue resolution time.
@@
 Research does support a relationship\cite{mockus2002two} between the stated priority and fix interval.
-As the GitHub issue tracker can only support prioritization by free-text \textit{tags} placed on issues, it is not feasable to programmattically gather issue priority metrics.\\
+As the GitHub issue tracker can only support prioritization by free-text \textit{tags} placed on issues, it is not feasible to programmatically gather issue priority metrics.\\
 
-\textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude.}\\
+\noindent \textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude.}\\
 Data does not support an exponentially increasing number of distinct Issue Reporters as compared to distinct Committers. 
@@
 \includegraphics[height=3in,width=3in]{images/issue_reporters_committers_scatterplot.png}
-\caption{Repostitory issue reporters and committers}
+\caption{Repository issue reporters and committers}
 \label{fig:issue_reporters_committers_scatterplot}
@@
 Figure\ref{fig:issue_reporters_committers_scatterplot}, does however suggest, however, that that the two variables may be related.
-The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation cofficient of \input{issue_reporters_committers_scatterplot-correlation}.
+The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation cofficient of \input{issue_reporters_committers_scatterplot-correlation}}\unskip.
 More investigation would be necessary to further examine this finding.\\
@@
 \subsection{Research Questions}
-\textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
+\noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
 The data in figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown.
@@
 
-To further analyze this research question, and to explore further the GitHub ``society'' and its impact on OSS development, an interview with GitHub staff was conducted.\\\\
-\textit{Interview Q1: How is GitHub changing the way OSS is being developed?}\\
-
-\textit{Interview Q2: Many new Software Developers are will be starting their education/training already having experience with GitHub.
-How will Software Engineering as a discipline (and how it is taught) change to accommodate the ideas popularized by GitHub?}\\
-
-\textit{Interview Q3: There are many non-code artifacts being hosted on GitHub, for example, Object Management Group Specifications, Laws and Statutes, books...
-etc.
-Do you see GitHub participation expanding in the non-code areas?}\\
-
-\textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
+\noindent \textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
 The OMG\textregistered standardization process is a process by which products strive to produce an interoperability statndard by way of an industry standards consortium.\cite{omgTechProcess}
-The standardization process itself aims to, at its conclusion, produce a set of model artificats containing sufficient detail such that when interpreted and implemented, produce interoperable software.
+The standardization process itself aims to, at its conclusion, produce a set of model artifacts containing sufficient detail such that when interpreted and implemented, produce interoperable software.
 At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process.
@@
 
-As mentioned, the CTS2 specification made heavy use of the GitHub issue tracker, as well as another OMG specification, ServD\footnote{https://github.com/servd/servd-specification}.
-For open source specifications, or requirements engineering activities, GitHub has shown to be a viable platform.
+As mentioned, the CTS2 specification made heavy use of the GitHub issue tracker, as well as another OMG specification, ServD.\footnote{https://github.com/servd/servd-specification} For open source specifications, or requirements engineering activities, GitHub has shown to be a viable platform.
 
@@
 \section{Acknowledgments}
+We thank Gilberto Fragoso, Ph.D. and Paul Alexander for insight and ideas, Harold Solbrig, Craig Stancl, and Dick Hedger for guidance, and Rick Kiefer for his review.
 
@@
 
+\input{interview}
+
 \end{document}

### 2013-04-11 code/plots.py
Updated code for plots.

@@ -6,7 +6,7 @@ committers_histogram.plot("committers_histogram")
 
 committers_percentage_query = '''
 select 
-	cast(cast( (count(*) / commits.cnt ) as decimal(2,2)) * 100 as unsigned integer)
+	cast(cast( (count(*) / commits.cnt ) as decimal(3,2)) * 100 as unsigned integer)
 from
     (select 
         count(*) as cnt, c1.repository_id
@@ -14,8 +14,8 @@ from
         commit c1
     group by c1.repository_id) commits
         inner join
-    commit c1 ON c1.repository_id = commits.repository_id
-group by c1.repository_id , c1.committer
+    commit c2 ON c2.repository_id = commits.repository_id
+group by c2.repository_id, c2.committer
 '''
 committers_percentage_histogram = Histograph(committers_percentage_query, "% Of Project Commits", "Committers")
 committers_percentage_histogram.plot("committers_percentage_histogram")
@@ -74,8 +74,5 @@ group by repository_id
 issue_close_time_forks_scatterplot = ScatterPlot(issue_close_time_forks_query, "Close Time", "Forks", (0,200), (0,200))
 print issue_close_time_forks_scatterplot.plot("issue_close_time_forks_scatterplot")
 
-
-
 committers_percentage_pie_chart = PieChart(committers_percentage_query)
 committers_percentage_pie_chart.plot("committers_percentage_pie_chart")
-

### 2013-04-11 code/plotter.py

@@ -65,14 +65,12 @@ class ScatterPlot:
 
 		fig = plt.figure() 
 		m,b = numpy.polyfit(x,y,1)
-		print("Slope: %s, B: %s") % (m,b)
-
-		plt.plot(x,y,'bo',x,m*numpy.array(x)+b,'-k',linewidth=2)
 
 		plt.scatter(x,y)
 
-		correlation_coefficient = numpy.corrcoef(x,y)[0,1]
-		title('Correlation Coefficient: ' + str(correlation_coefficient))
+		correlation_coefficient = "%.4f" % numpy.corrcoef(x,y)[0,1]
+		title('Correlation Coefficient: ' + correlation_coefficient)
+		plt.plot(x,y,'bo',x,m*numpy.array(x)+b,'-k',linewidth=2)
 
 		plt.xlabel(self.xlabel)
 		plt.ylabel(self.ylabel)

### 2013-04-11 code/research.py

@@ -147,4 +147,4 @@ def crawl(sample_size):
 		except Exception as e:
 			print e
 
-crawl(300)
+crawl(5000)

### 2013-04-11 paper/main.tex
First run through proof-reading.

@@
 \usepackage{amsmath}
+\usepackage{mathrsfs}
 %\setlength{\parskip}{\baselineskip}
@@
 
-GitHub\footnote{https://github.com/} is a popular code hosting website that uses Git Source Code Management (SCM)\footnote{http://git-scm.com/}.
-Along with hosting software repositories, GitHub incorporates social aspects of development.
+GitHub\footnote{https://github.com/} is a popular code hosting website that uses Git Source Code Management (SCM).\footnote{http://git-scm.com/} Along with hosting software repositories, GitHub incorporates social aspects of development.
 GitHub users have community-visible profiles, and user actions can be tracked and followed by other community actions.
@@
 Git also allows for a variety of work flows\cite{chacon2009pro}.
-These workflows may be tailored to the individual needs of a project.
+These work flows may be tailored to the individual needs of a project.
 The Linux Kernel, for example, uses a Dictator and Lieutenants Workflow\cite{platschekfloss}, which is hierarchical in nature.
@@
 In addition, three supporting research questions are proposed.
-These questions will draw upon data from this work, but also reference interviews with GitHub staff to better represent how GitHub may be changing how OSS development is conducted.
+These questions will draw upon data from this work, but also reference several case studies to better represent how GitHub may be changing how OSS development is conducted.
 These qualitative observations, paired with the more qualitative data analysis, allows for a broad description of the way that GitHub is changing the Software Engineering landscape for OSS.
@@
 This feeling of belonging can be expressed passively, as a \emph{watch}, or actively, as a \emph{fork} of a repository.
+
+In GitHub, a \emph{watch} is a user act of subscribing to changes in a repository.
+It is also referred to as a \emph{star}.
+A \emph{watch} allows users to receive status updates, such as when code has been committed.
+This is considered a \emph{passive} act, as it signifies involvement in the community, but no direct action to contribute.
+
+A \emph{fork} is the creation of a personal copy of a repository\cite{dabbish2012social}, from which they can change without impacting the original code source.
+A \emph{fork} is an \emph{active} action, as it signifies both involvement and intent to contribute.
+It should be noted, however, that it only signifies \emph{intent} -- a fork of a repository could be made and no other actions taken.
+
 As \emph{watches} don't necessarily signal intent to contribute, a positive direction for an OSS project is to not only increase the number of \emph{watchers}, but also the number of \emph{forks}.
@@
 ``Patches'' may be released at any time, by any member of the community\cite{mockus2000case}.
-``Patches'' in the context of GitHub could be equated to Pull Requests.
+``Patches'' in the context of GitHub could be equated to \emph{Pull Requests}.
+A \emph{Pull Requests} is an announcement by a \emph{fork} that there are changes that may be appropriate to add back to the main repository.
+The owners of the main repository may then audit these requests, and \emph{pull} some or all of them back into the main repository.
 
-This is, ultimately, a proof of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub.
+This hypothesis is, ultimately, an examination of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub.
 In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.\\
@@
 These, although backed by analyzed data where possible, are intended to show broader patterns, ideas, and motivations around the use of GitHub.
-Where possible, interviews\cite{begel2013social} where conducted to explore high level concepts and motivations.\\
+Where possible, interviews\cite{begel2013social} were conducted to explore high level concepts and motivations.\\
 
 \noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
-A smal core of developers, however, seems to be a historical fact of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}.
+A small core of developers, however, seems to be a historical fact of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}.
 At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers.
@@
 \noindent \textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
-The Object Management Group\textregistered\footnote{http://www.omg.org} computer industry consortium focused on technology standards.
+The Object Management Group\textsuperscript{\textregistered}\footnote{http://www.omg.org} computer industry consortium focused on technology standards.
 Potential standard submission teams must go through a vetting process\cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard.
 Gathering and recording this feedback is a challenge, but GitHub's issue tracking system may be able to streamline the process.
+This just one example of how GitHub is being used for non-code artifacts.
+Are organizations exploring this type of work flow?
 
@@
 \subsection{Collection and Storage}
-Data collection was done using the GitHub REST API\footnote{http://developer.github.com/v3/}. A random selection of 1000 repositories was selected using Algorithm \ref{algo:data_collection_algo}.
+Data collection was done using the GitHub REST API.\footnote{http://developer.github.com/v3/} A random selection of 1000 repositories was selected using Algorithm \ref{algo:data_collection_algo}.
 
@@
 \State $i \gets i + 1$
-\State $word \gets random(word\_list)$
-\State $repos \gets git\_search(word)$
-\State $j \gets $%\Call{random}{0... \Call{size}{repos}}$
+\State $word \gets \textsc{random}(word\_list)$
+\State $repos \gets \textsc{github\_search}(word)$
+\State $j \gets \textsc{random}(repos)$
 \State $repo \gets repos_j$
-\State \Call{store}{repo}
+\If { $\textsc{isOrgRepo}(repo)$  }
+\State $\textsc{store}(repo)$
+\EndIf 
 \EndWhile
@@
 
+Repositories were gathered at random, but only repositories owned by an \textit{Organization} GitHub User type were considered.
+The \textsc{isOrgRepo} function will filter out all non \textit{Organization} owned repositories.
 The \textsc{store} function accepts as an input a given random repository.
@@
 
-All source code and materials used in data collection and analysis are hosted at GitHub (naturally!), and may be accessed by the following URL: \url{https://github.com/kevinpeterson/github-process-research}
+All source code and materials used in data collection and analysis are hosted at GitHub (naturally!), and may be accessed by the following URL: \url{https://github.com/kevinpeterson/github-process-research}.
 
 \subsection{Analysis}
+As stated, the main focus of data collection was to gather metrics pertaining to GitHub \textit{Repositories}.
 In this context, a \textit{Repository} is defined as a Set of \textit{Attributes}.
@@
 See Section \ref{sec:attributes} for further information on individual \textit{Attributes}.
-These metrics of interest are used as an Indexing Set $A^\prime$, and have labels as follows:
+For analysis purposes, these metrics of interest are used as an Indexing Set $A^\prime$, and have labels as follows:
 
@@
 
-All analyzed \textit{Respositories} can be thought of as the Indexed Family of \textit{Attribute} Sets $A$, or a Set of Sets of \textit{Attributes}.
-Let the Set of these Sets of \textit{Attribute} Sets be called $S$, or the \textit{Repository Sample}.
-We can label the \textit{Attribute} Sets by arbitrary integers 0 through $|S|$, or:
+All analyzed \textit{Respositories} can be thought of as the Indexed Family of \textit{Attribute} Sets $A$, or a set of sets of \textit{Attributes}.
+Let the set of these sets of \textit{Attributes} be called $S$, or the \textit{Repository Sample}.
+We can label the \textit{Attribute} sets by arbitrary integers 0 through $|S|$, or:
 \[ \{A_i\}_{i \in \{0,1,2 ... |S|\} }\]
 
-In this way, we can reference a Set of \textit{Attributes} (or, an individual \textit{Repository}), by $A_i$.
+In this way, we can reference a set of \textit{Attributes} (or, an individual \textit{Repository}), by $A_i$.
 
@@
 The number of individual source code commits to a repository.
-It is important to note that a textit{commit} is a single transaction that changes the code contents of a repository.
-Actual commit ``size'' (lines of code modified, for example), is not taken into account.
+It is important to note that a \textit{commit} is a single transaction that changes the code contents of a repository.
+Actual commit ``size'' (lines of code modified, for example) is not taken into account.
 
@@
 It is important to note, however, that issue resolution time is not necessarily related to software product quality in the context of GitHub.
-We can assert that for serveral reasons.
+We can assert that for several reasons.
 \textit{First}, GitHub issues are not necessarily product defects.
-Issues may be feature requests.
-TODOs, or general support questions.
-\textit{second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all.
-\textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub -- something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric \cite{phamcreating}.
+Issues may be feature requests, comments, or general support questions.
+\textit{Second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all.
+\textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub -- something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric\cite{phamcreating}.
 
@@
 Figure\ref{fig:issue_reporters_committers_scatterplot}, does however suggest, however, that that the two variables may be related.
-The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation cofficient of \input{issue_reporters_committers_scatterplot-correlation}}\unskip.
+The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation coefficient of \input{issue_reporters_committers_scatterplot-correlation}}\unskip.
 More investigation would be necessary to further examine this finding.\\
@@
 \noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
-The data in figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown.
-The inner percentages represent the percentage of committers in a project that that have committed the outer percentage amount of commits to that project, or:
+The data in Figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown.
+The intent of this measurement is to show how the number of repository commits are distributed among it's committers.
+The slices represent the distribution of committers that that have committed the outer percentage amount of commits to a project, or:
 
 \begin{equation}
-P = \left\{ \frac{ C_u } { |C| } \times 100 \Big| u \in U \right\}
+P = \left\{ \frac{ C_u } { |\mathscr{C}| } \times 100 \Big| u \in U \right\}
 \label{eq:commit_percentage}
@@
 
-Where $U$ is the Set of \textit{Committers} to a given \textit{Repository} and $C$ is the MultiSet of \textit{Commits} by an individual \textit{Committers} to a given \textit{Repository}, or:
+Where $U$ is the Set of \textit{Committers} to a given \textit{Repository}, $\mathscr{C}$ is the set of all \textit{Commits} to a \textit{Repository}, and $C_u$ denotes sets of \textit{Commits} by an individual \textit{Committers} to a \textit{Repository}, or:
  \[ \{ C_u \}_{u \in U} \]
+
+Each \textit{Committers}, then, will produce a subset of all \textit{Commits} to a a \textit{Repository}, or $C_u \subset \mathscr{C}$.
 
@@
 \includegraphics[height=3in,width=3in]{images/committers_percentage_pie_chart.png}
-\caption{Individual committer commit percentage per repository}
+\caption{Distribution of individual committer commit percentage of a repository}
 \label{fig:committers_percentage_pie_chart}
@@
 
-To find the percentages Set $P$ of equation\ref{eq:commit_percentage}, the commits by each user $C_u$ are divided by the total number of commits $|C|$, yielding the set of contribution percentages $P$ for the repository.
+To find the percentages Set $P$ of Equation\ref{eq:commit_percentage}, the commits by each user $C_u$ are divided by the total number of commits $|\mathscr{C}|$, yielding the set of contribution percentages $P$ for the repository.
 Elements of $P$ are then assigned to ranges $0-5\%$, $5-10\%$, $10-50\%$, $50-95\%$, and $95-100\%$ (with shared values inclusive to the larger range).
@@
 
-There does appear to be evidence of a small set of core developers, denoted by the relatively high percentage of developers that have authored $95-100\%$ of the commits of a given repository.
-Most notably, however, is that the data shows that by a large percentage of developers committed small amounts ($0-5\%$ of the total repository commits).
+Most notably, the data shows that a large percentage of developers committed small amounts ($0-5\%$ of the total repository commit).
 This means that a small core are not entirely monopolizing the commits, but rather opportunity exists for users to contribute, even if in small ways.
+Instances of developer monopolization of commits to a repository (for example, where a developer contributes $95-100\%$ of the total commits), are shown to be small.
 
@@
 
-When looking strictly at the number of committers to a repository, however, the notion of a small set of core Committers is strongly supported, as the vast majority of repositories have fewer than 10 committers (figure \ref{fig:committers_histogram}).\\
+When looking strictly at the number of committers to a repository, however, the notion of a small set of core committers is strongly supported, as the vast majority of repositories have fewer than 10 committers (Figure \ref{fig:committers_histogram}).\\
 
-\textbf{Research Question 2: How is GitHub changing the OSS process?}\\
+\noindent \textbf{Research Question 2: How is GitHub changing the OSS process?}\\
 Thung \textit{et al.} show that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enable high collaboration rates.\cite{thung2013network} In fact, the developer connectivity was shown to be higher than SourceForge and even Facebook.
-In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal for.
+In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal.
 
-Ye also observed\cite{ye2003toward} that the contribution environment, or ``society,'' is important in providing a framework for develper trust building:
+Ye also observed\cite{ye2003toward} that the contribution environment, or ``society,'' is important in providing a framework for developer trust building:
 \begin{quote}
@@
 
+With this process change can also come cause for concern.
+The move to Git from a centralized version control system (such as SVN or CVS) is not without its challenges.
+Differences in philosophy and nomenclature may place added strain on developers\cite{bird2009promises}.
+Also, it is not to be assumed that process change is always welcome\cite{de2009software}, even if it is an improvement.
+Developers and organizations have existing processes, SCM tools, and workflow patterns -- which all change at a cost.
+
+Despite the potential migration cost, the National Center for Biomedical Ontology,\footnote{https://github.com/ncbo}, the National Cancer Informatics Program\footnote{https://github.com/ncip}, NASA\footnote{https://github.com/nasa}, and the WhiteHouse\footnote{https://github.com/WhiteHouse} are just some organizations that have moved all or portions of development to GitHub.
+As these and other organizations consider the move to GitHub, the distributed and ``social'' aspects of the OSS process will move continue to evolve.\\
+
 \noindent \textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
-The OMG\textregistered standardization process is a process by which products strive to produce an interoperability statndard by way of an industry standards consortium.\cite{omgTechProcess}
+The OMG\textsuperscript{\textregistered} standardization process is a process by which products strive to produce an interoperability standard by way of an industry standards consortium.\cite{omgTechProcess}
 The standardization process itself aims to, at its conclusion, produce a set of model artifacts containing sufficient detail such that when interpreted and implemented, produce interoperable software.
@@
 
-As a case study, the CTS2 OMG\textregistered Specification\cite{cts2} used a GitHub issue tracker\footnote{https://github.com/cts2/cts2-specification/issues} to track specification changes.
+As a case study, the CTS2 OMG\textsuperscript{\textregistered} Specification\cite{cts2} used a GitHub issue tracker\footnote{https://github.com/cts2/cts2-specification/issues} to track specification changes.
 Elicitation, Analysis and Negotiation were natural fits for the issue tracker.
@@
 Because customer involvement is critical to these activities\cite{paetsch2003requirements}, ease of contribution was a main concern.
-OMG has a predescribed format for requirements Documentation, but using the GitHub issue API\footnote{http://developer.github.com/v3/issues/}, much of this documentation can be automated.
+OMG\textsuperscript{\textregistered} has a predescribed format for requirements Documentation, but using the GitHub issue API\footnote{http://developer.github.com/v3/issues/}, much of this documentation can be automated.
 Validation and Management are areas of requirements engineering that GitHub has interesting solutions for.
@@
 This level of traceability is powerful, as it allows every change to the specification artifacts to be tracked consistently to an open issue.
-As the normal OMG process proceeds, requirements can be managed through ``tagging'' the issues.
-By allowing issues to be grouped by arbitrary tags, users are more free to adapt their own workflow.
+As the normal OMG\textsuperscript{\textregistered} process proceeds, requirements can be managed through ``tagging'' the issues.
+By allowing issues to be grouped by arbitrary tags, users are more free to adapt their own work flow.
 For instance, when community voting on issues needed to be done for CTS2, a simple ``ready for vote'' tag signified to the voting board exactly what was under consideration
 
-As mentioned, the CTS2 specification made heavy use of the GitHub issue tracker, as well as another OMG specification, ServD.\footnote{https://github.com/servd/servd-specification} For open source specifications, or requirements engineering activities, GitHub has shown to be a viable platform.
+As mentioned, the CTS2 specification made heavy use of the GitHub issue tracker, as well as another OMG\textsuperscript{\textregistered} specification, ServD.\footnote{https://github.com/servd/servd-specification} For open source specifications, or requirements engineering activities, GitHub has shown to be a viable platform.
 
 \section{Conclusion}
+The data suggests an evolution in process, but not necessarily a large shift.
+A small subset of core developers, a traditional aspect of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}, seems supported by both the summary data in Figure \ref{eq:variance}, and the distribution of committers per repository in Figure \ref{fig:committers_histogram}.
+Figure \ref{fig:committers_percentage_pie_chart} shows that a significant number of participation, albeit small, in many repositories.
+This may indicate a lowered entry barrier for participation.
+
+The data summary variance statement in Assertion: \ref{eq:variance} is important, as it suggests large amounts of variety in the observed variables.
+This may suggest that not all repositories on GitHub are intended to be consumable software artifacts for users, or ``consumers.'' A large number of repositories may not be intended for customer consumption, but merely as example code, experimentation, or exercises.
+More research into categorizing GitHub repositories is needed.
+This is one reason why only repositories owned by a GitHub \textit{Organization} were analyzed, as it was anticipated that this would yield a better sample of repositories for analysis.
+
+The non-code artifacts in GitHub may fall outside of the domain of software development, but it is a significant aspect of our findings.
+From government documents to software specifications, GitHub is being used and used successfully.
+More research is needed into what role GitHub will play in the work flow of these artifacts, and whether it will influence their traditional work flow, as it has for software development.
 

### 2013-04-11 code/plots.py

@@ -1,7 +1,7 @@
 from plotter import Histograph, ScatterPlot, PieChart
 
 committers_query = '''select count(distinct committer) from commit c group by repository_id'''
-committers_histogram = Histograph(committers_query, "Committers", "Repositories", (1,25))
+committers_histogram = Histograph(committers_query, "Committers", "Repositories", (1,20))
 committers_histogram.plot("committers_histogram")
 
 committers_percentage_query = '''
@@ -34,7 +34,7 @@ issue_close_time_histogram.plot("issue_close_time_histogram")
 
 watcher_forks_scatterplot_query = '''select watchers, forks from repository'''
 watcher_forks_scatterplot = ScatterPlot(watcher_forks_scatterplot_query, "Watchers", "Forks", (0,1000), (0,150))
-print watcher_forks_scatterplot.plot("watcher_forks_scatterplot")
+watcher_forks_scatterplot.plot("watcher_forks_scatterplot")
 
 issue_reporters_committers_query = '''
 select 
@@ -53,7 +53,7 @@ from
     group by repository_id) issues ON issues.repository_id = commits.repository_id
 '''
 issue_reporters_committers_scatterplot = ScatterPlot(issue_reporters_committers_query, "Committers", "Issue Reoprters", (0,140), (0,140))
-print issue_reporters_committers_scatterplot.plot("issue_reporters_committers_scatterplot")
+issue_reporters_committers_scatterplot.plot("issue_reporters_committers_scatterplot")
 
 issue_close_time_forks_query = '''
 select 
@@ -72,7 +72,7 @@ from
 group by repository_id
 '''
 issue_close_time_forks_scatterplot = ScatterPlot(issue_close_time_forks_query, "Close Time", "Forks", (0,200), (0,200))
-print issue_close_time_forks_scatterplot.plot("issue_close_time_forks_scatterplot")
+issue_close_time_forks_scatterplot.plot("issue_close_time_forks_scatterplot")
 
 committers_percentage_pie_chart = PieChart(committers_percentage_query)
 committers_percentage_pie_chart.plot("committers_percentage_pie_chart")

### 2013-04-11 code/plotter.py

@@ -34,7 +34,10 @@ class PieChart:
 				ranges['95-100'] += 1
 
 		fig = plt.figure()
-		plt.pie(ranges.values(), labels=[key + "%" for key in ranges.keys()], autopct='%.2f%%', shadow=True)
+
+		plt.pie(ranges.values(), labels=[key + "%" for key in ranges.keys()], autopct=None, shadow=True)
+
+		plt.legend(title="Contribution")
 
 		F = gcf()
 		F.savefig("../paper/images/"+filename+".png")
@@ -70,7 +73,9 @@ class ScatterPlot:
 
 		correlation_coefficient = "%.4f" % numpy.corrcoef(x,y)[0,1]
 		title('Correlation Coefficient: ' + correlation_coefficient)
-		plt.plot(x,y,'bo',x,m*numpy.array(x)+b,'-k',linewidth=2)
+
+		if float(correlation_coefficient) > 0.5:
+			plt.plot(x,y,'bo',x,m*numpy.array(x)+b,'-k',linewidth=2)
 
 		plt.xlabel(self.xlabel)
 		plt.ylabel(self.ylabel)

### 2013-04-12 paper/main.tex
Paper proofreading.

@@
 By analyzing a subset of GitHub repositories, this report will show how GitHub has influenced some intrinsic aspects of traditional OSS development, such as developer hierarchies and issue close velocity.
-We find that many of the traditional aspects of OSS development remain, such as most project development is done by a small group of core developers.
-Other assumptions about project hierarchy, such as a large number of Issue Reporters compared to Committers, seems unsupported by the data.
+We find that many of the traditional aspects of OSS development remain, such as most project development being done by a small group of core developers.
+Other traditional assumptions about OSS project hierarchy, such as a large number of Issue Reporters compared to Committers, seems unsupported by the GitHub data.
+We conclude that GitHub represents an evolution of the OSS development process, and not necessarily a large shift.
 \end{abstract}
@@
 A case study of the Apache Server project\cite{mockus2000case} shows that a dedicated community can produce software that rivals or exceeds commercial offerings.
-This study showed that even in the unstructured context of OSS, certain structures, hierarchies, and codes of conduct emerge.
+Furthermore, it showed that even in the unstructured context of OSS, certain structures, hierarchies, and codes of conduct emerge.
 Contrast the Apache Server project to a successful project on GitHub.
@@
 
-GitHub\footnote{https://github.com/} is a popular code hosting website that uses Git Source Code Management (SCM).\footnote{http://git-scm.com/} Along with hosting software repositories, GitHub incorporates social aspects of development.
+GitHub\footnote{https://github.com/} is a popular code hosting website that uses Git Source Code Management (SCM).\footnote{http://git-scm.com/} Along with hosting software repositories, GitHub incorporates the social aspects of development.
 GitHub users have community-visible profiles, and user actions can be tracked and followed by other community actions.
@@
 In addition, three supporting research questions are proposed.
-These questions will draw upon data from this work, but also reference several case studies to better represent how GitHub may be changing how OSS development is conducted.
+These questions will draw upon data from this work, but also reference several case studies to better represent how GitHub may be changing the way in which OSS development is conducted.
 These qualitative observations, paired with the more qualitative data analysis, allows for a broad description of the way that GitHub is changing the Software Engineering landscape for OSS.
@@
 The following hypotheses are presented in an effort to better understand the GitHub development process by analyzing data from actual GitHub projects.
-Each hypothesis will be backed by previous research in an attempt to show concrete ways in which the GitHub process may be changing the OSS development process.
-Where possible, results from previous research on the Apache Server[mockus2000case] is compared, as this work closely aligns with Hypothesis 2 and 3.\\
+Each hypothesis will be backed by previous research in an attempt to show concrete ways in which GitHub may be changing the OSS development process.
+Where possible, results from previous research on the Apache Server\cite{mockus2000case} is compared, as this work closely aligns with Hypothesis 2 and 3.\\
 
@@
 
-A \emph{fork} is the creation of a personal copy of a repository\cite{dabbish2012social}, from which they can change without impacting the original code source.
+A \emph{fork} is the creation of a personal copy of a repository\cite{dabbish2012social}, from which changes may be made without impacting the original code source.
 A \emph{fork} is an \emph{active} action, as it signifies both involvement and intent to contribute.
-It should be noted, however, that it only signifies \emph{intent} -- a fork of a repository could be made and no other actions taken.
+It should be noted, however, that it only signifies \emph{intent} -- a \emph{fork} of a repository could be made and no further actions taken.
 
 As \emph{watches} don't necessarily signal intent to contribute, a positive direction for an OSS project is to not only increase the number of \emph{watchers}, but also the number of \emph{forks}.
-As both signify a general general interest in the project, it is expected that there be some correlation between the two.\\
+As both signify a general general interest in the project, it is expected that there will be some correlation between the two.\\
 
@@
 ``Patches'' in the context of GitHub could be equated to \emph{Pull Requests}.
-A \emph{Pull Requests} is an announcement by a \emph{fork} that there are changes that may be appropriate to add back to the main repository.
+A \emph{Pull Request} is an announcement by a \emph{fork} that there have been changes made that may be appropriate to add back to the main repository.
 The owners of the main repository may then audit these requests, and \emph{pull} some or all of them back into the main repository.
@@
 \noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
-A small core of developers, however, seems to be a historical fact of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}.
-At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers.
+A small core of developers contributing to a project seems to be a well entrenched pattern of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}.
+At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers\cite{krishnamurthy2002cave}.
 Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?\\
@@
 \noindent \textbf{Research Question 2: How is GitHub changing the OSS process?}\\
-The social aspects of GitHub are an important part of the experience\cite{dabbish2012social}.
+The social aspects of GitHub are an important part of the development experience\cite{dabbish2012social}.
 User interactions and social pressures can drive OSS development in interesting ways.
@@
 \noindent \textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
-The Object Management Group\textsuperscript{\textregistered}\footnote{http://www.omg.org} computer industry consortium focused on technology standards.
+The Object Management Group\textsuperscript{\textregistered}\footnote{http://www.omg.org} is a computer industry consortium focused on technology standards.
 Potential standard submission teams must go through a vetting process\cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard.
@@
 \subsection{Collection and Storage}
-Data collection was done using the GitHub REST API.\footnote{http://developer.github.com/v3/} A random selection of 1000 repositories was selected using Algorithm \ref{algo:data_collection_algo}.
+Data collection was done using the GitHub REST API.\footnote{http://developer.github.com/v3/} A random selection of 1000 repositories was selected, and metrics were gathered.
+The selection process used for choosing the repository sample is described in Algorithm \ref{algo:data_collection_algo}.
 
@@
 Repositories were gathered at random, but only repositories owned by an \textit{Organization} GitHub User type were considered.
-The \textsc{isOrgRepo} function will filter out all non \textit{Organization} owned repositories.
-The \textsc{store} function accepts as an input a given random repository.
+After a random repository has been selected, the \textsc{isOrgRepo} function will filter out all non \textit{Organization} owned repositories.
+Next, the \textsc{store} function accepts as an input the given repository.
 The purpose of this function is to persist the given repository to a database for further analysis.
@@
 As stated, the main focus of data collection was to gather metrics pertaining to GitHub \textit{Repositories}.
-In this context, a \textit{Repository} is defined as a Set of \textit{Attributes}.
+In this context, a \textit{Repository} is defined as a set of \textit{Attributes}.
 An \textit{Attribute} is a metric of interest.
@@
 
-All analyzed \textit{Respositories} can be thought of as the Indexed Family of \textit{Attribute} Sets $A$, or a set of sets of \textit{Attributes}.
+All analyzed \textit{Respositories} can be thought of as the Indexed Family of \textit{Attribute} Sets, or a set of sets of \textit{Attributes}.
 Let the set of these sets of \textit{Attributes} be called $S$, or the \textit{Repository Sample}.
@@
 
-Each individual Attribute processed by some aggregate function $f$, or $f(  A_{i,r} )$.
+To gather the metics, each individual Attribute processed by some aggregate function $f$, or $f(  A_{i,r} )$.
 For this analysis, the aggregate functions are: \textsc{mean}, \textsc{min}, \textsc{max}, \textsc{stddev}.
@@
 \noindent \textbf{\textit{Issues:}}
-The number of issues posted to a repository's issue tracker\footnote{https://github.com/blog/411-github-issue-tracker}.
+The number of issues posted to a repository's issue tracker.\footnote{https://github.com/blog/411-github-issue-tracker}
 
@@
 Let $d(x)$ be a function that converts seconds to a Date, and let $s(x)$ be a function that converts a Date to seconds.
-Let $I$ equal the Set of \textit{Issues} in a given repository, with elements comprised of tuples $(\tau, \tau^\prime)$, where $\tau$ is \textit{Issue Open Time} and $\tau^\prime$ is \textit{Issue Close Time}.
+Let $I$ equal the Set of \textit{Issues} in a given repository, with elements comprised of tuples $(\tau, \tau^\prime)$, where $\tau$ is \textit{Issue Open Date} and $\tau^\prime$ is \textit{Issue Close Date}.
 We calculate \textit{Issue Close Time} by:
 
-\[ \Delta = d\left( \frac{\sum_{i \in I} s( p_{(1)}i ) - s( p_{(2)}i )  } {n} \right) \]
+\[ \Delta = d\left( \frac{\sum_{i \in I} s( p_{(2)}i ) - s( p_{(1)}i )  } {n} \right) \]
 
@@
 \textit{Second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all.
-\textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub -- something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric\cite{phamcreating}.
+\textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub.
+This, among other things, suggests that GitHub may be facilitating testing by making it more \emph{triable} by lowering entry barriers, and makng the process as a whole more \emph{observable}.
+This aspect is something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric\cite{phamcreating}.
 
@@
 Data does not support an exponentially increasing number of distinct Issue Reporters as compared to distinct Committers. 
-As shown in Figure\ref{fig:issue_reporters_committers_scatterplot}, the slope of the relationship between the two variables is less than one.
+As shown in Figure \ref{fig:issue_reporters_committers_scatterplot}, the slope of the relationship between the two variables is less than one.
 This does not seem to completely align with findings in the Apache Server project\cite{mockus2000case}.
@@
 
-Figure\ref{fig:issue_reporters_committers_scatterplot}, does however suggest, however, that that the two variables may be related.
+Figure \ref{fig:issue_reporters_committers_scatterplot}, does however suggest, however, that that the two variables may be related.
 The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation coefficient of \input{issue_reporters_committers_scatterplot-correlation}}\unskip.
@@
 
-Each \textit{Committers}, then, will produce a subset of all \textit{Commits} to a a \textit{Repository}, or $C_u \subset \mathscr{C}$.
+Each \textit{Committer}, then, will produce a subset of all \textit{Commits} to a a \textit{Repository}, or $C_u \subset \mathscr{C}$.
 
@@
 
-To find the percentages Set $P$ of Equation\ref{eq:commit_percentage}, the commits by each user $C_u$ are divided by the total number of commits $|\mathscr{C}|$, yielding the set of contribution percentages $P$ for the repository.
+To find the percentages, or set $P$ of Equation \ref{eq:commit_percentage}, the commits by each user $C_u$ are divided by the total number of commits $|\mathscr{C}|$, yielding the set of contribution percentages $P$ for the repository.
 Elements of $P$ are then assigned to ranges $0-5\%$, $5-10\%$, $10-50\%$, $50-95\%$, and $95-100\%$ (with shared values inclusive to the larger range).
@@
 
-Most notably, the data shows that a large percentage of developers committed small amounts ($0-5\%$ of the total repository commit).
+Most notably, the data shows that a large percentage of developers committed small amounts ($0-5\%$) of the total number of repository commits.
 This means that a small core are not entirely monopolizing the commits, but rather opportunity exists for users to contribute, even if in small ways.
@@
 
-Despite the potential migration cost, the National Center for Biomedical Ontology,\footnote{https://github.com/ncbo}, the National Cancer Informatics Program\footnote{https://github.com/ncip}, NASA\footnote{https://github.com/nasa}, and the WhiteHouse\footnote{https://github.com/WhiteHouse} are just some organizations that have moved all or portions of development to GitHub.
-As these and other organizations consider the move to GitHub, the distributed and ``social'' aspects of the OSS process will move continue to evolve.\\
+Despite the potential migration cost, the National Center for Biomedical Ontology,\footnote{https://github.com/ncbo} the National Cancer Informatics Program,\footnote{https://github.com/ncip} NASA,\footnote{https://github.com/nasa} and the WhiteHouse\footnote{https://github.com/WhiteHouse} are just some organizations that have moved all or portions of development to GitHub.
+As these and other organizations consider the move to GitHub, the distributed and ``social'' aspects of the OSS process will continue to evolve.\\
 
 \noindent \textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
-The OMG\textsuperscript{\textregistered} standardization process is a process by which products strive to produce an interoperability standard by way of an industry standards consortium.\cite{omgTechProcess}
+The OMG\textsuperscript{\textregistered} standardization process is a process by which products strive to produce an interoperability standard by way of an industry standards consortium\cite{omgTechProcess}.
 The standardization process itself aims to, at its conclusion, produce a set of model artifacts containing sufficient detail such that when interpreted and implemented, produce interoperable software.
@@
 Elicitation, Analysis and Negotiation were natural fits for the issue tracker.
-This was chosen in part because the barrier to participation was very small (only a GitHub account was needed), and issues tracker itself allowed for dialogs in the form of comments.
+This was chosen in part because the barrier to participation was very small (only a GitHub account was needed), and issue tracker itself allowed for dialogs in the form of comments.
 Because customer involvement is critical to these activities\cite{paetsch2003requirements}, ease of contribution was a main concern.
@@
 \section{Acknowledgments}
-We thank Gilberto Fragoso, Ph.D. and Paul Alexander for insight and ideas, Harold Solbrig, Craig Stancl, and Dick Hedger for guidance, and Rick Kiefer for his review.
+We thank Gilberto Fragoso, Ph.D., Paul Alexander, and the GitHub Staff for insight and ideas, Harold Solbrig, Craig Stancl, and Dick Hedger for guidance, and Rick Kiefer for his review.
 

### 2013-04-13 paper/main.tex
Sent for initial review.

@@
 At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers\cite{krishnamurthy2002cave}.
-Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?\\
+Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changed this?\\
 
@@
 
-\[A^\prime = \{i, c, i^\prime, i^{\prime\prime}, \Delta, \alpha, \beta\}\]
+\[A^\prime = \{i, c, i^\prime, i^{\prime\prime}, \Delta, \omega, \epsilon\}\]
 
-Where $i$ = {\textit{Issues}, $c$ = {\textit{Commits}, $i^\prime$ = {\textit{Closed Issues}, $i^{\prime\prime}$ = {\textit{Open Issues}, $\Delta$ = {\textit{Issue Close Time}, $\alpha$ = {\textit{Issue Creators}, and $\beta$ = {\textit{Committers}.
+Where $i$ = {\textit{Issues}, $c$ = {\textit{Commits}, $i^\prime$ = {\textit{Closed Issues}, $i^{\prime\prime}$ = {\textit{Open Issues}, $\Delta$ = {\textit{Issue Close Time}, $\omega$ = {\textit{Issue Creators}, and $\epsilon$ = {\textit{Committers}.
 
@@
 Going further, we can also denote an individual \textit{Attribute} of a given \textit{Repository} by:
-\[ \{A_{(i,r)}\}_{(i,r) \in A^\prime \times \{0,1,2 ... |S|\} } \]
-or, $A_{i,r}$. For example, $A_{\Delta,10}$ would denote the value of the \textit{Issue Close Time} ($\Delta$) of Repository \textit{10}.
+\[ \{A_{(i,\alpha)}\}_{(i,\alpha) \in A^\prime \times \{0,1,2 ... |S|\} } \]
+or, $A_{i,\alpha}$. For example, $A_{10,\Delta}$ would denote the value of the \textit{Issue Close Time} ($\Delta$) of Repository \textit{10}.
 
-To gather the metics, each individual Attribute processed by some aggregate function $f$, or $f(  A_{i,r} )$.
+To gather the metrics, the sets of all individual \textit{Attribute} across all \textit{Repositories} are processed by some aggregate function $f$.
+The input of this function is the union of all sets of a particular \textit{Attribute} over each \textit{Repository} in the sample.
 For this analysis, the aggregate functions are: \textsc{mean}, \textsc{min}, \textsc{max}, \textsc{stddev}.
-This analysis is applied to all $n$ repositories for which data was collected.
-Results were then averaged over all repositories.
+This analysis is applied to all repositories in the sample $S$ for which data was collected.
 Each attribute is analyzed independently, so given:
-\[ g(j) = \frac{\sum\limits_{r = 0}^{|S|} f(  A_{r,j}  ) }{|S|} \]
-each individual attribute is processed via $g(j)$ and will result in the average of the given aggregate function.
+\[ g(\alpha) =f \left( \bigcup\limits_{r = 0}^{|S|}  A_{r,\alpha} \right) \]
+each individual attribute is processed via $g(\alpha)$ and will result in the output of the given aggregate function.
+The result of each aggregate function when applied to the entire sample set of \textit{Respositories} of Table \ref{table:summary_stats}.
 
@@
 
-\[ \Delta = d\left( \frac{\sum_{i \in I} s( p_{(2)}i ) - s( p_{(1)}i )  } {n} \right) \]
+\[ \Delta = d\left( \frac{\sum_{i \in I} s\left(p_{2}(i) \right) - s\left(p_{1}(i) \right)  } {|I|} \right) \]
 
-Where for each issue $i$ in $I$, the \textit{Issue Open Date} $s( p_{(1)}i)$ is subtracted from the \textit{Issue Close Date} $s( p_{(2)}i)$, and $p_{(1)}i$ and $p_{(2)}i$ denote the elements of the tuple, or \textit{Issue Open Date} and \textit{Issue Close Date}, respectively.
+Where for each issue $i$ in $I$, the \textit{Issue Open Date} $s( p_{1}(i))$ is subtracted from the \textit{Issue Close Date} $s( p_{2}(i))$, where $p_{1}(i)$ and $p_{2}(i)$ denote projections of the tuple, or \textit{Issue Open Date} and \textit{Issue Close Date}, respectively.
 The result of this is averaged and converted into a Date via $d(x)$, resulting in the average issue close time interval $\Delta$.
@@
 
-\input{interview}
+%\input{interview}
 

### 2013-04-17 paper/github_interview.tex
Added the GitHub Staff interview transcript (Thanks to John Britton!).

\newpage

\section*{Appendix B: Interview Transcript}
Email interview with John Britton, GitHub Staff, April 17, 2013

\subsection*{Q: How is GitHub changing the way Open Source Software (OSS) is being developed today?}
GitHub moved the open source community away from a permission culture. Prior to GitHub contributing to an open source project was very difficult and often involved asking permission. By allowing people to fork a repository and make changes without asking for permission, GitHub has made contributing to someone else's project much easier than it has ever been. By submitting a Pull Request with suggested changes, project maintainers and contributors have a place to discuss changes to code that is already written and to refine it.

\subsection*{Q: Many new Software Developers will be starting their education/training already having experience with GitHub. How, in your opinion, will Software Engineering as a discipline (and how it is taught) change to accommodate the ideas popularized by GitHub?}

Software Engineering is moving more towards the development models seen in industry. It's much more useful to students to learn how to use the tools that they will see in industry than it is to learn how to use learning management systems and the like. I encourage teachers to assign project work in teams so that students gain experience working together and navigating the issues that arrive in a collaborative environment. While it's far from standard practice, I'm seeing more and more teachers encouraging and even requiring the use of version control, code review, and automated testing. Most of these tools have been standard in industry for quite a while, but only recently have they made their way into the classroom thanks to services like GitHub lowering the learning curve substantially.

\subsection*{There are many non-code artifacts being hosted on GitHub, for example, Object Management Group Specifications, Laws and Statutes, books... etc. Do you see GitHub participation expanding in the non-code areas?}
Most definitely. There are many examples of non-code artifacts hosted on GitHub:

\begin{itemize}
\item German law\footnote{https://github.com/bundestag/gesetze}
\item Images\footnote{https://github.com/blog/817-behold-image-view-modes}
\item 3d models\footnote{https://github.com/blog/1465-stl-file-viewing}
\item Research papers\footnote{https://github.com/karthikram/smb\_git}
\end{itemize}

That's just a small sample of some interesting things on GitHub, we're seeing new uses all the time.

### 2013-04-18 paper/main.tex
Added the GitHub Staff interview transcript (Thanks to John Britton!).
Finished adding in Rick's comments.

@@
 We find that many of the traditional aspects of OSS development remain, such as most project development being done by a small group of core developers.
-Other traditional assumptions about OSS project hierarchy, such as a large number of Issue Reporters compared to Committers, seems unsupported by the GitHub data.
+Other traditional assumptions about OSS developer hierarchy, such as a large number of Issue Reporters compared to Committers, seems unsupported by the GitHub data.
 We conclude that GitHub represents an evolution of the OSS development process, and not necessarily a large shift.
@@
 These questions will draw upon data from this work, but also reference several case studies to better represent how GitHub may be changing the way in which OSS development is conducted.
-These qualitative observations, paired with the more qualitative data analysis, allows for a broad description of the way that GitHub is changing the Software Engineering landscape for OSS.
+These qualitative observations, paired with the more quantitative data analysis, allows for a broad description of the way that GitHub is changing the Software Engineering landscape for OSS.
 
@@
 
-The Apache Project noted that a rapid response to problems can be obtained because OSS is not bound to release schedules in the same way commercial software is.
+The Apache Project noted that a rapid response to problems can be obtained because OSS is not bound to release schedules in the same way as commercial software.
 ``Patches'' may be released at any time, by any member of the community\cite{mockus2000case}.
@@
 User interactions and social pressures can drive OSS development in interesting ways.
-Social aspects of OSS development have existed before GitHub, with mailing lists\cite{mockus2000case}, gift giving mechanisms\cite{bergquist2008power}, and other code hosting sites.
-What is GitHub doing to facilitate change on OSS development, and how does it see OSS development changing moving forward?\\
+Social aspects of OSS development have existed before GitHub, with mailing lists\cite{mockus2000case}, gift giving mechanisms\cite{bergquist2008power}, and other collaboration tools.
+What is GitHub doing to facilitate change in OSS development, and how does it see OSS development changing moving forward?\\
 
-\noindent \textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
+\noindent \textbf{Research Question 3: Can GitHub be used for more than code artifacts?}\\
 The Object Management Group\textsuperscript{\textregistered}\footnote{http://www.omg.org} is a computer industry consortium focused on technology standards.
@@
 Data collection was done using the GitHub REST API.\footnote{http://developer.github.com/v3/} A random selection of 1000 repositories was selected, and metrics were gathered.
+Only repositories owned by a GitHub \textit{Organization} were considered.
+GitHub \textit{Organizations} are group-owned accounts\footnote{https://github.com/blog/674-introducing-organizations}, as opposed to normal individual user-owned accounts.
+\textit{Organization} owned repositories were chosen to increase the likelihood of analyzing high-participation projects.
 The selection process used for choosing the repository sample is described in Algorithm \ref{algo:data_collection_algo}.
@@
 
-Repositories were gathered at random, but only repositories owned by an \textit{Organization} GitHub User type were considered.
-After a random repository has been selected, the \textsc{isOrgRepo} function will filter out all non \textit{Organization} owned repositories.
+When a random repository has been selected, the \textsc{isOrgRepo} function will filter out all non \textit{Organization} owned repositories.
 Next, the \textsc{store} function accepts as an input the given repository.
@@
 It is important to note that a \textit{commit} is a single transaction that changes the code contents of a repository.
-Actual commit ``size'' (lines of code modified, for example) is not taken into account.
+Actual commit ``size'' (lines of code modified, number of files changed, for example) is not taken into account.
 
@@
 If the number of Watchers of a repository serves as a social cue for participation, even ``passive'' involvement with a project (Watching) could influence the ``active'' participation (Forking).
-This could possibly contribute to the correlation of the two variables.\\
+This could possibly contribute to the correlation of the two variables.
+It is important to note, however, that we cannot conclude any causal relationship between the two variables.
+Concluding whether the social cue for participation is the Forks, the Watchers, or some combination, is not possible given the data and methods used.\\
 
@@
 
-Figure \ref{fig:issue_reporters_committers_scatterplot}, does however suggest, however, that that the two variables may be related.
+Figure \ref{fig:issue_reporters_committers_scatterplot}, does however suggest, however, that that the two variables are related.
 The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation coefficient of \input{issue_reporters_committers_scatterplot-correlation}}\unskip.
@@
 \noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
-The data in Figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown.
+The data in Figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown, or an analysis of developer \textit{Contribution} percentages to a repository.
 The intent of this measurement is to show how the number of repository commits are distributed among it's committers.
-The slices represent the distribution of committers that that have committed the outer percentage amount of commits to a project, or:
+The slices represent the distribution of committers that have committed a given percentage of commits (or a \textit{Contribution} percentage, as labeled in the figure) to a project.
+For example, a small percentage of developers have contributed $95-100\%$ of the commits to a repository, so this is represented as a small slice in the figure.
+Contrast that to developers that have committed $0-5\%$ of the commits to a repository.
+This type of contribution is common, and represented as a larger slice.
+The \textit{Contribution} percentages are calculated by:
 
@@
 Despite the potential migration cost, the National Center for Biomedical Ontology,\footnote{https://github.com/ncbo} the National Cancer Informatics Program,\footnote{https://github.com/ncip} NASA,\footnote{https://github.com/nasa} and the WhiteHouse\footnote{https://github.com/WhiteHouse} are just some organizations that have moved all or portions of development to GitHub.
-As these and other organizations consider the move to GitHub, the distributed and ``social'' aspects of the OSS process will continue to evolve.\\
+As these and other organizations consider the move to GitHub, the distributed and ``social'' aspects of the OSS process will continue to evolve.
 
-\noindent \textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
+According to GitHub Staff member John Britton, lowering the barrier to participation is an important cultural shift:
+\begin{quote}
+GitHub moved the open source community away from a permission culture.
+Prior to GitHub, contributing to an open source project was very difficult and often involved asking permission.
+By allowing people to fork a repository and make changes without asking for permission, GitHub has made contributing to someone else's project much easier than it has ever been.
+By submitting a Pull Request with suggested changes, project maintainers and contributors have a place to discuss changes to code that is already written and to refine it.\cite{github_interview}
+\end{quote}
+
+Spending large amounts of time and resources to participate is a detriment to knowledge-sharing in general, not just OSS development\cite{ardichvili2003motivation}.
+Because of this, we conclude that this permissive culture of knowlege-sharing that GitHub fosters represents an important change in the OSS process.
+
+\noindent \textbf{Research Question 3: Can GitHub be used for more than code artifacts?}\\
 The OMG\textsuperscript{\textregistered} standardization process is a process by which products strive to produce an interoperability standard by way of an industry standards consortium\cite{omgTechProcess}.
@@
 \section{Conclusion}
-The data suggests an evolution in process, but not necessarily a large shift.
+The data suggests an \textit{evolution} in process, but not necessarily a large shift.
 A small subset of core developers, a traditional aspect of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}, seems supported by both the summary data in Figure \ref{eq:variance}, and the distribution of committers per repository in Figure \ref{fig:committers_histogram}.
-Figure \ref{fig:committers_percentage_pie_chart} shows that a significant number of participation, albeit small, in many repositories.
-This may indicate a lowered entry barrier for participation.
+Figure \ref{fig:committers_percentage_pie_chart} shows a significant number of participators, albeit with small contribution amounts, in many repositories.
+This may indicate a lowered entry barrier for participation, and is consistent with the conclusions reached in Research Question 2.
 
-The data summary variance statement in Assertion: \ref{eq:variance} is important, as it suggests large amounts of variety in the observed variables.
-This may suggest that not all repositories on GitHub are intended to be consumable software artifacts for users, or ``consumers.'' A large number of repositories may not be intended for customer consumption, but merely as example code, experimentation, or exercises.
-More research into categorizing GitHub repositories is needed.
-This is one reason why only repositories owned by a GitHub \textit{Organization} were analyzed, as it was anticipated that this would yield a better sample of repositories for analysis.
+The data does not reflect some traditional aspects of OSS development.
+Notably, it was not observed that there were far more \emph{Issue Reporters} than code \emph{Committers}, which is contrary to what the Apache Server report concluded\cite{mockus2000case}.
+A possible explanation is the amount of variability in the data collected.
+Assertion \ref{eq:variance}, which is derived from the summary data shown in Table \ref{table:summary_stats}, suggests large amounts of variety in the observed variables.
+This may suggest a wide range of repository purposes, or specifically, that not all repositories on GitHub are intended to be consumable software artifacts for users, or ``customers.'' A large number of repositories may not be intended for customer consumption, but merely as example code, experimentation, or exercises.
+Given this, we cannot state any conclusions regarding participation rates of the \emph{Issue Reporters}.
+More research into categorizing GitHub repositories, and the role of \emph{Issue Reporters} in them, is needed.
 
-The non-code artifacts in GitHub may fall outside of the domain of software development, but it is a significant aspect of our findings.
-From government documents to software specifications, GitHub is being used and used successfully.
-More research is needed into what role GitHub will play in the work flow of these artifacts, and whether it will influence their traditional work flow, as it has for software development.
+The most notable \textit{evolution} in OSS process driven by GitHub, we conclude, is the lowered barrier to participation.
+Qualitative analysis conducted in Research Question 2 suggest a cultural shift\cite{begel2013social,github_interview} in the participation process.
+The data also supports high levels of participation, especially Figure \ref{fig:committers_percentage_pie_chart}, which shows that many repositories have large numbers of developers committing small amounts of code.
+This, we conclude, represents an important \textit{evolution} in OSS development process.
 
 \section{Acknowledgments}
-We thank Gilberto Fragoso, Ph.D., Paul Alexander, and the GitHub Staff for insight and ideas, Harold Solbrig, Craig Stancl, and Dick Hedger for guidance, and Rick Kiefer for his review.
+We thank Gilberto Fragoso, Ph.D., Paul Alexander, John Britton, and all of the GitHub Staff for interviews, insights and ideas, Harold Solbrig, Craig Stancl, and Dick Hedger for guidance, and Rick Kiefer for his review.
 
@@
 
-%\input{interview}
+\input{interview}
+\input{github_interview}
 

### 2013-04-18 code/plotter.py

@@ -35,7 +35,7 @@ class PieChart:
 
 		fig = plt.figure()
 
-		plt.pie(ranges.values(), labels=[key + "%" for key in ranges.keys()], autopct=None, shadow=True)
+		plt.pie(ranges.values(), labels=[key + "%" for key in sorted(ranges.keys(), key=lambda key: int(key.split('-')[0]))], autopct=None, shadow=True)
 
 		plt.legend(title="Contribution")
 

### 2013-04-18 paper/github_interview.tex
Fixed some typos.

@@
-\newpage
+\clearpage
 
@@
 GitHub moved the open source community away from a permission culture.
-Prior to GitHub contributing to an open source project was very difficult and often involved asking permission.
+Prior to GitHub, contributing to an open source project was very difficult and often involved asking permission.
 By allowing people to fork a repository and make changes without asking for permission, GitHub has made contributing to someone else's project much easier than it has ever been.
@@
 Software Engineering is moving more towards the development models seen in industry.
-It's much more useful to students to learn how to use the tools that they will see in industry than it is to learn how to use learning management systems and the like.
+It's much more useful for students to learn how to use the tools that they will see in industry than it is to learn how to use learning management systems and the like.
 I encourage teachers to assign project work in teams so that students gain experience working together and navigating the issues that arrive in a collaborative environment.

### 2013-04-18 paper/main.tex
Fixed some typos.

@@
 \label{sec:attributes}
-The individual \textit{Attribute} names, or the elements of $A^\prime$, are described below. For detailed summary statistics of these \textit{Attributes}, see Figure \ref{table:summary_stats}.\\\\
+The individual \textit{Attribute} names, or the elements of $A^\prime$, are described below. For detailed summary statistics of these \textit{Attributes}, see Table \ref{table:summary_stats}.\\\\
 \noindent \textbf{\textit{Issues:}}
@@
 The most evident feature of the result is the large variance.
-In fact, for the set of tested variables $A$, each variable $A_j$ is observed to have a larger standard deviation than the variable mean.
+In fact, for the set of tested variables $A$, each variable $A_j$ is observed to have a larger standard deviation than the variable mean:
 \begin{equation}
@@
 Spending large amounts of time and resources to participate is a detriment to knowledge-sharing in general, not just OSS development\cite{ardichvili2003motivation}.
-Because of this, we conclude that this permissive culture of knowlege-sharing that GitHub fosters represents an important change in the OSS process.
+Because of this, we conclude that this permissive culture of knowlege-sharing that GitHub fosters represents an important change in the OSS process.\\
 

### 2013-04-18 paper/github_interview.tex
Added correct spaces before citations.

@@
 
-\subsection*{There are many non-code artifacts being hosted on GitHub, for example, Object Management Group Specifications, Laws and Statutes, books...
+\subsection*{Q: There are many non-code artifacts being hosted on GitHub, for example, Object Management Group Specifications, Laws and Statutes, books...
 etc.

### 2013-04-18 paper/main.tex
Added correct spaces before citations.

@@
 \section{Introduction}
-Open Source Software (OSS) has fundamentally changed how we view the Software Development Process\cite{raymond1999cathedral}. OSS projects are not only viable, but successful and thriving.
+Open Source Software (OSS) has fundamentally changed how we view the Software Development Process \cite{raymond1999cathedral}. OSS projects are not only viable, but successful and thriving.
 
-A case study of the Apache Server project\cite{mockus2000case} shows that a dedicated community can produce software that rivals or exceeds commercial offerings.
+A case study of the Apache Server project \cite{mockus2000case} shows that a dedicated community can produce software that rivals or exceeds commercial offerings.
 Furthermore, it showed that even in the unstructured context of OSS, certain structures, hierarchies, and codes of conduct emerge.
@@
 GitHub users have community-visible profiles, and user actions can be tracked and followed by other community actions.
-This type of social integration of a user's identity and actions is unique to GitHub\cite{dabbish2012social}.
+This type of social integration of a user's identity and actions is unique to GitHub \cite{dabbish2012social}.
 
-Git itself, the SCM of GitHub, lends itself to the collaborative nature of GitHub by allowing for development to take place in a more distributed manner than previously available in other SCM systems\cite{spinellis2012git}.
-Git also allows for a variety of work flows\cite{chacon2009pro}.
+Git itself, the SCM of GitHub, lends itself to the collaborative nature of GitHub by allowing for development to take place in a more distributed manner than previously available in other SCM systems \cite{spinellis2012git}.
+Git also allows for a variety of work flows \cite{chacon2009pro}.
 These work flows may be tailored to the individual needs of a project.
-The Linux Kernel, for example, uses a Dictator and Lieutenants Workflow\cite{platschekfloss}, which is hierarchical in nature.
+The Linux Kernel, for example, uses a Dictator and Lieutenants Workflow \cite{platschekfloss}, which is hierarchical in nature.
 Other work flows may be more distributed, or resemble more traditional centralized SCM systems.
@@
 To better understand quantitatively the GitHub process, three hypotheses are introduced.
-The goal of these hypotheses is to bridge the gap between the more qualitative research of GitHub\cite{dabbish2012social,begel2013social} and the statistical analysis of data from a random sampling of GitHub projects.
+The goal of these hypotheses is to bridge the gap between the more qualitative research of GitHub \cite{dabbish2012social,begel2013social} and the statistical analysis of data from a random sampling of GitHub projects.
 
@@
 Each hypothesis will be backed by previous research in an attempt to show concrete ways in which GitHub may be changing the OSS development process.
-Where possible, results from previous research on the Apache Server\cite{mockus2000case} is compared, as this work closely aligns with Hypothesis 2 and 3.\\
+Where possible, results from previous research on the Apache Server \cite{mockus2000case} is compared, as this work closely aligns with Hypothesis 2 and 3.\\
 
 \noindent \textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
-In Open Source Software, a feeling of community belonging can be intrinsically motivating to developers\cite{lakhani2003hackers}.
+In Open Source Software, a feeling of community belonging can be intrinsically motivating to developers \cite{lakhani2003hackers}.
 This feeling of belonging can be expressed passively, as a \emph{watch}, or actively, as a \emph{fork} of a repository.
@@
 
-A \emph{fork} is the creation of a personal copy of a repository\cite{dabbish2012social}, from which changes may be made without impacting the original code source.
+A \emph{fork} is the creation of a personal copy of a repository \cite{dabbish2012social}, from which changes may be made without impacting the original code source.
 A \emph{fork} is an \emph{active} action, as it signifies both involvement and intent to contribute.
@@
 \noindent \textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
-In OSS, a project \emph{fork} has at times carried negative connotations, and has even been referred to as a hazzard\cite{kogut2001open}.
+In OSS, a project \emph{fork} has at times carried negative connotations, and has even been referred to as a hazzard \cite{kogut2001open}.
 In the context of Git and GitHub, however, a \emph{fork} is a positive occurrence for a project, as it signifies greater project involvement.
@@
 The Apache Project noted that a rapid response to problems can be obtained because OSS is not bound to release schedules in the same way as commercial software.
-``Patches'' may be released at any time, by any member of the community\cite{mockus2000case}.
+``Patches'' may be released at any time, by any member of the community \cite{mockus2000case}.
 ``Patches'' in the context of GitHub could be equated to \emph{Pull Requests}.
@@
 
-This hypothesis is, ultimately, an examination of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub.
+This hypothesis is, ultimately, an examination of ``Linus Law" \cite{raymond1999cathedral} in the context of GitHub.
 In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.\\
@@
 \noindent \textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude}\\
-Research into the Apache Server project observed that there were far more \emph{Issue Reporters} than there were code \emph{Committers}\cite{mockus2000case}.
+Research into the Apache Server project observed that there were far more \emph{Issue Reporters} than there were code \emph{Committers} \cite{mockus2000case}.
 The Apache Server project is an excellent case study in this type of Developer Hierarchy.
@@
 These, although backed by analyzed data where possible, are intended to show broader patterns, ideas, and motivations around the use of GitHub.
-Where possible, interviews\cite{begel2013social} were conducted to explore high level concepts and motivations.\\
+Where possible, interviews \cite{begel2013social} were conducted to explore high level concepts and motivations.\\
 
 \noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
-A small core of developers contributing to a project seems to be a well entrenched pattern of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}.
-At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers\cite{krishnamurthy2002cave}.
+A small core of developers contributing to a project seems to be a well entrenched pattern of OSS development \cite{mockus2000case,mockus2002two,krishnamurthy2002cave}.
+At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers \cite{krishnamurthy2002cave}.
 Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changed this?\\
@@
 \noindent \textbf{Research Question 2: How is GitHub changing the OSS process?}\\
-The social aspects of GitHub are an important part of the development experience\cite{dabbish2012social}.
+The social aspects of GitHub are an important part of the development experience \cite{dabbish2012social}.
 User interactions and social pressures can drive OSS development in interesting ways.
-Social aspects of OSS development have existed before GitHub, with mailing lists\cite{mockus2000case}, gift giving mechanisms\cite{bergquist2008power}, and other collaboration tools.
+Social aspects of OSS development have existed before GitHub, with mailing lists \cite{mockus2000case}, gift giving mechanisms \cite{bergquist2008power}, and other collaboration tools.
 What is GitHub doing to facilitate change in OSS development, and how does it see OSS development changing moving forward?\\
@@
 The Object Management Group\textsuperscript{\textregistered}\footnote{http://www.omg.org} is a computer industry consortium focused on technology standards.
-Potential standard submission teams must go through a vetting process\cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard.
+Potential standard submission teams must go through a vetting process \cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard.
 Gathering and recording this feedback is a challenge, but GitHub's issue tracking system may be able to streamline the process.
@@
 
-Matplotlib\cite{Hunter2007}, a 2D graphics package for Python, was used to render graphs and charts.
-NumPy and SciPy\cite{scipy} were used for data analysis calculations, such as correlation coefficients.
+Matplotlib \cite{Hunter2007}, a 2D graphics package for Python, was used to render graphs and charts.
+NumPy and SciPy \cite{scipy} were used for data analysis calculations, such as correlation coefficients.
 
@@
 
-The sample ($N = 1000$) GitHub repositories were analyzed, out of a total of 5.7 million\cite{githubPress}.
+The sample ($N = 1000$) GitHub repositories were analyzed, out of a total of 5.7 million \cite{githubPress}.
 
@@
 \noindent \textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
-A correlation coefficient of \input{watcher_forks_scatterplot-correlation}}\unskip was observed between the number of Watchers and Forks of a repository.
+A correlation coefficient of \input{watcher_forks_scatterplot-correlation}} was observed between the number of Watchers and Forks of a repository.
 This suggests that a correlation between Watchers and Forks exists.
@@
 The data reinforces the hypothesis, as does other research in the area.
-Even though there are arguments that the number of Forks is the true measure of project success\cite{baudry2012towards}, the data suggests that the two measures are related.
+Even though there are arguments that the number of Forks is the true measure of project success \cite{baudry2012towards}, the data suggests that the two measures are related.
 Also, \textit{Dabbish et al.} have noted that the number of Watches can be a key for others to gauge interest in a project, and thus determine if they should participate \cite{dabbish2013leveraging}.
@@
 This, among other things, suggests that GitHub may be facilitating testing by making it more \emph{triable} by lowering entry barriers, and makng the process as a whole more \emph{observable}.
-This aspect is something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric\cite{phamcreating}.
+This aspect is something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric \cite{phamcreating}.
 
@@
 It is possible that other factors, such as issue prioritization, may have a greater impact on issue resolution time.
-Research does support a relationship\cite{mockus2002two} between the stated priority and fix interval.
+Research does support a relationship \cite{mockus2002two} between the stated priority and fix interval.
 As the GitHub issue tracker can only support prioritization by free-text \textit{tags} placed on issues, it is not feasible to programmatically gather issue priority metrics.\\
@@
 As shown in Figure \ref{fig:issue_reporters_committers_scatterplot}, the slope of the relationship between the two variables is less than one.
-This does not seem to completely align with findings in the Apache Server project\cite{mockus2000case}.
+This does not seem to completely align with findings in the Apache Server project \cite{mockus2000case}.
 
@@
 \noindent \textbf{Research Question 2: How is GitHub changing the OSS process?}\\
-Thung \textit{et al.} show that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enable high collaboration rates.\cite{thung2013network} In fact, the developer connectivity was shown to be higher than SourceForge and even Facebook.
-In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal.
+Thung \textit{et al.} show that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enable high collaboration rates \cite{thung2013network}.
+In fact, the developer connectivity was shown to be higher than SourceForge and even Facebook.
+In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful \cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal.
 
-Ye also observed\cite{ye2003toward} that the contribution environment, or ``society,'' is important in providing a framework for developer trust building:
+Ye also observed \cite{ye2003toward} that the contribution environment, or ``society,'' is important in providing a framework for developer trust building:
 \begin{quote}
@@
 The move to Git from a centralized version control system (such as SVN or CVS) is not without its challenges.
-Differences in philosophy and nomenclature may place added strain on developers\cite{bird2009promises}.
-Also, it is not to be assumed that process change is always welcome\cite{de2009software}, even if it is an improvement.
+Differences in philosophy and nomenclature may place added strain on developers \cite{bird2009promises}.
+Also, it is not to be assumed that process change is always welcome \cite{de2009software}, even if it is an improvement.
 Developers and organizations have existing processes, SCM tools, and workflow patterns -- which all change at a cost.
@@
 By allowing people to fork a repository and make changes without asking for permission, GitHub has made contributing to someone else's project much easier than it has ever been.
-By submitting a Pull Request with suggested changes, project maintainers and contributors have a place to discuss changes to code that is already written and to refine it.\cite{github_interview}
+By submitting a Pull Request with suggested changes, project maintainers and contributors have a place to discuss changes to code that is already written and to refine it.
+\cite{github_interview}
 \end{quote}
 
-Spending large amounts of time and resources to participate is a detriment to knowledge-sharing in general, not just OSS development\cite{ardichvili2003motivation}.
+Spending large amounts of time and resources to participate is a detriment to knowledge-sharing in general, not just OSS development \cite{ardichvili2003motivation}.
 Because of this, we conclude that this permissive culture of knowlege-sharing that GitHub fosters represents an important change in the OSS process.\\
@@
 \noindent \textbf{Research Question 3: Can GitHub be used for more than code artifacts?}\\
-The OMG\textsuperscript{\textregistered} standardization process is a process by which products strive to produce an interoperability standard by way of an industry standards consortium\cite{omgTechProcess}.
+The OMG\textsuperscript{\textregistered} standardization process is a process by which products strive to produce an interoperability standard by way of an industry standards consortium \cite{omgTechProcess}.
 The standardization process itself aims to, at its conclusion, produce a set of model artifacts containing sufficient detail such that when interpreted and implemented, produce interoperable software.
 At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process.
-If so, of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,\cite{sommerville1998requirements} we can show that GitHub is an appropriate forum for many of them.
+If so, of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management, \cite{sommerville1998requirements} we can show that GitHub is an appropriate forum for many of them.
 
-As a case study, the CTS2 OMG\textsuperscript{\textregistered} Specification\cite{cts2} used a GitHub issue tracker\footnote{https://github.com/cts2/cts2-specification/issues} to track specification changes.
+As a case study, the CTS2 OMG\textsuperscript{\textregistered} Specification \cite{cts2} used a GitHub issue tracker\footnote{https://github.com/cts2/cts2-specification/issues} to track specification changes.
 Elicitation, Analysis and Negotiation were natural fits for the issue tracker.
 This was chosen in part because the barrier to participation was very small (only a GitHub account was needed), and issue tracker itself allowed for dialogs in the form of comments.
-Because customer involvement is critical to these activities\cite{paetsch2003requirements}, ease of contribution was a main concern.
+Because customer involvement is critical to these activities \cite{paetsch2003requirements}, ease of contribution was a main concern.
 OMG\textsuperscript{\textregistered} has a predescribed format for requirements Documentation, but using the GitHub issue API\footnote{http://developer.github.com/v3/issues/}, much of this documentation can be automated.
@@
 The data suggests an \textit{evolution} in process, but not necessarily a large shift.
-A small subset of core developers, a traditional aspect of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}, seems supported by both the summary data in Figure \ref{eq:variance}, and the distribution of committers per repository in Figure \ref{fig:committers_histogram}.
+A small subset of core developers, a traditional aspect of OSS development \cite{mockus2000case,mockus2002two,krishnamurthy2002cave}, seems supported by both the summary data in Figure \ref{eq:variance}, and the distribution of committers per repository in Figure \ref{fig:committers_histogram}.
 Figure \ref{fig:committers_percentage_pie_chart} shows a significant number of participators, albeit with small contribution amounts, in many repositories.
@@
 The data does not reflect some traditional aspects of OSS development.
-Notably, it was not observed that there were far more \emph{Issue Reporters} than code \emph{Committers}, which is contrary to what the Apache Server report concluded\cite{mockus2000case}.
+Notably, it was not observed that there were far more \emph{Issue Reporters} than code \emph{Committers}, which is contrary to what the Apache Server report concluded \cite{mockus2000case}.
 A possible explanation is the amount of variability in the data collected.
@@
 The most notable \textit{evolution} in OSS process driven by GitHub, we conclude, is the lowered barrier to participation.
-Qualitative analysis conducted in Research Question 2 suggest a cultural shift\cite{begel2013social,github_interview} in the participation process.
+Qualitative analysis conducted in Research Question 2 suggest a cultural shift \cite{begel2013social,github_interview} in the participation process.
 The data also supports high levels of participation, especially Figure \ref{fig:committers_percentage_pie_chart}, which shows that many repositories have large numbers of developers committing small amounts of code.

### 2013-04-18 paper/main.tex
Change citations to non-breaking spaces.

@@
 \usepackage{mathrsfs}
-%\setlength{\parskip}{\baselineskip}
-%\setlength{\parindent}{0pt}
 
@@
 \section{Introduction}
-Open Source Software (OSS) has fundamentally changed how we view the Software Development Process \cite{raymond1999cathedral}. OSS projects are not only viable, but successful and thriving.
+Open Source Software (OSS) has fundamentally changed how we view the Software Development Process~\cite{raymond1999cathedral}. OSS projects are not only viable, but successful and thriving.
 
-A case study of the Apache Server project \cite{mockus2000case} shows that a dedicated community can produce software that rivals or exceeds commercial offerings.
+A case study of the Apache Server project~\cite{mockus2000case} shows that a dedicated community can produce software that rivals or exceeds commercial offerings.
 Furthermore, it showed that even in the unstructured context of OSS, certain structures, hierarchies, and codes of conduct emerge.
@@
 GitHub users have community-visible profiles, and user actions can be tracked and followed by other community actions.
-This type of social integration of a user's identity and actions is unique to GitHub \cite{dabbish2012social}.
+This type of social integration of a user's identity and actions is unique to GitHub~\cite{dabbish2012social}.
 
-Git itself, the SCM of GitHub, lends itself to the collaborative nature of GitHub by allowing for development to take place in a more distributed manner than previously available in other SCM systems \cite{spinellis2012git}.
-Git also allows for a variety of work flows \cite{chacon2009pro}.
+Git itself, the SCM of GitHub, lends itself to the collaborative nature of GitHub by allowing for development to take place in a more distributed manner than previously available in other SCM systems~\cite{spinellis2012git}.
+Git also allows for a variety of work flows~\cite{chacon2009pro}.
 These work flows may be tailored to the individual needs of a project.
-The Linux Kernel, for example, uses a Dictator and Lieutenants Workflow \cite{platschekfloss}, which is hierarchical in nature.
+The Linux Kernel, for example, uses a Dictator and Lieutenants Workflow~\cite{platschekfloss}, which is hierarchical in nature.
 Other work flows may be more distributed, or resemble more traditional centralized SCM systems.
@@
 To better understand quantitatively the GitHub process, three hypotheses are introduced.
-The goal of these hypotheses is to bridge the gap between the more qualitative research of GitHub \cite{dabbish2012social,begel2013social} and the statistical analysis of data from a random sampling of GitHub projects.
+The goal of these hypotheses is to bridge the gap between the more qualitative research of GitHub~\cite{dabbish2012social,begel2013social} and the statistical analysis of data from a random sampling of GitHub projects.
 
@@
 Each hypothesis will be backed by previous research in an attempt to show concrete ways in which GitHub may be changing the OSS development process.
-Where possible, results from previous research on the Apache Server \cite{mockus2000case} is compared, as this work closely aligns with Hypothesis 2 and 3.\\
+Where possible, results from previous research on the Apache Server~\cite{mockus2000case} is compared, as this work closely aligns with Hypothesis 2 and 3.\\
 
 \noindent \textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
-In Open Source Software, a feeling of community belonging can be intrinsically motivating to developers \cite{lakhani2003hackers}.
+In Open Source Software, a feeling of community belonging can be intrinsically motivating to developers~\cite{lakhani2003hackers}.
 This feeling of belonging can be expressed passively, as a \emph{watch}, or actively, as a \emph{fork} of a repository.
@@
 
-A \emph{fork} is the creation of a personal copy of a repository \cite{dabbish2012social}, from which changes may be made without impacting the original code source.
+A \emph{fork} is the creation of a personal copy of a repository~\cite{dabbish2012social}, from which changes may be made without impacting the original code source.
 A \emph{fork} is an \emph{active} action, as it signifies both involvement and intent to contribute.
@@
 \noindent \textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
-In OSS, a project \emph{fork} has at times carried negative connotations, and has even been referred to as a hazzard \cite{kogut2001open}.
+In OSS, a project \emph{fork} has at times carried negative connotations, and has even been referred to as a hazzard~\cite{kogut2001open}.
 In the context of Git and GitHub, however, a \emph{fork} is a positive occurrence for a project, as it signifies greater project involvement.
@@
 The Apache Project noted that a rapid response to problems can be obtained because OSS is not bound to release schedules in the same way as commercial software.
-``Patches'' may be released at any time, by any member of the community \cite{mockus2000case}.
+``Patches'' may be released at any time, by any member of the community~\cite{mockus2000case}.
 ``Patches'' in the context of GitHub could be equated to \emph{Pull Requests}.
@@
 
-This hypothesis is, ultimately, an examination of ``Linus Law" \cite{raymond1999cathedral} in the context of GitHub.
+This hypothesis is, ultimately, an examination of ``Linus Law"~\cite{raymond1999cathedral} in the context of GitHub.
 In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.\\
@@
 \noindent \textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude}\\
-Research into the Apache Server project observed that there were far more \emph{Issue Reporters} than there were code \emph{Committers} \cite{mockus2000case}.
+Research into the Apache Server project observed that there were far more \emph{Issue Reporters} than there were code \emph{Committers}~\cite{mockus2000case}.
 The Apache Server project is an excellent case study in this type of Developer Hierarchy.
@@
 These, although backed by analyzed data where possible, are intended to show broader patterns, ideas, and motivations around the use of GitHub.
-Where possible, interviews \cite{begel2013social} were conducted to explore high level concepts and motivations.\\
+Where possible, interviews~\cite{begel2013social} were conducted to explore high level concepts and motivations.\\
 
 \noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
-A small core of developers contributing to a project seems to be a well entrenched pattern of OSS development \cite{mockus2000case,mockus2002two,krishnamurthy2002cave}.
-At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers \cite{krishnamurthy2002cave}.
+A small core of developers contributing to a project seems to be a well entrenched pattern of OSS development~\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}.
+At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers~\cite{krishnamurthy2002cave}.
 Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changed this?\\
@@
 \noindent \textbf{Research Question 2: How is GitHub changing the OSS process?}\\
-The social aspects of GitHub are an important part of the development experience \cite{dabbish2012social}.
+The social aspects of GitHub are an important part of the development experience~\cite{dabbish2012social}.
 User interactions and social pressures can drive OSS development in interesting ways.
-Social aspects of OSS development have existed before GitHub, with mailing lists \cite{mockus2000case}, gift giving mechanisms \cite{bergquist2008power}, and other collaboration tools.
+Social aspects of OSS development have existed before GitHub, with mailing lists~\cite{mockus2000case}, gift giving mechanisms~\cite{bergquist2008power}, and other collaboration tools.
 What is GitHub doing to facilitate change in OSS development, and how does it see OSS development changing moving forward?\\
@@
 The Object Management Group\textsuperscript{\textregistered}\footnote{http://www.omg.org} is a computer industry consortium focused on technology standards.
-Potential standard submission teams must go through a vetting process \cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard.
+Potential standard submission teams must go through a vetting process~\cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard.
 Gathering and recording this feedback is a challenge, but GitHub's issue tracking system may be able to streamline the process.
@@
 
-Matplotlib \cite{Hunter2007}, a 2D graphics package for Python, was used to render graphs and charts.
-NumPy and SciPy \cite{scipy} were used for data analysis calculations, such as correlation coefficients.
+Matplotlib~\cite{Hunter2007}, a 2D graphics package for Python, was used to render graphs and charts.
+NumPy and SciPy~\cite{scipy} were used for data analysis calculations, such as correlation coefficients.
 
@@
 
-All analyzed \textit{Respositories} can be thought of as the Indexed Family of \textit{Attribute} Sets, or a set of sets of \textit{Attributes}.
+All analyzed \textit{Repositories} can be thought of as the Indexed Family of \textit{Attribute} Sets, or a set of sets of \textit{Attributes}.
 Let the set of these sets of \textit{Attributes} be called $S$, or the \textit{Repository Sample}.
@@
 
-The sample ($N = 1000$) GitHub repositories were analyzed, out of a total of 5.7 million \cite{githubPress}.
+The sample ($N = 1000$) GitHub repositories were analyzed, out of a total of 5.7 million~\cite{githubPress}.
 
@@
 The data reinforces the hypothesis, as does other research in the area.
-Even though there are arguments that the number of Forks is the true measure of project success \cite{baudry2012towards}, the data suggests that the two measures are related.
-Also, \textit{Dabbish et al.} have noted that the number of Watches can be a key for others to gauge interest in a project, and thus determine if they should participate \cite{dabbish2013leveraging}.
+Even though there are arguments that the number of Forks is the true measure of project success~\cite{baudry2012towards}, the data suggests that the two measures are related.
+Also, \textit{Dabbish et al.} have noted that the number of Watches can be a key for others to gauge interest in a project, and thus determine if they should participate~\cite{dabbish2013leveraging}.
 If the number of Watchers of a repository serves as a social cue for participation, even ``passive'' involvement with a project (Watching) could influence the ``active'' participation (Forking).
@@
 This, among other things, suggests that GitHub may be facilitating testing by making it more \emph{triable} by lowering entry barriers, and makng the process as a whole more \emph{observable}.
-This aspect is something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric \cite{phamcreating}.
+This aspect is something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric~\cite{phamcreating}.
 
@@
 It is possible that other factors, such as issue prioritization, may have a greater impact on issue resolution time.
-Research does support a relationship \cite{mockus2002two} between the stated priority and fix interval.
+Research does support a relationship~\cite{mockus2002two} between the stated priority and fix interval.
 As the GitHub issue tracker can only support prioritization by free-text \textit{tags} placed on issues, it is not feasible to programmatically gather issue priority metrics.\\
@@
 As shown in Figure \ref{fig:issue_reporters_committers_scatterplot}, the slope of the relationship between the two variables is less than one.
-This does not seem to completely align with findings in the Apache Server project \cite{mockus2000case}.
+This does not seem to completely align with findings in the Apache Server project~\cite{mockus2000case}.
 
@@
 \noindent \textbf{Research Question 2: How is GitHub changing the OSS process?}\\
-Thung \textit{et al.} show that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enable high collaboration rates \cite{thung2013network}.
+Thung \textit{et al.} show that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enable high collaboration rates~\cite{thung2013network}.
 In fact, the developer connectivity was shown to be higher than SourceForge and even Facebook.
-In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful \cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal.
+In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful~\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal.
 
-Ye also observed \cite{ye2003toward} that the contribution environment, or ``society,'' is important in providing a framework for developer trust building:
+Ye also observed~\cite{ye2003toward} that the contribution environment, or ``society,'' is important in providing a framework for developer trust building:
 \begin{quote}
@@
 The move to Git from a centralized version control system (such as SVN or CVS) is not without its challenges.
-Differences in philosophy and nomenclature may place added strain on developers \cite{bird2009promises}.
-Also, it is not to be assumed that process change is always welcome \cite{de2009software}, even if it is an improvement.
+Differences in philosophy and nomenclature may place added strain on developers~\cite{bird2009promises}.
+Also, it is not to be assumed that process change is always welcome~\cite{de2009software}, even if it is an improvement.
 Developers and organizations have existing processes, SCM tools, and workflow patterns -- which all change at a cost.
@@
 By allowing people to fork a repository and make changes without asking for permission, GitHub has made contributing to someone else's project much easier than it has ever been.
-By submitting a Pull Request with suggested changes, project maintainers and contributors have a place to discuss changes to code that is already written and to refine it.
-\cite{github_interview}
+By submitting a Pull Request with suggested changes, project maintainers and contributors have a place to discuss changes to code that is already written and to refine it.~\cite{github_interview}
 \end{quote}
 
-Spending large amounts of time and resources to participate is a detriment to knowledge-sharing in general, not just OSS development \cite{ardichvili2003motivation}.
+Spending large amounts of time and resources to participate is a detriment to knowledge-sharing in general, not just OSS development~\cite{ardichvili2003motivation}.
 Because of this, we conclude that this permissive culture of knowlege-sharing that GitHub fosters represents an important change in the OSS process.\\
@@
 \noindent \textbf{Research Question 3: Can GitHub be used for more than code artifacts?}\\
-The OMG\textsuperscript{\textregistered} standardization process is a process by which products strive to produce an interoperability standard by way of an industry standards consortium \cite{omgTechProcess}.
+The OMG\textsuperscript{\textregistered} standardization process is a process by which products strive to produce an interoperability standard by way of an industry standards consortium~\cite{omgTechProcess}.
 The standardization process itself aims to, at its conclusion, produce a set of model artifacts containing sufficient detail such that when interpreted and implemented, produce interoperable software.
 At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process.
-If so, of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management, \cite{sommerville1998requirements} we can show that GitHub is an appropriate forum for many of them.
+If so, of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,~\cite{sommerville1998requirements} we can show that GitHub is an appropriate forum for many of them.
 
-As a case study, the CTS2 OMG\textsuperscript{\textregistered} Specification \cite{cts2} used a GitHub issue tracker\footnote{https://github.com/cts2/cts2-specification/issues} to track specification changes.
+As a case study, the CTS2 OMG\textsuperscript{\textregistered} Specification~\cite{cts2} used a GitHub issue tracker\footnote{https://github.com/cts2/cts2-specification/issues} to track specification changes.
 Elicitation, Analysis and Negotiation were natural fits for the issue tracker.
 This was chosen in part because the barrier to participation was very small (only a GitHub account was needed), and issue tracker itself allowed for dialogs in the form of comments.
-Because customer involvement is critical to these activities \cite{paetsch2003requirements}, ease of contribution was a main concern.
+Because customer involvement is critical to these activities~\cite{paetsch2003requirements}, ease of contribution was a main concern.
 OMG\textsuperscript{\textregistered} has a predescribed format for requirements Documentation, but using the GitHub issue API\footnote{http://developer.github.com/v3/issues/}, much of this documentation can be automated.
@@
 The data suggests an \textit{evolution} in process, but not necessarily a large shift.
-A small subset of core developers, a traditional aspect of OSS development \cite{mockus2000case,mockus2002two,krishnamurthy2002cave}, seems supported by both the summary data in Figure \ref{eq:variance}, and the distribution of committers per repository in Figure \ref{fig:committers_histogram}.
+A small subset of core developers, a traditional aspect of OSS development~\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}, seems supported by both the summary data in Figure \ref{eq:variance}, and the distribution of committers per repository in Figure \ref{fig:committers_histogram}.
 Figure \ref{fig:committers_percentage_pie_chart} shows a significant number of participators, albeit with small contribution amounts, in many repositories.
@@
 The data does not reflect some traditional aspects of OSS development.
-Notably, it was not observed that there were far more \emph{Issue Reporters} than code \emph{Committers}, which is contrary to what the Apache Server report concluded \cite{mockus2000case}.
+Notably, it was not observed that there were far more \emph{Issue Reporters} than code \emph{Committers}, which is contrary to what the Apache Server report concluded~\cite{mockus2000case}.
 A possible explanation is the amount of variability in the data collected.
@@
 The most notable \textit{evolution} in OSS process driven by GitHub, we conclude, is the lowered barrier to participation.
-Qualitative analysis conducted in Research Question 2 suggest a cultural shift \cite{begel2013social,github_interview} in the participation process.
+Qualitative analysis conducted in Research Question 2 suggest a cultural shift~\cite{begel2013social,github_interview} in the participation process.
 The data also supports high levels of participation, especially Figure \ref{fig:committers_percentage_pie_chart}, which shows that many repositories have large numbers of developers committing small amounts of code.

### 2013-04-18 paper/main.tex
Fixed some typos.

@@
 Only repositories owned by a GitHub \textit{Organization} were considered.
-GitHub \textit{Organizations} are group-owned accounts\footnote{https://github.com/blog/674-introducing-organizations}, as opposed to normal individual user-owned accounts.
+GitHub \textit{Organizations} are group-owned accounts,\footnote{https://github.com/blog/674-introducing-organizations} as opposed to normal individual user-owned accounts.
 \textit{Organization} owned repositories were chosen to increase the likelihood of analyzing high-participation projects.
@@
 each individual attribute is processed via $g(\alpha)$ and will result in the output of the given aggregate function.
-The result of each aggregate function when applied to the entire sample set of \textit{Respositories} of Table \ref{table:summary_stats}.
+The result of each aggregate function when applied to the entire sample set of \textit{Repositories} of Table \ref{table:summary_stats}.
 
@@
 \noindent \textbf{\textit{Issue Close Time:}}
-The average time taken to close a given issue in a repository. Non-\textsc{closed} issues were not counted.\\
+The average time (in Days) taken to close a given issue in a repository. Non-\textsc{closed} issues were not counted.\\
 
@@
 
-Table \ref{table:summary_stats} describes summary statistics of the repository variables above.
+Table \ref{table:summary_stats} describes summary statistics of the repository variables described in Section \ref{sec:attributes}.
 The most evident feature of the result is the large variance.
@@
 
-The sample ($N = 1000$) GitHub repositories were analyzed, out of a total of 5.7 million~\cite{githubPress}.
+The sample ($N = 1000$) GitHub repositories were selected, at random, out of a total of 5.7 million~\cite{githubPress} public GitHub repositories.
 
@@
 \includegraphics[height=3in,width=3in]{images/issue_close_time_forks_scatterplot.png}
-\caption{Repostitory issue close time and number of forks}
+\caption{Repository issue close time and number of forks}
 \label{fig:issue_close_time_forks_scatterplot}

### 2013-04-19 paper/main.tex
Added non-breaking spaces before figures.

@@
 \textit{Organization} owned repositories were chosen to increase the likelihood of analyzing high-participation projects.
-The selection process used for choosing the repository sample is described in Algorithm \ref{algo:data_collection_algo}.
+The selection process used for choosing the repository sample is described in Algorithm~\ref{algo:data_collection_algo}.
 
@@
 Results were stored in a MySQL relational database management system.
-The database schema to store the data is described in Figure \ref{fig:er_diagram}.
+The database schema to store the data is described in Figure~\ref{fig:er_diagram}.
 
@@
 An \textit{Attribute} is a metric of interest.
-See Section \ref{sec:attributes} for further information on individual \textit{Attributes}.
+See Section~\ref{sec:attributes} for further information on individual \textit{Attributes}.
 For analysis purposes, these metrics of interest are used as an Indexing Set $A^\prime$, and have labels as follows:
@@
 This analysis is applied to all repositories in the sample $S$ for which data was collected.
-Each attribute is analyzed independently, so given:
+Each attribute $\alpha$ is analyzed independently, so given:
 \[ g(\alpha) =f \left( \bigcup\limits_{r = 0}^{|S|}  A_{r,\alpha} \right) \]
 each individual attribute is processed via $g(\alpha)$ and will result in the output of the given aggregate function.
-The result of each aggregate function when applied to the entire sample set of \textit{Repositories} of Table \ref{table:summary_stats}.
+The result of each aggregate function when applied to the entire sample set of \textit{Repositories} is shown in Table~\ref{table:summary_stats}.
 
@@
 \label{sec:attributes}
-The individual \textit{Attribute} names, or the elements of $A^\prime$, are described below. For detailed summary statistics of these \textit{Attributes}, see Table \ref{table:summary_stats}.\\\\
+The individual \textit{Attribute} names, or the elements of $A^\prime$, are described below. For detailed summary statistics of these \textit{Attributes}, see Table~\ref{table:summary_stats}.\\\\
 \noindent \textbf{\textit{Issues:}}
@@
 
-Table \ref{table:summary_stats} describes summary statistics of the repository variables described in Section \ref{sec:attributes}.
+Table~\ref{table:summary_stats} describes summary statistics of the repository variables described in Section~\ref{sec:attributes}.
 The most evident feature of the result is the large variance.
@@
 Data does not support an exponentially increasing number of distinct Issue Reporters as compared to distinct Committers. 
-As shown in Figure \ref{fig:issue_reporters_committers_scatterplot}, the slope of the relationship between the two variables is less than one.
+As shown in Figure~\ref{fig:issue_reporters_committers_scatterplot}, the slope of the relationship between the two variables is less than one.
 This does not seem to completely align with findings in the Apache Server project~\cite{mockus2000case}.
@@
 
-Figure \ref{fig:issue_reporters_committers_scatterplot}, does however suggest, however, that that the two variables are related.
+Figure~\ref{fig:issue_reporters_committers_scatterplot}, does however suggest, however, that that the two variables are related.
 The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation coefficient of \input{issue_reporters_committers_scatterplot-correlation}}\unskip.
@@
 \noindent \textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
-The data in Figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown, or an analysis of developer \textit{Contribution} percentages to a repository.
+The data in Figure~\ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown, or an analysis of developer \textit{Contribution} percentages to a repository.
 The intent of this measurement is to show how the number of repository commits are distributed among it's committers.
@@
 
-To find the percentages, or set $P$ of Equation \ref{eq:commit_percentage}, the commits by each user $C_u$ are divided by the total number of commits $|\mathscr{C}|$, yielding the set of contribution percentages $P$ for the repository.
+To find the percentages, or set $P$ of Equation~\ref{eq:commit_percentage}, the commits by each user $C_u$ are divided by the total number of commits $|\mathscr{C}|$, yielding the set of contribution percentages $P$ for the repository.
 Elements of $P$ are then assigned to ranges $0-5\%$, $5-10\%$, $10-50\%$, $50-95\%$, and $95-100\%$ (with shared values inclusive to the larger range).
-The distribution of this range assignments is plotted on Figure \ref{fig:committers_percentage_pie_chart}.
+The distribution of this range assignments is plotted on Figure~\ref{fig:committers_percentage_pie_chart}.
 
@@
 
-When looking strictly at the number of committers to a repository, however, the notion of a small set of core committers is strongly supported, as the vast majority of repositories have fewer than 10 committers (Figure \ref{fig:committers_histogram}).\\
+When looking strictly at the number of committers to a repository, however, the notion of a small set of core committers is strongly supported, as the vast majority of repositories have fewer than 10 committers (Figure~\ref{fig:committers_histogram}).\\
 
@@
 
-Ye also observed~\cite{ye2003toward} that the contribution environment, or ``society,'' is important in providing a framework for developer trust building:
+Ye also observed that the contribution environment, or ``society,'' is important in providing a framework for developer trust building:
 \begin{quote}
-Only in a society where technical supremacy is highly appreciated can developers acquire good reputations among their peers by displaying their skills through free distribution, and often wider acceptance, of their systems.The good reputation attracts attention, trust, and cooperation from others and lays the foundation for advancing the original developers agenda and the establishment and development of OSS communities.
+Only in a society where technical supremacy is highly appreciated can developers acquire good reputations among their peers by displaying their skills through free distribution, and often wider acceptance, of their systems.The good reputation attracts attention, trust, and cooperation from others and lays the foundation for advancing the original developers agenda and the establishment and development of OSS communities.~\cite{ye2003toward}
 \end{quote}
@@
 The data suggests an \textit{evolution} in process, but not necessarily a large shift.
-A small subset of core developers, a traditional aspect of OSS development~\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}, seems supported by both the summary data in Figure \ref{eq:variance}, and the distribution of committers per repository in Figure \ref{fig:committers_histogram}.
-Figure \ref{fig:committers_percentage_pie_chart} shows a significant number of participators, albeit with small contribution amounts, in many repositories.
+A small subset of core developers, a traditional aspect of OSS development~\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}, seems supported by both the summary data in Figure~\ref{eq:variance}, and the distribution of committers per repository in Figure~\ref{fig:committers_histogram}.
+Figure~\ref{fig:committers_percentage_pie_chart} shows a significant number of participators, albeit with small contribution amounts, in many repositories.
 This may indicate a lowered entry barrier for participation, and is consistent with the conclusions reached in Research Question 2.
@@
 A possible explanation is the amount of variability in the data collected.
-Assertion \ref{eq:variance}, which is derived from the summary data shown in Table \ref{table:summary_stats}, suggests large amounts of variety in the observed variables.
+Assertion~\ref{eq:variance}, which is derived from the summary data shown in Table~\ref{table:summary_stats}, suggests large amounts of variety in the observed variables.
 This may suggest a wide range of repository purposes, or specifically, that not all repositories on GitHub are intended to be consumable software artifacts for users, or ``customers.'' A large number of repositories may not be intended for customer consumption, but merely as example code, experimentation, or exercises.
@@
 Qualitative analysis conducted in Research Question 2 suggest a cultural shift~\cite{begel2013social,github_interview} in the participation process.
-The data also supports high levels of participation, especially Figure \ref{fig:committers_percentage_pie_chart}, which shows that many repositories have large numbers of developers committing small amounts of code.
+The data also supports high levels of participation, especially Figure~\ref{fig:committers_percentage_pie_chart}, which shows that many repositories have large numbers of developers committing small amounts of code.
 This, we conclude, represents an important \textit{evolution} in OSS development process.

### 2013-04-20 paper/main.tex
Updated the conclusion.

@@
 Because customer involvement is critical to these activities~\cite{paetsch2003requirements}, ease of contribution was a main concern.
-OMG\textsuperscript{\textregistered} has a predescribed format for requirements Documentation, but using the GitHub issue API\footnote{http://developer.github.com/v3/issues/}, much of this documentation can be automated.
+OMG\textsuperscript{\textregistered} has a predescribed format for requirements Documentation, but using the GitHub issue API,\footnote{http://developer.github.com/v3/issues/} much of this documentation can be automated.
 Validation and Management are areas of requirements engineering that GitHub has interesting solutions for.
@@
 \section{Conclusion}
-The data suggests an \textit{evolution} in process, but not necessarily a large shift.
-A small subset of core developers, a traditional aspect of OSS development~\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}, seems supported by both the summary data in Figure~\ref{eq:variance}, and the distribution of committers per repository in Figure~\ref{fig:committers_histogram}.
+The analysis suggests an \textit{evolution} in process, but not necessarily a large shift.
+We classify this as an \textit{evolution} because although many of the traditional aspects of the OSS development process remain, there are notable differences.
+For example, a small subset of core developers, a traditional aspect of OSS development~\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}, seems supported by both the summary data in Figure~\ref{eq:variance}, and the distribution of committers per repository in Figure~\ref{fig:committers_histogram}.
 Figure~\ref{fig:committers_percentage_pie_chart} shows a significant number of participators, albeit with small contribution amounts, in many repositories.
-This may indicate a lowered entry barrier for participation, and is consistent with the conclusions reached in Research Question 2.
+This may indicate a lowered entry barrier for participation, and is consistent with the conclusions reached in Research Questions 1 and 2.
 
-The data does not reflect some traditional aspects of OSS development.
-Notably, it was not observed that there were far more \emph{Issue Reporters} than code \emph{Committers}, which is contrary to what the Apache Server report concluded~\cite{mockus2000case}.
+The data does not, however, reflect some other traditional aspects of OSS development.
+Notably, it was observed that data gathered for Hypothesis 2 and Hypothesis 3 seemed contrary to what the Apache Server report concluded~\cite{mockus2000case}.
 A possible explanation is the amount of variability in the data collected.
@@
 This may suggest a wide range of repository purposes, or specifically, that not all repositories on GitHub are intended to be consumable software artifacts for users, or ``customers.'' A large number of repositories may not be intended for customer consumption, but merely as example code, experimentation, or exercises.
-Given this, we cannot state any conclusions regarding participation rates of the \emph{Issue Reporters}.
-More research into categorizing GitHub repositories, and the role of \emph{Issue Reporters} in them, is needed.
+Given this, and specific reasons outlined in the hypotheses themselves, we cannot state any definitive conclusions regarding these two hypotheses.
+More research into categorizing GitHub repositories is needed.
 
-The most notable \textit{evolution} in OSS process driven by GitHub, we conclude, is the lowered barrier to participation.
-Qualitative analysis conducted in Research Question 2 suggest a cultural shift~\cite{begel2013social,github_interview} in the participation process.
+The most notable \textit{difference} in OSS process driven by GitHub, considering both the quantitative and qualitative analysis, is the lowered barrier to participation.
+Qualitative analysis conducted in Research Question 2 suggests a cultural shift~\cite{begel2013social,github_interview} in the participation process.
 The data also supports high levels of participation, especially Figure~\ref{fig:committers_percentage_pie_chart}, which shows that many repositories have large numbers of developers committing small amounts of code.
-This, we conclude, represents an important \textit{evolution} in OSS development process.
+This, we conclude, represents an important \textit{evolution} in the OSS development process.
 
 \section{Acknowledgments}
-We thank Gilberto Fragoso, Ph.D., Paul Alexander, John Britton, and all of the GitHub Staff for interviews, insights and ideas, Harold Solbrig, Craig Stancl, and Dick Hedger for guidance, and Rick Kiefer for his review.
+We thank Gilberto Fragoso, Ph.D.
+(NIH/NCI), Paul Alexander (NCBO), John Britton (GitHub), and all of the GitHub Staff for interviews, insights and ideas, Harold Solbrig, Craig Stancl, and Dick Hedger for guidance, and Rick Kiefer for his review.
 

### 2013-04-22 paper/main.tex
Fixed a few tense inconsistencies.

@@
 Open Source Software (OSS) projects, or software projects with publicly available source code, are realizing ever more important roles in both personal and business computing.
-As such, shifts in the way OSS is developed could have impacts on both the quantity and the quality of OSS projects.
+As such, shifts in the way in which OSS is developed could have impacts on both the quantity and the quality of OSS projects.
 
@@
 
-A case study of the Apache Server project~\cite{mockus2000case} shows that a dedicated community can produce software that rivals or exceeds commercial offerings.
+A case study of the Apache Server project~\cite{mockus2000case} showed that a dedicated community can produce software that rivals or exceeds commercial offerings.
 Furthermore, it showed that even in the unstructured context of OSS, certain structures, hierarchies, and codes of conduct emerge.
@@
 These questions will draw upon data from this work, but also reference several case studies to better represent how GitHub may be changing the way in which OSS development is conducted.
-These qualitative observations, paired with the more quantitative data analysis, allows for a broad description of the way that GitHub is changing the Software Engineering landscape for OSS.
+These qualitative observations, paired with the more quantitative data analysis, allow for a broad description of the way that GitHub is changing the Software Engineering landscape for OSS.
 
@@
 \subsection{Summary Statistics}
+\label{sec:summary_statistics}
 \begin{table}[!ht]
@@
 \end{equation}
-All measured variables demonstrated this characteristic.
+All measured variables demonstrated this characteristic. We will refer to this characteristic as Summary Data Assertion \ref{eq:variance}.
 
@@
 \includegraphics[height=3in,width=3in]{images/watcher_forks_scatterplot.png}
-\caption{Repostitory watchers and forks}
+\caption{Repository watchers and forks}
 \end{figure}
@@
 Elements of $P$ are then assigned to ranges $0-5\%$, $5-10\%$, $10-50\%$, $50-95\%$, and $95-100\%$ (with shared values inclusive to the larger range).
-The distribution of this range assignments is plotted on Figure~\ref{fig:committers_percentage_pie_chart}.
+The distribution of this range assignment is plotted on Figure~\ref{fig:committers_percentage_pie_chart}.
 
@@
 \noindent \textbf{Research Question 2: How is GitHub changing the OSS process?}\\
-Thung \textit{et al.} show that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enable high collaboration rates~\cite{thung2013network}.
+Thung \textit{et al.} showed that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enabled high collaboration rates~\cite{thung2013network}.
 In fact, the developer connectivity was shown to be higher than SourceForge and even Facebook.
-In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful~\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal.
+In a recent interview with GitHub staff, the social aspects of GitHub were referred to as ``unique and powerful~\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal.
 
@@
 Spending large amounts of time and resources to participate is a detriment to knowledge-sharing in general, not just OSS development~\cite{ardichvili2003motivation}.
-Because of this, we conclude that this permissive culture of knowlege-sharing that GitHub fosters represents an important change in the OSS process.\\
+Because of this, we conclude that this permissive culture of knowledge-sharing that GitHub fosters represents an important change in the OSS process.\\
 
@@
 Elicitation, Analysis and Negotiation were natural fits for the issue tracker.
-This was chosen in part because the barrier to participation was very small (only a GitHub account was needed), and issue tracker itself allowed for dialogs in the form of comments.
+This was chosen in part because the barrier to participation was very small (only a GitHub account was needed), and the issue tracker itself allowed for dialogs in the form of comments.
 Because customer involvement is critical to these activities~\cite{paetsch2003requirements}, ease of contribution was a main concern.
@@
 A possible explanation is the amount of variability in the data collected.
-Assertion~\ref{eq:variance}, which is derived from the summary data shown in Table~\ref{table:summary_stats}, suggests large amounts of variety in the observed variables.
+Summary Data Assertion~\ref{eq:variance} (Section~\ref{sec:summary_statistics}), which is derived from the summary data shown in Table~\ref{table:summary_stats}, suggests large amounts of variety in the observed variables.
 This may suggest a wide range of repository purposes, or specifically, that not all repositories on GitHub are intended to be consumable software artifacts for users, or ``customers.'' A large number of repositories may not be intended for customer consumption, but merely as example code, experimentation, or exercises.

### 2013-04-28 code/plotter.py
Added larger labels to diagrams.

@@ -72,13 +72,13 @@ class ScatterPlot:
 		plt.scatter(x,y)
 
 		correlation_coefficient = "%.4f" % numpy.corrcoef(x,y)[0,1]
-		title('Correlation Coefficient: ' + correlation_coefficient)
+		title('Correlation Coefficient: ' + correlation_coefficient, fontsize=18)
 
 		if float(correlation_coefficient) > 0.5:
 			plt.plot(x,y,'bo',x,m*numpy.array(x)+b,'-k',linewidth=2)
 
-		plt.xlabel(self.xlabel)
-		plt.ylabel(self.ylabel)
+		plt.xlabel(self.xlabel,fontsize=18)
+		plt.ylabel(self.ylabel,fontsize=18)
 		plt.xlim(self.xrange)
 		plt.ylim(self.yrange)
 	
@@ -112,8 +112,8 @@ class Histograph:
 
 		fig = plt.figure() 
 		plt.hist(x, range=self.range)
-		plt.xlabel(self.xlabel)
-		plt.ylabel(self.ylabel)
+		plt.xlabel(self.xlabel,fontsize=18)
+		plt.ylabel(self.ylabel,fontsize=18)
 
 		F = gcf()
 		F.savefig("../paper/images/"+filename+".png")

### 2013-04-28 paper/main.tex
Finished another round of typo checks.

@@
 GitHub\footnote{https://github.com/} is a popular code hosting website that uses Git Source Code Management (SCM).\footnote{http://git-scm.com/} Along with hosting software repositories, GitHub incorporates the social aspects of development.
-GitHub users have community-visible profiles, and user actions can be tracked and followed by other community actions.
+GitHub users have community-visible profiles, and user actions can be tracked and followed by other community members.
 This type of social integration of a user's identity and actions is unique to GitHub~\cite{dabbish2012social}.
@@
 This hypothesis is, ultimately, an examination of ``Linus Law"~\cite{raymond1999cathedral} in the context of GitHub.
-In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.\\
+In other words, the more exposure the code gets (in terms of \emph{forks}), the easier bugs will be to find and fix.\\
 
@@
 User interactions and social pressures can drive OSS development in interesting ways.
-Social aspects of OSS development have existed before GitHub, with mailing lists~\cite{mockus2000case}, gift giving mechanisms~\cite{bergquist2008power}, and other collaboration tools.
-What is GitHub doing to facilitate change in OSS development, and how does it see OSS development changing moving forward?\\
+Social aspects of OSS development have existed, however, before GitHub, with mailing lists~\cite{mockus2000case}, gift giving mechanisms~\cite{bergquist2008power}, and other collaboration tools.
+What is GitHub doing to facilitate and capitalize on the social aspects of OSS development, and how might this influence the OSS development process?\\
 
@@
 Gathering and recording this feedback is a challenge, but GitHub's issue tracking system may be able to streamline the process.
-This just one example of how GitHub is being used for non-code artifacts.
+This is just one example of how GitHub is being used for non-code artifacts.
 Are organizations exploring this type of work flow?
@@
 
-To gather the metrics, the sets of all individual \textit{Attribute} across all \textit{Repositories} are processed by some aggregate function $f$.
+To analyze the metrics, the sets of all individual \textit{Attribute} across all \textit{Repositories} are processed by some aggregate function $f$.
 The input of this function is the union of all sets of a particular \textit{Attribute} over each \textit{Repository} in the sample.
@@
 
-The average is calculated using the following method.
-Let $d(x)$ be a function that converts seconds to a Date, and let $s(x)$ be a function that converts a Date to seconds.
+The average \textit{issue close time} is calculated using the following method: Let $d(x)$ be a function that converts seconds to a Date, and let $s(x)$ be a function that converts a Date to seconds.
 Let $I$ equal the Set of \textit{Issues} in a given repository, with elements comprised of tuples $(\tau, \tau^\prime)$, where $\tau$ is \textit{Issue Open Date} and $\tau^\prime$ is \textit{Issue Close Date}.
@@
 \textit{Second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all.
-\textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub.
-This, among other things, suggests that GitHub may be facilitating testing by making it more \emph{triable} by lowering entry barriers, and makng the process as a whole more \emph{observable}.
+\textit{Third}, Cultural shifts in how testing is approached may have significant impact on quality.
+Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub, which suggests that GitHub may be facilitating testing by making it more \emph{triable} by lowering entry barriers, and makng the process as a whole more \emph{observable}.
 This aspect is something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric~\cite{phamcreating}.
@@
 \includegraphics[height=3in,width=3in]{images/committers_percentage_pie_chart.png}
-\caption{Distribution of individual committer commit percentage of a repository}
+\caption{Distribution of individual committer contribution percentage to a repository}
 \label{fig:committers_percentage_pie_chart}
@@
 
-According to GitHub Staff member John Britton, lowering the barrier to participation is an important cultural shift:
+According to GitHub Education Liaison John Britton, lowering the barrier to participation is an important cultural shift:
 \begin{quote}
@@
 At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process.
-If so, of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,~\cite{sommerville1998requirements} we can show that GitHub is an appropriate forum for many of them.
+If so, of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management~\cite{sommerville1998requirements}, we can show that GitHub is an appropriate forum for many of them.
 
@@
 The data does not, however, reflect some other traditional aspects of OSS development.
-Notably, it was observed that data gathered for Hypothesis 2 and Hypothesis 3 seemed contrary to what the Apache Server report concluded~\cite{mockus2000case}.
+Notably, it was observed that data gathered for Hypothesis~2 (\textit{As the number Repository Forks increase, the Issue Resolution Time will decrease}) and Hypothesis~3 (\textit{There will be more Issue Reporters than Committers by an order of magnitude}) seemed contrary to what the Apache Server report concluded~\cite{mockus2000case}.
 A possible explanation is the amount of variability in the data collected.

### 2013-04-29 paper/main.tex
Final set of typo fixes??? Maybe.

@@
 As \emph{watches} don't necessarily signal intent to contribute, a positive direction for an OSS project is to not only increase the number of \emph{watchers}, but also the number of \emph{forks}.
-As both signify a general general interest in the project, it is expected that there will be some correlation between the two.\\
+As both signify a general interest in the project, it is expected that there will be some correlation between the two.\\
 
@@
 A small core of developers contributing to a project seems to be a well entrenched pattern of OSS development~\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}.
-At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers~\cite{krishnamurthy2002cave}.
+At the extreme of this, a study of projects on the software hosting site SourceForge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers~\cite{krishnamurthy2002cave}.
 Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changed this?\\
@@
 
+To select a repository, a random word is selected from an arbitrary list via the \textsc{random}($word\_list$) function.
+Using the GitHub search API, a list of repositories containing that word in the description is retrieved using the \textsc{github\_search}($word$) function, from which a one is randomly select using \textsc{random}($repos$).
 When a random repository has been selected, the \textsc{isOrgRepo} function will filter out all non \textit{Organization} owned repositories.
@@
 \includegraphics[height=3in,width=3in]{images/er.png}
-\caption{GitHub Date Entity Relationship diagram}
+\caption{GitHub data entity relationship diagram}
 \label{fig:er_diagram}
@@
 \noindent \textbf{\textit{Issue Close Time:}}
-The average time (in Days) taken to close a given issue in a repository. Non-\textsc{closed} issues were not counted.\\
+The average time (in days) taken to close a given issue in a repository. Non-\textsc{closed} issues were not counted.
 
-The average \textit{issue close time} is calculated using the following method: Let $d(x)$ be a function that converts seconds to a Date, and let $s(x)$ be a function that converts a Date to seconds.
+The average \textit{Issue Close Time} is calculated using the following method: Let $d(x)$ be a function that converts seconds to a Date, and let $s(x)$ be a function that converts a Date to seconds.
 Let $I$ equal the Set of \textit{Issues} in a given repository, with elements comprised of tuples $(\tau, \tau^\prime)$, where $\tau$ is \textit{Issue Open Date} and $\tau^\prime$ is \textit{Issue Close Date}.
@@
 N = 1000 Repositories
-\caption{GitHub Summary Statistics}
+\caption{GitHub summary statistics}
 \label{table:summary_stats}
@@
 \noindent \textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
-No significant correlation was found between the number of repository forks and issue resolution time.
+No significant correlation was found between the number of repository forks and issue resolution time, as shown by the data represented in Figure~\ref{fig:issue_close_time_forks_scatterplot}.
 It is important to note, however, that issue resolution time is not necessarily related to software product quality in the context of GitHub.
@@
 \textit{Second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all.
-\textit{Third}, Cultural shifts in how testing is approached may have significant impact on quality.
-Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub, which suggests that GitHub may be facilitating testing by making it more \emph{triable} by lowering entry barriers, and makng the process as a whole more \emph{observable}.
+\textit{Third}, cultural shifts in how testing is approached in GitHub may have significant impacts on software quality.
+Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub, which suggests that GitHub may be facilitating testing by making it more \emph{triable} by lowering entry barriers, and making the process as a whole more \emph{observable}.
 This aspect is something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric~\cite{phamcreating}.
@@
 \includegraphics[height=3in,width=3in]{images/issue_close_time_forks_scatterplot.png}
-\caption{Repository issue close time and number of forks}
+\caption{Repository issue close time (in days) and number of forks}
 \label{fig:issue_close_time_forks_scatterplot}
@@
 
-Where $U$ is the Set of \textit{Committers} to a given \textit{Repository}, $\mathscr{C}$ is the set of all \textit{Commits} to a \textit{Repository}, and $C_u$ denotes sets of \textit{Commits} by an individual \textit{Committers} to a \textit{Repository}, or:
+Where $U$ is the set of \textit{Committers} to a given \textit{Repository}, $\mathscr{C}$ is the set of all \textit{Commits} to a \textit{Repository}, and $C_u$ denotes sets of \textit{Commits} by an individual \textit{Committers} to a \textit{Repository}, or:
  \[ \{ C_u \}_{u \in U} \]

### 2013-04-29 paper/main.tex
Last minute touchups.

@@
 Each hypothesis will be backed by previous research in an attempt to show concrete ways in which GitHub may be changing the OSS development process.
-Where possible, results from previous research on the Apache Server~\cite{mockus2000case} is compared, as this work closely aligns with Hypothesis 2 and 3.\\
+Where possible, results from previous research on the Apache Server~\cite{mockus2000case} are compared, as this work closely aligns with Hypothesis 2 and 3.\\
 
@@
 
-\noindent \textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude}\\
+\noindent \textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude.}\\
 Research into the Apache Server project observed that there were far more \emph{Issue Reporters} than there were code \emph{Committers}~\cite{mockus2000case}.
@@
 To select a repository, a random word is selected from an arbitrary list via the \textsc{random}($word\_list$) function.
-Using the GitHub search API, a list of repositories containing that word in the description is retrieved using the \textsc{github\_search}($word$) function, from which a one is randomly select using \textsc{random}($repos$).
+Using the GitHub search API, a list of repositories containing that word in the description is retrieved using the \textsc{github\_search}($word$) function, from which one repository is randomly selected using \textsc{random}($repos$).
 When a random repository has been selected, the \textsc{isOrgRepo} function will filter out all non \textit{Organization} owned repositories.
@@
 The average \textit{Issue Close Time} is calculated using the following method: Let $d(x)$ be a function that converts seconds to a Date, and let $s(x)$ be a function that converts a Date to seconds.
-Let $I$ equal the Set of \textit{Issues} in a given repository, with elements comprised of tuples $(\tau, \tau^\prime)$, where $\tau$ is \textit{Issue Open Date} and $\tau^\prime$ is \textit{Issue Close Date}.
+Let $I$ equal the set of \textit{Issues} in a given repository, with elements comprised of tuples $(\tau, \tau^\prime)$, where $\tau$ is \textit{Issue Open Date} and $\tau^\prime$ is \textit{Issue Close Date}.
 We calculate \textit{Issue Close Time} by:

### 2013-04-29 paper/main.tex
Fixed a bad SQL statement.

@@
 
-Where $i$ = {\textit{Issues}, $c$ = {\textit{Commits}, $i^\prime$ = {\textit{Closed Issues}, $i^{\prime\prime}$ = {\textit{Open Issues}, $\Delta$ = {\textit{Issue Close Time}, $\omega$ = {\textit{Issue Creators}, and $\epsilon$ = {\textit{Committers}.
+Where $i$ = {\textit{Issues}, $c$ = {\textit{Commits}, $i^\prime$ = {\textit{Closed Issues}, $i^{\prime\prime}$ = {\textit{Open Issues}, $\Delta$ = {\textit{Issue Close Time}, $\omega$ = {\textit{Issue Reporters}, and $\epsilon$ = {\textit{Committers}.
 
@@
 
-\noindent \textbf{\textit{Issue Creators:}}
+\noindent \textbf{\textit{Issue Reporters:}}
 The number of distinct users that have opened at least one \textit{Issue}.
@@
 Data does not support an exponentially increasing number of distinct Issue Reporters as compared to distinct Committers. 
-As shown in Figure~\ref{fig:issue_reporters_committers_scatterplot}, the slope of the relationship between the two variables is less than one.
+As shown in Figure~\ref{fig:issue_reporters_committers_scatterplot}, the number of Issue Reporters is larger than the number of Committers, but not by an order of magnitude.
+Table~\ref{table:summary_stats} reinforces this, as the averages of the two variables are closer than the hypothesis states.
 This does not seem to completely align with findings in the Apache Server project~\cite{mockus2000case}.

### 2013-04-29 code/plots.py

@@ -52,7 +52,7 @@ from
         issue
     group by repository_id) issues ON issues.repository_id = commits.repository_id
 '''
-issue_reporters_committers_scatterplot = ScatterPlot(issue_reporters_committers_query, "Committers", "Issue Reoprters", (0,140), (0,140))
+issue_reporters_committers_scatterplot = ScatterPlot(issue_reporters_committers_query, "Issue Reporters", "Committers", (0,140), (0,140))
 issue_reporters_committers_scatterplot.plot("issue_reporters_committers_scatterplot")
 
 issue_close_time_forks_query = '''

### 2014-08-14 paper/main.tex
Fixed a typo.
Removed some contact info from the header of main, as well as the interviews.

@@
 \author{Kevin Peterson\\
-Department of Information Technology\\
-Clinical Informatics Support Systems\\
-Mayo Clinic\\
-200 1st Street\\
-Rochester, MN  55905\\
-\small \texttt{peterson.kevin@mayo.edu}
+\small \texttt{kevin@kevinp.me}
 }
@@
 
-Figure~\ref{fig:issue_reporters_committers_scatterplot}, does however suggest, however, that that the two variables are related.
+Figure~\ref{fig:issue_reporters_committers_scatterplot} does suggest, however, that that the two variables are related.
 The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation coefficient of \input{issue_reporters_committers_scatterplot-correlation}}\unskip.
@@
 
-\input{interview}
-\input{github_interview}
+%\input{interview}
+%\input{github_interview}
 

### 2014-08-15 paper/publish.sh
PDF commit

@@ -1,1 +1,7 @@
 latexmk main.tex -output-directory=build -pdf
+cp build/main.pdf /tmp/
+cd ..
+git checkout gh-pages
+mv /tmp/main.pdf github-process-research.pdf
+git commit -am "PDF commit"
+git checkout master

### 2014-08-15 paper/publish.sh
PDF commit

@@ -1,7 +1,9 @@
 latexmk main.tex -output-directory=build -pdf
-cp build/main.pdf /tmp/
-cd ..
-git checkout gh-pages
-mv /tmp/main.pdf github-process-research.pdf
-git commit -am "PDF commit"
-git checkout master
+if [[ $1 == "deploy" ]]; then
+	cp build/main.pdf /tmp/
+	cd ..
+	git checkout gh-pages
+	mv /tmp/main.pdf github-process-research.pdf
+	git commit -am "PDF commit"
+	git checkout master
+fi
